---
title: "Vancouver KSI: Data Check"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
- "Brice Batomen brice.kuimi@utoronto.ca"
- "Linda Rothman linda.rothman@torontomu.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../../docs/archive/check-vancouver-ksi",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Libraries

Install R packages if needed.

```{r, results = FALSE}

# Required packages
required_packages <- c(
	"rmarkdown",
	"bookdown",
	"knitr",
	"lubridate",
	"tidyverse",
	"purrr",
	"glue",
	"lubridate",
	"sf",
	"tmap",
	"leaflet",
	"leaflet.extras"
)

# Try to install packages if not installed
default_options <- options()
tryCatch(
	{
		# Disable interactivity
		options(install.packages.compile.from.source = "always")
		
		# Install package if not installed
		for (package in required_packages) {
			is_package_installed <- require(package, character.only = TRUE)
			if (!is_package_installed) {
				cat(paste0("Installing package: ", package, "\n"))
				install.packages(package)
			} else {
				cat(paste0("Package already installed: ", package, "\n"))
			}
		}
	},
	error = function(cond) {
		stop(cond)
	},
	finally = {
		options(default_options) # reset interactivity
	}
)
```

Load R libraries.

```{r}
library(ggplot2)
library(glue)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(sf)
library(tidyverse)
library(tmap)
```

# Data {.tabset}

* `vancksi_raw1`: is the file received from Brice Batomen <brice.kuimi@utoronto.ca> on July 3, 2024 via email, which was originally obtained directly from ICBC
* `vancksi_raw2`: is the file downloaded [here](https://public.tableau.com/app/profile/icbc/viz/ICBCReportedCrashes/ICBCReportedCrashes-LowerMainland) with the documentation provided at ICBC's Tableau website [here](https://public.tableau.com/app/profile/icbc/viz/ICBCReportedCrashes/ICBCReportedCrashes) on September 13, 2024

```{r}
vancksi_raw1 <- read.table("../../tmp/vanc-test.dat", sep = "|", header = T)
vancksi_raw2 <- read_delim("../../tmp/vanc-ksi-lowermainland-2024-09-13.csv", delim = "\t", locale = locale(encoding="UTF-16"))
vancbounds_raw <- read_sf("../../tmp/vanc-localareabound-2024-09-13.geojson")
```

## ICBC (Direct)

```{r}
vancksi_raw1 %>% head
```

## ICBC (Tableau)

```{r}
vancksi_raw2 %>% head
```

## Vancouver Bounds

Use the local area boundary as the City of Vancouver boundaries obtained from the open data portal [here](https://opendata.vancouver.ca/explore/dataset/local-area-boundary/information/?disjunctive.name&location=12,49.2474,-123.12402).

```{r}
vancbounds_raw %>% head
```

# Cleaning

## Vancouver Bounds

Buffer boundaries by 100 meters to capture errorneous points.

```{r}
vancbounds <- vancbounds_raw %>% st_buffer(100)
```

## ICBC (Direct)

### Convert to Spatial Data {.tabset}

```{r}
vancksi1 <- st_as_sf(
	vancksi_raw1 %>% filter(!is.na(LATITUDE) & !is.na(LONGITUDE)),
	coords = c("LONGITUDE", "LATITUDE"),
	crs = 4326
)
```

There are missing coordinates in the data.

```{r}
cat(
	"\nICBC (Direct) Missing Coordinates: ", vancksi_raw1 %>% filter(is.na(LATITUDE) | is.na(LONGITUDE)) %>% nrow,
	"\nICBC (Direct) Missing Longitudes: ", vancksi_raw1 %>% filter(is.na(LONGITUDE)) %>% nrow,
	"\nICBC (Direct) Missing Latitudes: ", vancksi_raw1 %>% filter(is.na(LATITUDE)) %>% nrow
)
```

### Crop to City Bounds

```{r}
before_crop <- vancksi1 %>% nrow
vancksi1 <- vancksi1 %>%
	st_crop(vancbounds)
after_crop <- vancksi1 %>% nrow
cat("Rows cropped: ", before_crop - after_crop)
```

### Killed or Serious Injuries Only {.tabset}

Injuries only (`Fatal injury`, `Serious injury - Overnight at hospital`) for column `INJURY_TYPE`:

```{r}
vancksi1 <- vancksi1 %>%
	filter(INJURY_TYPE %in% c(
		"Fatal injury",
		"Serious injury - Overnight at hospital"
	))
```

#### Preview

```{r}
vancksi1 %>% as_tibble %>% select(-geometry) %>% head
```

#### Details

```{r}
vancksi_raw1 %>%
	group_by(INJURY_TYPE) %>%
	count %>%
	mutate(
		perc = n / nrow(vancksi_raw1) * 100
	)
```

## ICBC (Tableau)

### Convert to Spatial Data

```{r}
vancksi2 <- st_as_sf(
	vancksi_raw2 %>% filter(!is.na(Latitude) & !is.na(Longitude)),
	coords = c("Longitude", "Latitude"),
	crs = 4326
)
```

There are missing coordinates in the data.

```{r}
cat(
	"\nICBC (Tableau) Missing Coordinates: ", vancksi_raw2 %>% filter(is.na(Latitude) | is.na(Longitude)) %>% nrow,
	"\nICBC (Tableau) Missing Longitudes: ", vancksi_raw2 %>% filter(is.na(Longitude)) %>% nrow,
	"\nICBC (Tableau) Missing Latitudes: ", vancksi_raw2 %>% filter(is.na(Latitude)) %>% nrow
)
```

### Crop to City Bounds

```{r}
before_crop <- vancksi2 %>% nrow
vancksi2 <- vancksi2 %>%
	st_crop(st_bbox(vancbounds))
after_crop <- vancksi2 %>% nrow
cat("Rows cropped: ", before_crop - after_crop)
```

### Killed or Serious Injuries Only {.tabset}

Only collisions with `CASUALTY CRASH` according to the data dictionary [here](https://public.tableau.com/app/profile/icbc/viz/ICBCReportedCrashes/ICBCReportedCrashes) for column `Crash Severity`:

```{r}
vancksi2 <- vancksi2 %>%
	filter(`Crash Severity` == "CASUALTY CRASH")
```

#### Preview

```{r}
vancksi2 %>% as_tibble %>% select(-geometry) %>% head
```

#### Details

```{r}
vancksi_raw2 %>%
	group_by(`Crash Severity`) %>%
	count %>%
	mutate(
		perc = n / nrow(vancksi_raw2) * 100
	)
```

# Exploration

## Column Differences {.tabset}

The Tableau data has 11 more columns that the direct data with more temporal details.

### Details

```{r}

cat(
	"\nColumns in ICBC (Direct): ", ncol(vancksi1),
	"\nColumns in ICBC (Tableau): ", ncol(vancksi2),
	"\nColumns difference: ", ncol(vancksi2) - ncol(vancksi1)
)

```

### Columns

Time, day, month, and year are all given in the Tableau data, while the direct data only contains the month and year.

ICBC (Direct):

```{r}
vancksi1 %>% colnames
```

ICBC (Tableau):

```{r}
vancksi2 %>% colnames
```

## Row Differences

The Tableau data has more records than the direct data.

### Details

```{r}

cat(
	"\nRows in ICBC (Direct): ", nrow(vancksi1),
	"\nRows in ICBC (Tableau): ", nrow(vancksi2),
	"\nRows difference: ", nrow(vancksi2) - nrow(vancksi1)
)

```

### Rows

ICBC (Direct):

```{r}
vancksi1 %>% as_tibble %>% select(-geometry) %>% head
```

ICBC (Tableau):

```{r}
vancksi2 %>% as_tibble %>% select(-geometry) %>% head
```

### Temporal

The direct data covers a longer time span (2010 to 2021) than the tableau data (2019 to 2023).

```{r}
cat(
	"\nICBC (Direct) Years: ", min(vancksi1$YEAR), "to", max(vancksi1$YEAR),
	"\nICBC (Tableau) Years: ", min(vancksi2$`Date Of Loss Year`), "to", max(vancksi2$`Date Of Loss Year`)
)
```

ICBC (Direct):

```{r}
vancksi1 %>%
	group_by(YEAR) %>%
	count
```

ICBC (Tableau):

```{r}
vancksi2 %>%
	group_by(`Date Of Loss Year`) %>%
	count
```

## Location {.tabset}

Each record seems to represent collision locations as there are victim counts for each row in both datasets.

### Map

```{r}
tmap_mode("view")
tm_basemap("CartoDB.Positron") +
	tm_shape(
		vancksi2 %>% filter(`Date Of Loss Year` == 2019),
		name = "ICBC (Tableau)"
	) +
	tm_dots(
		popup.vars = T
	) +
	tm_shape(
		vancksi1 %>% filter(YEAR == 2019),
		name = "ICBC (Direct)"
	) +
	tm_dots(
		col = "red",
		popup.vars = T
	)
```
