---
title: "Toronto Bikeways: Data Check"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
- "Brice Batomen brice.kuimi@utoronto.ca"
- "Linda Rothman linda.rothman@torontomu.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../../docs/archive/check-toronto-bikeways",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Libraries

Install R packages if needed.

```{r, results = FALSE}

# Required packages
required_packages <- c(
	"rmarkdown",
	"bookdown",
	"knitr",
	"lubridate",
	"tidyverse",
	"purrr",
	"glue",
	"lubridate",
	"sf",
	"tmap",
	"leaflet",
	"leaflet.extras"
)

# Try to install packages if not installed
default_options <- options()
tryCatch(
	{
		# Disable interactivity
		options(install.packages.compile.from.source = "always")
		
		# Install package if not installed
		for (package in required_packages) {
			is_package_installed <- require(package, character.only = TRUE)
			if (!is_package_installed) {
				cat(paste0("Installing package: ", package, "\n"))
				install.packages(package)
			} else {
				cat(paste0("Package already installed: ", package, "\n"))
			}
		}
	},
	error = function(cond) {
		stop(cond)
	},
	finally = {
		options(default_options) # reset interactivity
	}
)
```

Load R libraries.

```{r}
library(ggplot2)
library(glue)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(sf)
library(tidyverse)
library(tmap)
```

# Data

Read data from the `data` folder.

```{r}
ddesc <- read_csv("../../data/data.csv")
ddesc
```

## Toronto Bikeways {.tabset}

`r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(notes)`

```{r}

# Read data
toronbike_raw <- read_sf("../../data/toronto-bikeways-2024-06-02.geojson")

# Get download date
toronbike_dldate <- ddesc %>% filter(
	file == "toronto-bikeways-2024-06-02.geojson"
) %>% pull(download_date)
```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(toronbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(toronbike_raw)`
* Rows: `r nrow(toronbike_raw)`

```{r}
toronbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
toronbike_ddict <- read_csv("../../data/toronto-bikeways-2024-06-02-datadict.csv")
toronbike_ddict
```

### Details

```{r}
print(toronbike_raw)
```

### Files

The data files are available below:

- [toronto-bikeways-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-06-02.geojson)
- [toronto-bikeways-2024-06-02-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-06-02-datadict.csv)

## Verified Dates {.tabset}

`r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(notes)`

```{r}

# Read data
vdates_raw <- read_csv("../../data/verify-dates-2024-06-12.csv")

# Get download date
vdates_dldate <- ddesc %>% filter(
	file == "verify-dates-2024-06-12.csv"
) %>% pull(download_date)
```

### Data

* Columns: `r ncol(vdates_raw)`
* Rows: `r nrow(vdates_raw)`

```{r}
vdates_raw
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
vdates_ddict <- read_csv("../../data/verify-dates-2024-06-12-datadict.csv")
vdates_ddict
```

### Files

The data files are available below:

- [verify-dates-2024-06-12.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12.csv)
- [verify-dates-2024-06-12-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12-datadict.csv)

# Cleaning

## Filter Empty Types

Filter out empty install/upgrade types.

```{r}

# Filter out none or na
toronbike <- toronbike_raw %>%
	filter(
		!verify_install_type %in% c("None", NA) |
		!verify_upgrade1_type %in% c("None", NA) |
		!verify_upgrade2_type %in% c("None", NA)
	)

# Display non empty types in at least one of install or upgrade
toronbike %>%
	as_tibble %>%
	select(-geometry) %>%
	select(
		id,
		verify_install_type,
		verify_upgrade1_type,
		verify_upgrade2_type
	)
```

## Add Dates

Add cleaned post-2011 dates to verified bikeways.

```{r}

# Add cleaned post-2011 dates to bikeways
toronbike <- toronbike %>%
    left_join( # clean install dates
        vdates_raw %>%
            rename_all(~str_replace(., "verify_", "clean_install_")),
        by = join_by(verify_install_date == clean_install_date_raw)
    ) %>%
	left_join( # clean upgrade1 dates
        vdates_raw %>%
            rename_all(~str_replace(., "verify_", "clean_upgrade1_")),
        by = join_by(verify_upgrade1_date == clean_upgrade1_date_raw)
    ) %>%
	left_join( # clean upgrade2 dates
        vdates_raw %>%
            rename_all(~str_replace(., "verify_", "clean_upgrade2_")),
        by = join_by(verify_upgrade2_date == clean_upgrade2_date_raw)
    )

# Display cleaned dates columns
toronbike %>%
	as_tibble %>%
	select(-geometry) %>%
	select(id, verify_install_date, verify_upgrade1_date, verify_upgrade2_date, starts_with("clean_"))
```

## Add Quarters

Assign quarters to each bikeway date, where a value of:

* `1`: represents November (this year) to April of next year
* `2`: represents May to October of next year

```{r}

# Add quarters to bike based on clean dates
toronbike <- toronbike %>%
	mutate(
		clean_install_quarter = case_when( # install quarter
			month(clean_install_date) %in% c(11:12, 1:4) |
			(
				month(clean_install_date_start) %in% c(11:12, 1:4) &
				( # Nov to Dec of this year
					month(clean_install_date_end) %in% 11:12 &
					year(clean_install_date_end) == year(clean_install_date_start)
				) |
				( # Jan to Apr of this or next year
					month(clean_install_date_end) %in% 1:4 &
					year(clean_install_date_end) == year(clean_install_date_start) |
					year(clean_install_date_end) == (year(clean_install_date_start) + 1)
				)
			) ~ 2, # Nov to Apr of next year
			month(clean_install_date) %in% 5:10 |
			(
				month(clean_install_date_start) %in% 5:10 &
				month(clean_install_date_end) %in% 5:10 &
				year(clean_install_date_start) == year(clean_install_date_end)
			) ~ 1 # May to Oct
		),
		clean_upgrade1_quarter = case_when( # upgrade1 quarter
			month(clean_upgrade1_date) %in% c(11:12, 1:4) |
			(
				month(clean_upgrade1_date_start) %in% c(11:12, 1:4) &
				( # Nov to Dec of this year
					month(clean_upgrade1_date_end) %in% 11:12 &
					year(clean_upgrade1_date_end) == year(clean_upgrade1_date_start)
				) |
				( # Jan to Apr of this or next year
					month(clean_upgrade1_date_end) %in% 1:4 &
					year(clean_upgrade1_date_end) == year(clean_upgrade1_date_start) |
					year(clean_upgrade1_date_end) == (year(clean_upgrade1_date_start) + 1)
				)
			) ~ 2, # Nov to Apr of next year
			month(clean_upgrade1_date) %in% 5:10 |
			(
				month(clean_upgrade1_date_start) %in% 5:10 &
				month(clean_upgrade1_date_end) %in% 5:10 &
				year(clean_upgrade1_date_start) == year(clean_upgrade1_date_end)
			) ~ 1 # May to Oct
		),
		clean_upgrade2_quarter = case_when( # upgrade2 quarter
			month(clean_upgrade2_date) %in% c(11:12, 1:4) |
			(
				month(clean_upgrade2_date_start) %in% c(11:12, 1:4) &
				( # Nov to Dec of this year
					month(clean_upgrade2_date_end) %in% 11:12 &
					year(clean_upgrade2_date_end) == year(clean_upgrade2_date_start)
				) |
				( # Jan to Apr of this or next year
					month(clean_upgrade2_date_end) %in% 1:4 &
					year(clean_upgrade2_date_end) == year(clean_upgrade2_date_start) |
					year(clean_upgrade2_date_end) == (year(clean_upgrade2_date_start) + 1)
				)
			) ~ 2, # Nov to Apr of next year
			month(clean_upgrade2_date) %in% 5:10 |
			(
				month(clean_upgrade2_date_start) %in% 5:10 &
				month(clean_upgrade2_date_end) %in% 5:10 &
				year(clean_upgrade2_date_start) == year(clean_upgrade2_date_end)
			) ~ 1 # May to Oct
		)
	)

# Display quarters
toronbike %>%
	as_tibble %>%
	select(-geometry) %>%
	select(
		id,
		clean_install_date,
		clean_install_date_start,
		clean_install_date_end,
		clean_install_quarter,
		clean_upgrade1_date,
		clean_upgrade1_date_start,
		clean_upgrade1_date_end,
		clean_upgrade1_quarter,
		clean_upgrade2_date,
		clean_upgrade2_date_start,
		clean_upgrade2_date_end,
		clean_upgrade2_quarter
	)
```

# Exploration

Explore accuracy of Toronto bikeway data compared.

## Inaccurate Install Years {.tabset}

Inspect all bikeways where the original installation year is not equal to the verified installation year for years 2010 to 2022.

```{r}

# Filter bike for unmatched install year and add diff in years
toronbike_instyearx <- toronbike %>%
	filter(
		install_year != verify_install_year &
		!verify_install_type %in% c("None", NA) &
		verify_install_year > 2009 & verify_install_year <= 2022
	) %>%
	mutate(
		verify_install_year_diff = verify_install_year - install_year,
		verify_install_year_diff_group = case_when(
			verify_install_year_diff <= 1 & verify_install_year_diff >= -1 ~ "±1",
			verify_install_year_diff <= 5 & verify_install_year_diff >= -5 ~ "±5",
			verify_install_year_diff > 5 | verify_install_year_diff < -5 ~ paste0(
				min(verify_install_year_diff),
				" to ",
				"-6 or 6 to ",
				max(verify_install_year_diff)
			)
		),
		verify_install_year_diff_group = factor(
			verify_install_year_diff_group,
			levels = c(
				"±1",
				"±5",
				paste0(
					min(verify_install_year_diff),
					" to ",
					"-6 or 6 to ",
					max(verify_install_year_diff)
				)
			)
		)
	) %>%
	relocate(
		install_year,
		verify_install_year,
		verify_install_year,
		verify_install_year_diff,
		verify_install_year_diff_group,
		install_type,
		verify_install_type,
		.after = street_to
	)

# Calc seg totals
verify_rows <- toronbike %>% nrow
all_rows <- toronbike %>%
	filter(verify_install_year > 2009 & verify_install_year <= 2022) %>%
	nrow
instyearx_rows <- toronbike_instyearx %>% nrow
instyear_rows <- all_rows - instyearx_rows
instyearx1_rows <- toronbike_instyearx %>%
	filter(verify_install_year_diff_group == "±1") %>%
	nrow
instyearx5_rows <- toronbike_instyearx %>%
	filter(verify_install_year_diff_group == "±5") %>%
	nrow
instyearxr_rows <- toronbike_instyearx %>%
	filter(
		verify_install_year_diff_group != "±5" &
		verify_install_year_diff_group != "±1"
	) %>%
	nrow

# Calc seg perc
all_rows_perc <- round(all_rows / verify_rows * 100, 2)
instyearx_perc <- round(instyearx_rows / all_rows * 100, 2)
instyear_perc <- round(instyear_rows / all_rows * 100, 2)
instyearx1_perc <- round(instyearx1_rows / instyearx_rows * 100, 2)
instyearx5_perc <- round(instyearx5_rows / instyearx_rows * 100, 2)
instyearxr_perc <- round(instyearxr_rows / instyearx_rows * 100, 2)

```

* Verified installs: `r verify_rows` (100%)
* Verified installs between 2009 and 2022: `r all_rows` of  (`r all_rows_perc`%) of `r verify_rows` verified installs
* Installs with correct year: `r instyear_rows` (`r instyear_perc`%) of `r all_rows` verified installs between 2009 and 2022
* Installs with incorrect year: `r instyearx_rows` (`r instyearx_perc`%) of `r all_rows` verified installs between 2009 and 2022
* Installs with incorrect year (±1): `r instyearx1_rows` (`r instyearx1_perc`%) of `r instyearx_rows` installs with incorrect year
* Installs with incorrect year (±5): `r instyearx5_rows` (`r instyearx5_perc`%) of `r instyearx_rows` installs with incorrect year
* Installs with incorrect year (>±5): `r instyearxr_rows` (`r instyearxr_perc`%) of `r instyearx_rows` installs with incorrect year

### Map

```{r}
	
# Map bike with unmatched install years
tmap_mode("view")
toronbike_instyearx_map <- tm_basemap("CartoDB.Positron") +
	tm_shape(
		toronbike_instyearx %>%
			select(!ends_with("_comment")) %>%
			st_buffer(25),
		name = "Install with Inaccurate Year"
	) +
	tm_polygons(
		col = "verify_install_year_diff_group",
		title = "Difference in install years",
		border.col = NULL,
		popup.vars = T,
		palette = c("green", "orange", "red")
	)

# Add fullscreen control to map
tmap_leaflet(toronbike_instyearx_map) %>%
	addFullscreenControl()
```

### Data

- [toronto-bikeways-instyearx-2024-06-02.csv](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-instyearx-2024-06-02.csv)
- [toronto-bikeways-instyearx-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-instyearx-2024-06-02.geojson)

```{r}

# Save unmatched install year bike csv
toronbike_instyearx %>%
	mutate(geometry_wkb = st_as_text(geometry)) %>%
	select(-geometry) %>%
    write_sf("../../data/archive/toronto-bikeways-instyearx-2024-06-02.csv", na = "", append = F)

# Save unmatched bike install year geojson
toronbike_instyearx %>%
    write_sf("../../data/archive/toronto-bikeways-instyearx-2024-06-02.geojson", na = "", append = F)

# Display bike with unmatched years
toronbike_instyearx %>%
	as_tibble %>%
	select(-geometry)
```

## No Quarters Assigned {.tabset}

Inspect post-2011 bikeways where segments had no quarters assigned.

```{r}

# Filter bike for post-2011 and no quarter
toronbike_noquarterp2011 <- toronbike %>%
	filter( # post-2011
		verify_install_year > 2011 |
		verify_upgrade1_year > 2011 |
		verify_upgrade2_year > 2011
	) %>%
	filter( # no quarter
		is.na(clean_install_quarter) |
		is.na(clean_upgrade1_quarter) |
		is.na(clean_upgrade2_quarter)
	) %>%
	relocate(
		verify_install_year,
		verify_install_date,
		clean_install_date_start,
		clean_install_date_end,
		clean_install_quarter,
		verify_upgrade1_year,
		verify_upgrade1_date,
		clean_upgrade1_date_start,
		clean_upgrade1_date_end,
		clean_upgrade1_quarter,
		verify_upgrade2_year,
		verify_upgrade2_date,
		clean_upgrade2_date_start,
		clean_upgrade2_date_end,
		clean_upgrade2_quarter,
		.after = street_to
	)

# Assign base cols for map
noquarter_cols <- c(
	"id",
	"street",
	"street_from",
	"street_to"
)

# Filter post-2011 install with no quarters
toronbike_noquartinstp2011 <- toronbike_noquarterp2011 %>%
	select(
		all_of(noquarter_cols),
		starts_with("install"),
		starts_with("verify_install"),
		starts_with("clean_install")
	) %>%
	filter(
		is.na(clean_install_quarter) &
		verify_install_year > 2011 &
		!verify_install_type %in% c("None", NA)
	)

# Filter post-2011 upgrade1 with no quarters
toronbike_noquartu1p2011 <- toronbike_noquarterp2011 %>%
	select(
		all_of(noquarter_cols),
		starts_with("upgrade1"),
		starts_with("verify_upgrade1"),
		starts_with("clean_upgrade1")
	) %>%
	filter(
		is.na(clean_upgrade1_quarter) &
		verify_upgrade1_year > 2011 &
		!verify_upgrade1_type %in% c("None", NA)
	)

# Filter post-2011 upgrade2 with no quarters
toronbike_noquartu2p2011 <- toronbike_noquarterp2011 %>%
	select(
		all_of(noquarter_cols),
		starts_with("upgrade2"),
		starts_with("verify_upgrade2"),
		starts_with("clean_upgrade2")
	) %>%
	filter(
		is.na(clean_upgrade2_quarter) &
		verify_upgrade2_year > 2011 &
		!verify_upgrade2_type %in% c("None", NA)
	)

# Calc seg totals
verify_rows <- toronbike %>% nrow
all_rows <- toronbike %>% filter(
		verify_install_year > 2011 |
		verify_upgrade1_year > 2011 |
		verify_upgrade2_year > 2011
	) %>%
	nrow
noquartinst_rows <- toronbike_noquartinstp2011 %>% nrow
noquartu1_rows <- toronbike_noquartu1p2011 %>% nrow
noquartu2_rows <- toronbike_noquartu2p2011 %>% nrow

# Calc seg perc
all_rows_perc <- round(all_rows / verify_rows * 100, 2)
noquartinst_perc <- round(noquartinst_rows / all_rows * 100, 2)
noquartu1_perc <- round(noquartu1_rows / noquartinst_rows * 100, 2)
noquartu2_perc <- round(noquartu2_rows / noquartinst_rows * 100, 2)

```

* Verified installs: `r verify_rows` (100%)
* Verified installs or upgrades post-2011: `r all_rows` (`r all_rows_perc`%) of `r verify_rows` verified installs
* Installs with no quarters: `r noquartinst_rows` (`r noquartinst_perc`%) of `r all_rows` post-2011 verified installs or upgrades
* 1st upgrades with no quarters: `r noquartu1_rows` (`r noquartu1_perc`%) of `r noquartinst_rows` post-2011 no quarter installs
* 2nd upgrades with no quarters: `r noquartu2_rows` (`r noquartu2_perc`%) of `r noquartinst_rows` post-2011 no quarter installs

### Map

```{r}

# Map bike with unmatched install years
tmap_mode("view")
toronbike_noquarter_map <- tm_basemap("CartoDB.Positron") +
	tm_shape(
		toronbike_noquartinstp2011 %>%
			select(!ends_with("_comment")) %>%
			st_buffer(25),
		name = "Install (Green)"
	) +
	tm_polygons(
		col = "green",
		border.col = "green",
		popup.vars = T
	) +
	tm_shape(
		toronbike_noquartu1p2011 %>%
			select(!ends_with("_comment")) %>%
			st_buffer(25),
		name = "1st Upgrade (Orange)"
	) +
	tm_polygons(
		col = "orange",
		border.col = "orange",
		popup.vars = T
	) +
	tm_shape(
		toronbike_noquartu2p2011 %>%
			select(!ends_with("_comment")) %>%
			st_buffer(25),
		name = "2nd Upgrade (Red)"
	) +
	tm_polygons(
		col = "red",
		border.col = "red",
		popup.vars = T
	)

# Add fullscreen control to map
tmap_leaflet(toronbike_noquarter_map) %>%
	addFullscreenControl()
```

### Data {.tabset .tabset-pills}

#### Install

- [toronto-bikeways-noquartinstp2011-2024-06-02.csv](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartinstp2011-2024-06-02.csv.csv)
- [toronto-bikeways-noquartinstp2011-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartinstp2011-2024-06-02.geojson)

```{r}

# Save post2011 no quarter install csv
toronbike_noquartinstp2011 %>%
	mutate(geometry_wkb = st_as_text(geometry)) %>%
	select(-geometry) %>%
    write_sf("../../data/archive/toronto-bikeways-noquartinstp2011-2024-06-02.csv", na = "", append = F)

# Save post2011 no quarter install geojson
toronbike_noquartinstp2011 %>%
    write_sf("../../data/archive/toronto-bikeways-noquartinstp2011-2024-06-02.geojson", na = "", append = F)

# Display data
toronbike_noquartinstp2011 %>%
	as_tibble %>%
	select(-geometry)

```

#### 1st Upgrade

- [toronto-bikeways-noquartu1p2011-2024-06-02.csv](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartu1p2011-2024-06-02.csv.csv)
- [toronto-bikeways-noquartu1p2011-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartu1p2011-2024-06-02.geojson)

```{r}

# Save post2011 no quarter upgrade1 csv
toronbike_noquartu1p2011 %>%
	mutate(geometry_wkb = st_as_text(geometry)) %>%
	select(-geometry) %>%
    write_sf("../../data/archive/toronto-bikeways-noquartu1p2011-2024-06-02.csv", na = "", append = F)

# Save post2011 no quarter upgrade1 geojson
toronbike_noquartu1p2011 %>%
    write_sf("../../data/archive/toronto-bikeways-noquartu1p2011-2024-06-02.geojson", na = "", append = F)

# Display data
toronbike_noquartu1p2011 %>%
	as_tibble %>%
	select(-geometry)

```

#### 2nd Upgrade

- [toronto-bikeways-noquartu2p2011-2024-06-02.csv](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartu2p2011-2024-06-02.csv.csv)
- [toronto-bikeways-noquartu2p2011-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/archive/toronto-bikeways-noquartu2p2011-2024-06-02.geojson)


```{r}

# Save post2011 no quarter upgrade2 csv
toronbike_noquartu2p2011 %>%
	mutate(geometry_wkb = st_as_text(geometry)) %>%
	select(-geometry) %>%
    write_sf("../../data/archive/toronto-bikeways-noquartu2p2011-2024-06-02.csv", na = "", append = F)

# Save post2011 no quarter upgrade2 geojson
toronbike_noquartu2p2011 %>%
    write_sf("../../data/archive/toronto-bikeways-noquartu2p2011-2024-06-02.geojson", na = "", append = F)

# Display data
toronbike_noquartu2p2011 %>%
	as_tibble %>%
	select(-geometry)

```
