---
title: "Quarterly Collisions in Bloor, University, Danforth, and Yonge"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../../docs/archive/collisions-bloorunidanyonge",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Libraries

Load libraries, the main ones being used are `ggplot2` for plotting, `leaflet` and `tmap` for interactive maps, `sf` for spatial data processing, `tidyverse` for data processing, and `DT` for interactive tables.

```{r}

# Load libraries
library(ggplot2)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(sf)
library(glue)
library(tidyverse)
library(tmap)
library(DT)

```

# Functions

Create a function to assign default behaviour for our `DT` datatables.

```{r}

# Function for default datatable params
datatable <- function(...) {
    
    # Build arg list
    args <- list(...)
    
    # Default custom filename
    filename <- if (!"filename" %in% names(args)) "data" else args$filename
    args[["filename"]] <- NULL
    
    # Add default extensions
    args$extensions <- if (!"extensions" %in% names(args)) "Buttons" else args$extensions
    
    # Add default args
    args$filter <- if (!"filter" %in% names(args)) "top" else args$filter
    args$fillContainer <- if (!"fillContainer" %in% names(args)) T else args$fillContainer
    
    # Add default options
    if (!"options" %in% names(args)) {
        args$options <-  list(
            scrollY = "350px",
            buttons = list(
                list(
                    extend = "csv",
                    filename = filename,
                    exportOptions = list(columns = ":not(.rownames)")
                ),
                list(
                    extend = "excel",
                    filename = filename,
                    exportOptions = list(columns = ":not(.rownames)"),
                    title = ""
                )
            ),
            columnDefs = list(
                list(
                    targets = 0,
                    className = "rownames"
                )
            ),
            dom = "Bfrtip"
        )
    }
    return(do.call(DT::datatable, args))
}

```

# Data

Load the following data for the City of Toronto:

* boundary ([url](https://open.toronto.ca/dataset/regional-municipal-boundary/))
* centrelines ([url](https://open.toronto.ca/dataset/toronto-centreline-tcl/))
* all collisions (private)
* verified bikeways ([url](https://github.com/rrwen/recovr-infracycle/blob/main/data/toronto-bikeways-preprocess-v3.geojson))

Toronto city boundaries data (last updated July 23, 2019) were downloaded November 9, 2024, while Toronto centrelines data (last updated December 6, 2024) were downlaoded December 6, 2024.

All collisions data (2010-2022) was provided by David McElroy <David.McElroy@toronto.ca> from the City of Toronto on November 7, 2024.

Verified bikeways data (2001 to 2022) originally downloaded by Konrad Samsel <konrad.samsel@mail.utoronto.ca> in January 2023 based on Toronto Cycling Network data ([url](https://open.toronto.ca/dataset/cycling-network/)), and last updated by Konrad on March 25, 2024. This data was later modified by Richard Wen <richard.wen@utoronto.ca>, and was last updated October 27, 2024.

```{r}

# Load boundaries
bounds_raw <- read_sf("../../data/toronto-boundary-2019-07-23/citygcs_regional_mun_wgs84.shp")

# Load toronto centrelines data
ctl_raw <- read_sf("../../data/toronto-centrelines-2024-12-06.geojson")

# Load toronto bikeways data
bike_raw <- read_sf("../../data/toronto-bikeways-2024-12-16.geojson")

# Load all toronto collisions from 2022 to 2024
colli_raw <- read_sf(
	"../../tmp/i0327_collisionsrep_acc_export.csv",
	options = c(
		"X_POSSIBLE_NAMES=LONGITUDE",
		"Y_POSSIBLE_NAMES=LATITUDE"
	),
	crs = 4326
)
```

# Cleaning

For the collisions:

1. Crop all collision points to the boundaries of the City of Toronto
2. Convert the accident date `ACCDATE` to `date` type
3. Add a column to separate the collisions data for Killed or Seriously Injured (KSI) (`INJURY` = 3 or 4) and non-KSI individuals (`INJURY` = 1 or 2)
4. Add columns to identify drivers (and passengers, `INVTYPE` = 01 or 02), pedestrians (`INVTYPE` = 03), cyclists (and cyclist passengers, `INVTYPE` = 04 or 05), and other road users (all other `INVTYPE` values).

For the verified bikeways:

1. Filter for Bloor Street, Danforth Street, Danforth Avenue, University Avenue, and Yonge Street
2. Calculate the `final_type` and `final_year` for each verified bikeway segment
	* The `final_type` is the latest infrastructure type after all installations and upgrades
	* If the `verify_upgrade2_type` is the same as the `verify_upgrade1_type` or the `verify_upgrade1_type` is the same as the `verify_install_type`, these are not considered upgrades and either the lower `verify_upgrade1_type` or `verify_install_type` respectively will be considered as the `final_type`
3. Filter for all segments where the `final_type` is cycle track
4. Combine both painted lane `PL` and buffered lanes `BUF` into one category of `painted lane`

For the centrelines:

1. Extract target streets Bloor Street, Danforth Street/Avenue, University Avenue, and Yonge Street
2. Spatially merge all segments for each target street

```{r}

# Reproj city bounds to 4326
bounds <- bounds_raw %>% st_transform(4326)

# Crop colli to toronto bounds and convert date type
colli <- colli_raw %>%
	st_intersection(bounds) %>%
	mutate(colli_date = as_date(ACCDATE))

# Add a columns for ksi and road users
ksi_codes <- c("3", "4")
nonksi_codes <- c("1", "2")
colli <- colli %>%
	mutate(
		ksi = case_when(
			INJURY %in% ksi_codes ~ "ksi",
			INJURY %in% nonksi_codes ~ "non_ksi",
			.default = NA
		), # ksi col
		user = case_when( # road user col
			INVTYPE %in% c("01", "02") ~ "driver",
			INVTYPE == "03" ~ "pedestrian",
			INVTYPE %in% c("04", "05") ~ "cyclist",
			.default = "other"
		)
	) %>%
	filter(!is.na(ksi))

# Get cycle tracks only after all upgrades
bike <- bike_raw %>%
	filter( # verified only
		(!is.na(verify_install_year) & !is.na(verify_install_type)) |
		(!is.na(verify_upgrade1_year) & !is.na(verify_upgrade1_type)) |
		(!is.na(verify_upgrade2_year) & !is.na(verify_upgrade2_type))
	) %>%
	mutate( # create col for target streets
		target_street = case_when(
			str_starts(street, "Bloor") ~
				"Bloor Street",
			str_starts(street, "University") ~
				"University Avenue",
			str_starts(street, "Yonge") ~
				"Yonge Street",
			str_starts(street, "Danforth") ~
				"Danforth Street/Avenue",
			.default = NA
		)
	) %>%
	filter(!is.na(target_street)) %>%
	mutate(across( # reclass to pl or ct
		c(verify_install_type, verify_upgrade1_type, verify_upgrade2_type),
		~ case_when(
			. %in% c("BUF", "PL") ~ "painted lane",
			. == "PBL" ~ "cycle track",
			.default = "none"
		)
	)) %>%
	mutate(
		final_type = case_when( # col for final type without improvements
			verify_upgrade2_type != "none" &
			verify_upgrade2_type != verify_upgrade1_type
				~ verify_upgrade2_type,
			verify_upgrade1_year != "none" &
			verify_upgrade1_type != verify_install_type
				~ verify_upgrade1_type,
			verify_install_year != "none"
				~ verify_install_type,
			.default = NA
		),
		final_year = case_when( # col for final year without improvements
			!is.na(verify_upgrade2_year) &
			verify_upgrade2_type != verify_upgrade1_type
				~ verify_upgrade2_year,
			!is.na(verify_upgrade1_year) &
			verify_upgrade1_type != verify_install_type
				~ verify_upgrade1_year,
			!is.na(verify_install_year)
				~ verify_install_year,
			.default = NA
		),
	) %>%
	filter( # filter for cycle tracks only
		final_type == "cycle track"
	)

# Add history to bike lanes without improvements
bike <- bike %>%
	mutate( # remove improvements
		verify_upgrade1_year = if_else(
			!is.na(verify_upgrade1_type) &
			!is.na(verify_install_type) &
			verify_install_type == verify_upgrade1_type,
			NA,
			verify_upgrade1_year
		),
		verify_upgrade1_type = if_else(
			!is.na(verify_upgrade1_type) &
			!is.na(verify_install_type) &
			verify_install_type == verify_upgrade1_type,
			"none",
			verify_upgrade1_type
		),
		verify_upgrade2_year = if_else(
			!is.na(verify_upgrade2_type) &
			!is.na(verify_upgrade1_type) &
			verify_upgrade2_type == verify_upgrade1_type,
			NA,
			verify_upgrade2_year
		),
		verify_upgrade2_type = if_else(
			!is.na(verify_upgrade2_type) &
			!is.na(verify_upgrade1_type) &
			verify_upgrade2_type == verify_upgrade1_type,
			"none",
			verify_upgrade2_type
		)
	) %>%
	mutate(
        history = glue(
            "{verify_install_type},{verify_install_year}",
            " -> {verify_upgrade1_type},{verify_upgrade1_year}",
            " -> {verify_upgrade2_type},{verify_upgrade2_year}",
        ) %>%
            str_remove_all(" -> none,NA|none,NA -> |none,NA")
    )

# Filter centrelines for Bloor, University and Yonge
streets <- ctl_raw %>%
	filter( # filter for streets only
		FEATURE_CODE_DESC %in% c(
			"Major Arterial",
		    "Major Arterial Ramp",
		    "Minor Arterial",
		    "Minor Arterial Ramp",
		    "Collector",
		    "Access Road",
		    "Other",
		    "Laneway",
		    "Local"
		)
	) %>%
	mutate( # create col for streets bloor, uni, and yonge
		target_street = case_when(
			str_starts(
				LINEAR_NAME_FULL_LEGAL,
				"Bloor Street"
			) ~ "Bloor Street",
			str_starts(
				LINEAR_NAME_FULL_LEGAL,
				"University Avenue"
			) ~ "University Avenue",
			str_starts(
				LINEAR_NAME_FULL_LEGAL,
				"Yonge Street"
			) ~ "Yonge Street",
			str_starts( # Prince Edward Viaduct
				LINEAR_NAME_FULL_LEGAL,
				"Danforth Avenue|Danforth Street"
			) ~ "Danforth Street/Avenue",
			.default = NA
		)
	) %>%
	filter(!is.na(target_street)) %>%
	group_by(target_street) %>% # merge geoms on streets
	summarize(geometry = st_union(geometry))

```

# Processing

Find all collision points within 25 meters of the verified bikeways on Yonge, Bloor, Danforth, and University.

```{r}

# Get row indices of nearest bikeway to each collision
near_idx <- colli %>% st_nearest_feature(bike)

# Calc distances to nearest bikeway for each collision
colli_bike <- colli %>%
	mutate( # calc dist to nearest bikeway for each ksi
		`near_bike_meters` = st_distance(
			geometry,
			bike[near_idx, ],
			by_element = T
		) %>% as.numeric
	)

# Get df version of bike and calc lens
bike_df <- bike %>%
	mutate(
		bike_len_km = as.numeric(st_length(geometry)) / 1000
	) %>%
	as_tibble %>%
	select(-geometry)

# Add bike columns to collisions
colli_bike <- colli_bike %>%
	mutate( # add ids for bike
		bike_id = bike_df[near_idx, ]$id
	) %>%
	left_join( # add bike cols to colli
		bike_df,
		by = join_by(bike_id == id)
	)

# Filter for collisions within 25 meters of the bikeways
colli_bike <- colli_bike %>%
	filter(near_bike_meters <= 25)

```

Calculate bikeway infrastructure types for all years by segment using the installation and upgrade history.

All years before the first installation are considered to have no infrastructure, while all years on or after the first installation is considered the type of the first installation, up until the first upgrade.

All years on or after the first upgrade is considered the type of the first upgrade up until the second upgrade.

Finally, all years on or after the second upgrade is considered the type of the second upgrade onwards.

```{r}

# Convert into bike events
bike_proc <- bike %>%
	pivot_longer(
	    c(
	        verify_install_type,
	        verify_upgrade1_type,
	        verify_upgrade2_type
	    ),
	    names_to = "type_column",
	    values_to = "type"
	) %>%
	filter(!is.na(type)) %>%
	mutate(
		year = case_when(
			type_column == "verify_install_type" ~ verify_install_year,
			type_column == "verify_upgrade1_type" ~ verify_upgrade1_year,
			type_column == "verify_upgrade2_type" ~ verify_upgrade2_year
		)
	) %>%
	rename(
		street_orig = street,
		street = target_street,
		segment_id = id
	) %>%
	select(
		street,
		year,
		segment_id,
		type
	) %>%
	as_tibble %>%
	select(-geometry) %>%
	filter(!is.na(year))

# Fill out the type for all years
min_year <- min(year(colli_bike$colli_date), na.rm = T)
max_year <- max(year(colli_bike$colli_date), na.rm = T)
bike_proc <- bike_proc %>%
    group_by(street) %>%
    group_map(~{ # get all possible years
        expand.grid(
            street = unique(.x$street),
            year = min_year:max_year,
            segment_id = unique(.x$segment_id)
        )
    }, .keep = T) %>%
    bind_rows %>%
    left_join( # join bike events
    	bike_proc,
    	by = c("street", "year", "segment_id")
    ) %>%
    group_by(street, segment_id) %>%
    group_map(~{ # fill out types
        .x %>%
            arrange(year) %>%
            fill(type) %>%
            mutate(type = replace_na(type, "none"))
    }, .keep = T) %>%
    bind_rows %>%
	left_join(
		bike %>%
			select(
				segment_id = id,
				segment_len_km = geometry_len_km
			),
		by = "segment_id"
	) %>%
	select(street, year, segment_id, segment_len_km, type) %>%
	arrange(street, year, segment_id)

# Show bike events
datatable(bike_proc)

```

Count injured individuals for each bikeway segment by year, user, and KSI/Non-KSI.

```{r}

# Count by yearly injured
colli_proc <- colli_bike %>%
	as_tibble %>%
	select(-geometry) %>%
	rename(
		orig_street = street,
		street = target_street,
		segment_id = bike_id
	) %>%
	mutate(
		year = year(colli_date)
	) %>%
	group_by(street, segment_id, year, user, ksi) %>%
	count %>%
	arrange(street, segment_id, year, user, ksi) %>%
	ungroup

# Pivot longer for colli by user and ksi
colli_proc <- colli_proc %>%
	pivot_wider(
		names_from = c(user, ksi),
		values_from = n
	)

# Display colli by seg, user, and ksi
datatable(colli_proc)
```

Add injured individual counts to the bikeway infrastructure type history.

```{r}

# Combine bike types and colli counts
bike_counts <- bike_proc %>%
	left_join(
		colli_proc,
		by = c("street", "year", "segment_id")
	) %>%
	mutate(across(
		ends_with("ksi"),
		~ replace_na(., 0)
	)) %>%
	arrange(street, segment_id, year)

# Display
datatable(bike_counts)
```

# Results

The results contain:

1. Data of injury counts and rates by infrastructure type and year
2. Plot showing injury counts by street and road user

## Data By Segments

The resulting data consists of the following columns:

* `year` the year of the collisions
* `street`: the street of interest
* `type`: the type of infrastructure
* `segment_id`: the segment identifier
* `segment_len_km`: the segment length in kilometers
* `ksi`: the total number of Killed or Seriously Injured (KSI) collisions
* `non_ksi`: the total number of non-KSI collisions
* `<USER>_ksi`: the number of KSI individuals for a road user group indicated by `<USER>`
* `<USER>_non_ksi`: the number of non-KSI individuals for a road user group indicated by `<USER>`

```{r fig.height = 8, fig.width = 8}

# Create output data
out <- bike_counts %>%
	mutate( # calc total ksi
        ksi = select(., !ends_with ("_non_ksi") & ends_with("_ksi")) %>%
            rowSums(na.rm = T),
        non_ksi = select(., ends_with("_non_ksi")) %>%
            rowSums(na.rm = T)
    ) %>%
	select(order(colnames(.))) %>%
	select(
		year,
		street,
		type,
		segment_id,
		segment_len_km,
		ksi,
		non_ksi,
		ends_with("_ksi"),
		everything()
	)

# Show data table
datatable(
	out,
	filename = glue("toronto-collisions-bloorunidanyonge-{today()}")
)
```

## Data By Kilometers

The rates data consists of injured individuals per km-year for each street and infrastructure type:

* `street`: the street of interest
* `type`: the type of infrastructure
* `segments`: the number of segments for the infrastructure type
* `mean_segments_per_year`: the average number of segments per year for the infrastructure type
* `len_km`: the total length of segments for the infrastructure type
* `mean_len_km_per_year`: the average length of segments per year for the infrastructure type
* `ksi`: the total number of Killed or Seriously Injured (KSI) collisions
* `non_ksi`: the total number of non-KSI collisions
* `<USER>_ksi`: the number of KSI individuals for a road user group indicated by `<USER>`
* `<USER>_non_ksi`: the number of non-KSI individuals for a road user group indicated by `<USER>`
* `<USER>_ksi_per_km_year`: the number of KSI individuals for a road user group indicated by `<USER>` per km-year
* `<USER>_non_ksi_per_km_year`: the number of non-KSI individuals for a road user group indicated by `<USER>` per km-year

Injury rates per segment-year for a street and infrastructure type are calculated using the following equation:

$$
Rate_{street,type} = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}x_{i,j}}{l_{i,j}}
$$

* $Rate_{street,type}$ is the number of injuries per km-year for a street and infrastructure type
* $n$ is the number of years in the collision data
* $i$ is the year index of the collision data range (e.g. $i=1$ is the first year)
* $m$ is the number of segments in the collision data
* $j$ is the segment index of the collision data range (e.g. $j=1$ is the first segment)
* $x_{i,j}$ is the number of injuries for year $i$ and segment $j$ for the given street and infrastructure type
* $l_{i,j}$ is the length of segment $i$ for year $j$ for the given street and infrastructure type

```{r}

# Calc rates for output
out_rates <- expand.grid(
	street = unique(out$street),
	type = c("none", "painted lane", "cycle track")
) %>%
	left_join(
		out %>%
			group_by(street, year, type) %>%
			summarize( # yearly calcs
		        segments = n(),
		        len_km = sum(segment_len_km, na.rm = T),
		        across(
		            starts_with("ksi") | ends_with("_ksi"),
		            ~ sum(., na.rm = T)
		        )
		    ) %>%
			group_by(street, type) %>%
		    summarize( # total calcs
		        mean_segments_per_year = mean(segments, na.rm = T),
		        segments = sum(segments, na.rm = T),
		        mean_len_km_per_year = mean(len_km, na.rm = T),
		        len_km = sum(len_km, na.rm = T),
		        across(
		            starts_with("ksi") | ends_with("_ksi"),
		            ~ sum(., na.rm = T)
		        ),
		        across(
		            starts_with("ksi") | ends_with("_ksi"),
		            ~ . / len_km,
		            .names = "{.col}_per_km_year"
		        )
		    ),
		by = c("street", "type")
	) %>%
	select(order(colnames(.))) %>%
	select(
		street,
		type,
		segments,
		mean_segments_per_year,
		len_km,
		mean_len_km_per_year,
		starts_with("ksi"),
		starts_with("non_ksi"),
		everything()
	) %>%
	mutate(
		type = factor(
			type,
			levels = c(
				"none",
				"painted lane",
				"cycle track"
			)
		)
	) %>%
	arrange(street, type)

# Show data table
datatable(
	out_rates,
	filename = glue("toronto-collisions-bloorunidanyonge-rates-{today()}")
)

```

## Plots {.tabset .tabset-dropdown}

These plots show the number of KSI collisions by road user for cycle tracks on Bloor Street, Danforth Street/Avenue, University Avenue, and Yonge Street.

Vertical dotted lines show added cycling infrastructure that was completed for by that year, where CT is Cycle Tracks added and PL is Painted Lanes added.

**Note: Use the dropdown to view different plots by street and road user.**

```{r}

# Prep plot data
plot_data <- colli_bike %>%
	as_tibble %>%
	select(-geometry) %>%
	mutate(
		year = year(colli_date),
		quarter = quarter(colli_date),
		user = str_to_title(user)
	) %>%
	filter(user != "Other") %>%
	rename(
		street_orig = street,
		street = target_street
	) %>%
	group_by(street, year, quarter, user) %>%
	count %>%
	mutate(
		quarter_date = as_date(glue("{year}-{quarter * 3}-01"))
	)

# Calc bike events
bike_events <- bike %>%
	pivot_longer(
	    c(
	        verify_install_type,
	        verify_upgrade1_type,
	        verify_upgrade2_type
	    ),
	    names_to = "type_column",
	    values_to = "type"
	) %>%
	filter(!is.na(type)) %>%
	mutate(
		year = case_when(
			type_column == "verify_install_type" ~ verify_install_year,
			type_column == "verify_upgrade1_type" ~ verify_upgrade1_year,
			type_column == "verify_upgrade2_type" ~ verify_upgrade2_year
		),
		len_km = as.numeric(st_length(geometry)) / 1000
	) %>%
	rename(
		street_orig = street,
		street = target_street
	) %>%
	select(
		street,
		year,
		type,
		len_km
	) %>%
	as_tibble %>%
	select(-geometry) %>%
	filter(!is.na(year)) %>%
	group_by(street, year, type) %>%
	summarize(len_km = sum(len_km, na.rm = T))

# Calc cumulative bike infra lens
min_year <- min(year(colli_bike$colli_date), na.rm = T)
max_year <- max(year(colli_bike$colli_date), na.rm = T)
bike_events <- expand.grid(
	year = min_year:max_year,
	street = unique(bike_events$street),
	type = unique(bike_events$type)
) %>%
	filter(type != "none") %>%
	left_join(bike_events, by = c("street", "year", "type")) %>%
	mutate(clen_km = replace_na(len_km, 0)) %>%
    arrange(type, street, year) %>%
    group_by(street, type) %>%
    mutate(clen_km = cumsum(clen_km)) %>%
	ungroup

# Pivot bike events wider to have types in own cols
bike_events <- bike_events %>%
	mutate(
		event_year = if_else(!is.na(len_km), year, NA),
		event_date = if_else(
			!is.na(event_year),
			as_date(glue("{year}-01-01")),
			NA
		)
	) %>%
    pivot_wider(
        names_from = type,
        names_glue = "{str_replace_all(type, ' ', '_')}_{str_replace(.value, 'event_', '')}",
        values_from = c(event_year, event_date, len_km, clen_km)
    ) %>%
	mutate(
		type_label = case_when(
			!is.na(painted_lane_len_km) & !is.na(cycle_track_len_km)
				~ glue(
					"PL: +{round(painted_lane_len_km, 1)} km\n",
					"CT: +{round(cycle_track_len_km, 1)} km"
				),
			!is.na(painted_lane_len_km)
				~ glue("PL: +{round(painted_lane_len_km, 1)} km"),
			!is.na(cycle_track_len_km)
				~ glue("CT: +{round(cycle_track_len_km, 1)} km"),
			.default = NA
		),
		type_year = if_else(
			!is.na(painted_lane_len_km) | !is.na(cycle_track_len_km),
			year,
			NA
		),
		type_date = if_else(
			!is.na(painted_lane_len_km) | !is.na(cycle_track_len_km),
			as_date(glue("{year}-01-01")),
			NA
		)
	)

# Get bike lens
bike_len <- bike %>%
	rename(
		street_orig = street,
		street = target_street
	) %>%
	group_by(street) %>%
	summarize(
		street_len_km = sum(as.numeric(st_length(geometry)), na.rm = T) / 1000
	) %>%
	mutate(
		street_label = glue("{street} ({round(street_len_km, 1)} km)")
	)

# Add bike events and lens to plot data
plot_data <- plot_data %>%
	left_join(
		bike_events,
		by = c("street", "year")
	) %>%
	left_join(
		bike_len,
		by = "street"
	)

```

```{r}

# Function to generate plot
plot_colli <- function(
		df = plot_data,
		y_title = "Injuries",
		title = NULL
) {
	
	# Generate vline label
	has_ct <- !(df %>% pull(cycle_track_year) %>% is.na(.) %>% all)
	has_pl <- !(df %>% pull(painted_lane_year) %>% is.na(.) %>% all)
	vlabel <- case_when(
		has_ct & has_pl ~ "Cycle Track (CT) / Painted Lane (PL)",
		has_ct ~ "Cycle Track (CT)",
		has_pl ~ "Painted Lane (PL)"
	)
	
	# Generate plot
	ggplot(
		df %>% mutate(vline_label = glue("{vlabel} Added")),
		aes(x = quarter_date, y = n)
	) +
		geom_vline(
			aes(xintercept = type_date, color = vline_label),
			size = 0.5,
			linetype = "dashed",
			alpha = 0.8
		) +
		geom_label(
			aes(
				x = type_date,
				label = type_label
			),
			color = "gray30",
			angle = 90,
			y = max(df$n * 1.4, na.rm = T)
		) +
		geom_line(
			color = "gray70",
			alpha = 0.8
		) +
		geom_text(
			aes(label = n),
			color = "gray10",
			size = 3,
			angle = 90,
			hjust = -1
		) +
		theme_minimal() +
		scale_color_manual(values = "gray30") +
		scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
		scale_y_continuous(expand = expansion(mult = c(0.1, 0.75))) +
		labs(
			x = "Year",
			y = y_title,
			title = title
		) +
		theme(
			plot.title = element_text(hjust = 0.5),
			axis.text.x = element_text(size = 10),
			legend.position = "bottom",
			legend.margin = margin(t = 0),
			legend.title = element_blank()
		)
}

```

```{r toronto-collisions-bloorunidanyonge-plots, results = "asis", fig.width = 8, fig.height = 4.5}

# Generate street and user combo for plots
plot_combos <- expand.grid(
	street = unique(plot_data$street),
	user = unique(plot_data$user)
)

# Create folder to store plots
dir.create("../../figures/collisions-bloorunidanyonge", recursive = TRUE)

# Plot by street and user
for (i in 1:nrow(plot_combos)) {
	
	# Get street and user combo
	streeti <- as.character(plot_combos$street[i])
	useri <- as.character(plot_combos$user[i])
	
	# Prep plot data for street and user
	df <- plot_data %>%
		filter(user == useri & street == streeti)
	
	# Generate the plot for street and user
	useri_label <- case_when(
	    useri == "Pedestrian" ~ "Pedestrians",
	    useri == "Driver" ~ "Car Occupants",
	    useri == "Cyclist" ~ "Bicycle Occupants"
	)
	p <- plot_colli(
		df,
		title = df$street_label %>% unique,
		y_title = glue("{useri_label} (n={sum(df$n, na.rm = T)})")
	)
	
	# Save the plot
	streeti_file <- str_split_i(as.character(streeti), " ", 1) %>%
		str_to_lower
	useri_file <- str_to_lower(useri)
	ggsave(
		glue(
			"../../figures/collisions-bloorunidanyonge/",
			"toronto-collisions-{streeti_file}-{useri_file}.png"
		),
		plot = p,
		width = 8,
		height = 4.5,
		dpi = 300
	)
	
	# Create section and display plot
	cat("\n\n\n\n### ", streeti, " (", useri, ")\n\n\n\n", sep = "")
	cat("\n\n\n\n")
	print(p)
	cat("\n\n\n\n")
	
}

```



```{r, echo = F, eval = F}

## Maps {.tabset .tabset-dropdown}

These map show the location of cycle tracks on Bloor Street, Danforth Street/Avenue, University Avenue, and Yonge Street, along with the collision density (determined by the number of individuals involved) within 25 meters of each cycle track.

Pre means before the noted year, and post means on or after the noted year (e.g. Pre-2022 is before 2022 and Post-2022 is on or after 2022).

# Function to produce a heatmap
plot_map <- function(
		user_type = NULL,
		ksi_type = NULL,
		pre = NULL,
		post = NULL,
		radius = 8,
		heatmap = T
	) {
	
	# Filter colli data
	plot_colli <- colli_bike %>%
		mutate(
			year = year(colli_date)
		)
	if (!is.null(user_type)) {
		plot_colli <- plot_colli %>% filter(user %in% user_type)
	}
	if (!is.null(ksi_type)) {
		plot_colli <- plot_colli %>% filter(ksi %in% ksi_type)
	}
	if(!is.null(pre)) {
		plot_colli <- plot_colli %>% filter(year < pre)
	}
	if(!is.null(post)) {
		plot_colli <- plot_colli %>% filter(year >= post)
	}
	
	# Plot map
	tmap_mode("view")
	p <- tm_shape(bounds) +
		tm_polygons(alpha = 0, popup.vars = F, id = "") +
		tm_shape(streets) +
		tm_lines(col = "#1d1d1d", popup.vars = T) +
		tm_text("target_street", size = 1) +
		tm_shape(
			bike %>%
				select(-ends_with("_comment")) %>%
				filter(target_street != "Bloor & Danforth")
		) +
		tm_lines(col = "red", lwd = 2, popup.vars = T)
	
	# If no heatmap, use points
	if (!heatmap) {
		p <- p +
			tm_shape(plot_colli) +
			tm_dots(clustering = T)
	}
	
	# Convert to leaflet
	p <- tmap_leaflet(p)
	
	# Add controls and legends
	p <- p %>%
		addFullscreenControl %>%
		addLegend(
			position = "topright",
			colors = c("red", "#1d1d1d"),
			labels = c("Cycle Track", "Street")
		) %>%
		addLegend(
			position = "topright",
			colors = colorNumeric(
				c("blue", "green", "yellow", "orange"),
				domain = NULL
			)(seq(1, 100, length.out = 5)),
			values = c(1, 10),
			labels = c("Low", "", "", "", "High"),
			title = "Collision<br/>Density"
		)
	
	# Add heatmap
	if (heatmap) {
		p <- p %>% addHeatmap(
			data = plot_colli,
			lat = ~LATITUDE,
			lng = ~LONGITUDE,
			radius = radius
		)
	} else {
		p <- p %>%
			htmlwidgets::onRender("
			    function(el, x) {
					var css = '.marker-cluster span { color: black; font-size: 12px; }';
					var style = document.createElement('style');
					style.type = 'text/css';
					if (style.styleSheet) {
						style.styleSheet.cssText = css;
					} else {
						style.appendChild(document.createTextNode(css));
					}
					document.head.appendChild(style);
			    }
			")
	}
	return(p)
}

### All Non-KSI

plot_map(ksi_type = "non_ksi")

# Graphviz
digraph G {
  
  subgraph cluster1{
    label = "Step 1: Data Cleaning"
    
    bike1[label="Bloor, University,\nDanforth, and Yonge\nCycle Tracks\n(n=16)" shape=rect]
    colli1[label="Minimal, Minor\nMajor, and Fatal Injuries\nInside Toronto\n(n=170,462)" shape=rect]
  }
  
  subgraph cluster2{
      label = "Step 2: Spatial Data Processing"
      colli2[label="Injuries within\n25 Meters of Cycle Track\n(n=5,504)" shape=rect]
  }
  
  subgraph cluster3{
      label = "Step 3: Injury Rate Calculations"
      colli3d[label="Driver-related\nCollisions\n(n=3,257)" shape=rect]
      colli3p[label="Pedestrian-related\nCollisions\n(n=879)" shape=rect]
      colli3c[label="Cyclist-related\nInjuries\n(n=1,049)" shape=rect]
      colli3o[label="Other\nInjuries\n(n=319)" shape=rect]
  }
  
  bike[label="Verified Bikeways\n(n=1,323)" shape=cylinder]
  colli[label="All Collisions\n(n=1,355,219)" shape=cylinder]
  bounds[label="Toronto City Bounds\n(n=1)" shape=cylinder]
  
  bike -> bike1 -> colli2:w
  colli,bounds -> colli1 -> colli2:e
  colli2 -> colli3d:w
  colli2 -> colli3p
  colli2 -> colli3c:e
  colli2 -> colli3o
}
```
