---
title: "Quarterly Collisions in Bloor, University and Yonge"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../../docs/archive/collisions-blooruniyonge",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Libraries

Load libraries, the main ones being used are `ggplot2` for plotting, `leaflet` and `tmap` for interactive maps, `sf` for spatial data processing, `tidyverse` for data processing, and `DT` for interactive tables.

```{r}

# Load libraries
library(ggplot2)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(sf)
library(glue)
library(tidyverse)
library(tmap)
library(DT)

```

# Functions

Create a function to assign default behaviour for our `DT` datatables.

```{r}

# Function for default datatable params
datatable <- function(...) {
    
    # Build arg list
    args <- list(...)
    
    # Default custom filename
    filename <- if (!"filename" %in% names(args)) "data" else args$filename
    args[["filename"]] <- NULL
    
    # Add default extensions
    args$extensions <- if (!"extensions" %in% names(args)) "Buttons" else args$extensions
    
    # Add default args
    args$filter <- if (!"filter" %in% names(args)) "top" else args$filter
    args$fillContainer <- if (!"fillContainer" %in% names(args)) T else args$fillContainer
    
    # Add default options
    if (!"options" %in% names(args)) {
        args$options <-  list(
            scrollY = "350px",
            buttons = list(
                list(
                    extend = "csv",
                    filename = filename,
                    exportOptions = list(columns = ":not(.rownames)")
                ),
                list(
                    extend = "excel",
                    filename = filename,
                    exportOptions = list(columns = ":not(.rownames)"),
                    title = ""
                )
            ),
            columnDefs = list(
                list(
                    targets = 0,
                    className = "rownames"
                )
            ),
            dom = "Bfrtip"
        )
    }
    return(do.call(DT::datatable, args))
}

```

# Data

Load the following data for the City of Toronto:

* boundary
* centrelines
* all collisions
* verified bikeways

All collisions data was provided by David McElroy <David.McElroy@toronto.ca> from the City of Toronto on November 7, 2024.

Verified bikeways data originally from Konrad Samsel <konrad.samsel@mail.utoronto.ca> and later modified by Richard Wen <richard.wen@utoronto.ca> on October 27, 2024.

```{r}

# Load boundaries
bounds_raw <- read_sf("../../data/toronto-boundary-2019-07-23/citygcs_regional_mun_wgs84.shp")

# Load toronto centrelines data
ctl_raw <- read_sf("../../data/toronto-centrelines-2024-12-06.geojson")

# Load toronto bikeways data
bike_raw <- read_sf("../../data/toronto-bikeways-2024-10-27.geojson")

# Load all toronto collisions from 2022 to 2024
colli_raw <- read_sf(
	"../../tmp/i0327_collisionsrep_acc_export.csv",
	options = c(
		"X_POSSIBLE_NAMES=LONGITUDE",
		"Y_POSSIBLE_NAMES=LATITUDE"
	),
	crs = 4326
)
```

# Cleaning

For the collisions:

1. Crop all collision points to the boundaries of the City of Toronto
2. Convert the accident date `ACCDATE` to `date` type
3. Add a column to separate the collisions data for Killed or Seriously Injured (KSI) and non-KSI individuals
4. Add columns to identify drivers (and passengers), pedestrians, cyclists (and cyclist passengers), and other road users

For the verified bikeways:

1. Filter for Bloor Street, University Avenue, and Yonge Street
2. Calculate the final type and year for each verified bikeway, where improvements (when the same type occurs between upgrades) are not considered an upgrade
3. Filter for cycle tracks only based on the final type

For the centrelines, extract target streets Bloor Street, University Avenue, and Yonge Street, and merge all segments for each target street.

```{r}

# Reproj city bounds to 4326
bounds <- bounds_raw %>% st_transform(4326)

# Crop colli to toronto bounds and convert date type
colli <- colli_raw %>%
	st_intersection(bounds) %>%
	mutate(colli_date = as_date(ACCDATE))

# Add a columns for ksi and road users
ksi_codes <- c("3", "4")
colli <- colli %>%
	mutate(
		ksi = if_else(INJURY %in% ksi_codes, "ksi", "non_ksi"), # ksi col
		user = case_when( # road user col
			INVTYPE %in% c("01", "02") ~ "driver",
			INVTYPE == "03" ~ "pedestrian",
			INVTYPE %in% c("04", "05") ~ "cyclist",
			.default = "other"
		)
	)

# Get cycle tracks only after all upgrades for yonge, bloor, uni
bike <- bike_raw %>%
	mutate( # create col for target streets bloor, uni, and yonge
		target_street = case_when(
			str_starts(street, "Bloor") ~
				"Bloor Street",
			str_starts(street, "University") ~
				"University Avenue",
			str_starts(street, "Yonge") ~
				"Yonge Street",
			.default = NA
		)
	) %>%
	filter(!is.na(target_street)) %>%
	mutate(
		final_type = case_when( # col for final type without improvements
			!is.na(verify_upgrade2_year) &
			verify_upgrade2_type != verify_upgrade1_type
				~ verify_upgrade2_type,
			!is.na(verify_upgrade1_year) &
			verify_upgrade1_type != verify_install_type
				~ verify_upgrade1_type,
			!is.na(verify_install_year)
				~ verify_install_type,
			.default = NA
		),
		final_type = case_when( # remap infra types to actual names
			final_type %in% c("PL", "BUF") ~ "Painted Lane",
			final_type == "PBL" ~ "Cycle Track",
			.default = NA
		)
	) %>%
	mutate( # col for final year without improvements
		final_year = case_when(
			!is.na(verify_upgrade2_year) &
			verify_upgrade2_type != verify_upgrade1_type
				~ verify_upgrade2_year,
			!is.na(verify_upgrade1_year) &
			verify_upgrade1_type != verify_install_type
				~ verify_upgrade1_year,
			!is.na(verify_install_year)
				~ verify_install_year,
			.default = NA
		),
	) %>%
	filter( # filter for cycle tracks only
		final_type == "Cycle Track"
	)

# Filter centrelines for Bloor, University and Yonge
streets <- ctl_raw %>%
	filter( # filter for streets only
		FEATURE_CODE_DESC %in% c(
			"Major Arterial",
		    "Major Arterial Ramp",
		    "Minor Arterial",
		    "Minor Arterial Ramp",
		    "Collector",
		    "Access Road",
		    "Other",
		    "Laneway",
		    "Local"
		)
	) %>%
	mutate( # create col for target streets bloor, uni, and yonge
		target_street = case_when(
			str_starts(LINEAR_NAME_FULL_LEGAL, "Bloor Street") ~
				"Bloor Street",
			str_starts(LINEAR_NAME_FULL_LEGAL, "University Avenue") ~
				"University Avenue",
			str_starts(LINEAR_NAME_FULL_LEGAL, "Yonge Street") ~
				"Yonge Street",
			.default = NA
		)
	) %>%
	filter(!is.na(target_street)) %>%
	group_by(target_street) %>% # merge geoms based on target streets
	summarize(geometry = st_union(geometry))

```

# Processing

Find all collision points within 25 meters of the verified bikeways on Yonge, Bloor and University, then count the number of individuals involved by the collision year, KSI/Non-KSI, and road user.

```{r}

# Get row indices of nearest bikeway to each collision
near_idx <- colli %>% st_nearest_feature(bike)

# Calc distances to nearest bikeway for each collision
colli_bike <- colli %>%
	mutate( # calc dist to nearest bikeway for each ksi
		`near_bike_meters` = st_distance(
			geometry,
			bike[near_idx, ],
			by_element = T
		) %>% as.numeric
	)

# Get df version of bike
bike_df <- bike %>% as_tibble %>% select(-geometry)

# Add bike columns to collisions
colli_bike <- colli_bike %>%
	mutate( # add ids for bike
		bike_id = bike_df[near_idx, ]$id
	) %>%
	left_join( # add bike cols to colli
		bike_df,
		by = join_by(bike_id == id)
	)

# Filter for collisions within 25 meters of the bikeways
colli_bike <- colli_bike %>%
	filter(near_bike_meters <= 25)

# Count them by collision year
colli_counts <- colli_bike %>%
	as_tibble %>%
	select(-geometry) %>%
	mutate(year = year(colli_date)) %>%
	group_by(year, target_street, user, ksi) %>%
	count %>%
	rename(
		street = target_street
	) %>%
	arrange(year, street, user, ksi) %>%
	ungroup

```

# Results {.tabset}

The results contain:

1. Data in wide format with counts divided by road user and KSI/Non-KSI for each street by year
2. A plot showing when the data in 1. with when the cycle tracks were implemented for each street
3. A map showing the locations of the cycle tracks for each street and a heatmap of the collision density determined by individuals involved

## Data

The resulting data consists of the following columns:

* `year`: the year of the collisions
* `street`: the street, one of Yonge Street, University Avenue, or Bloor Street
* `ksi`: the number of Killed or Seriously Injured (KSI) individuals
* `non_ksi`: the number of non-KSI individuals
* `<USER>_ksi`: the number of KSI individuals for a road user group indicated by `<USER>`
* `<USER>_non_ksi`: the number of non-KSI individuals for a road user group indicated by `<USER>`

```{r fig.height = 8, fig.width = 8}

# Create output data
out <- colli_counts %>%
	mutate(
		group = glue("{user}_{ksi}")
	) %>%
	pivot_wider(
		values_from = n,
		names_from = group
	) %>%
	select(-user, -ksi) %>%
	group_by(year, street) %>%
	summarize(across(everything(), ~ sum(., na.rm = T))) %>%
	ungroup %>%
	mutate(
		ksi = select(., !ends_with ("_non_ksi") & ends_with("_ksi")) %>%
			rowSums(na.rm = T),
		non_ksi = select(., ends_with("_non_ksi")) %>%
			rowSums(na.rm = T)
	) %>%
	select(order(colnames(.))) %>%
	select(year, street, ksi, non_ksi, everything())

# Show data table
datatable(
	out,
	filename = glue("toronto-collisions-blooruniyonge-{today()}")
)
```

## Plot

This plot shows the number of KSI individuals by road user for cycle tracks on Bloor Streeet, University Avenue, and Yonge Street. The dark gray dotted vertical lines show which year a cycle track is added to the street along with the amount of cycle track kilometers added.

```{r fig.width = 18, fig.height = 9}

# Calc cycle track final year events by road segment length
bike_events <- bike %>%
	group_by(target_street, final_year) %>%
	summarize(
		len_km = sum(st_length(geometry)) / 1000 %>%
			as.numeric
	) %>%
	mutate(
		len_km_label = glue("+{round(len_km, 2)} km"),
		len_km_legend = "Cycle Track Added"
	) %>%
	rename(year = final_year, street = target_street)

# Calc label position for ct events
bike_events_y <- colli_counts %>%
	group_by(user) %>%
	summarize(
		len_km_label_y = max(n, na.rm = T) + 0.3 * (max(n, na.rm = T))
	)

# Prep plot data
plot_data <- colli_counts %>% 
	filter(!user %in% c("other")) %>%
	left_join(bike_events, by = c("year", "street")) %>%
	left_join(bike_events_y, by = "user") %>%
	mutate(
		user = str_to_title(user),
		ksi = if_else(ksi == "non_ksi", "Non-KSI", "KSI")
	)

# Generate plot
p <- ggplot(
	plot_data,
	aes(x = year, y = n, fill = ksi, color = ksi)
) +
	geom_line() +
	geom_point(show.legend = F) +
	geom_vline(
		data = bike_events,
		aes(xintercept = year, color = len_km_legend),
		linetype = "dashed"
	) +
	geom_text(
		aes(label = n),
		size = 3,
		vjust = 2
	) +
	facet_grid(
		user ~ street,
		scales = "free",
		switch = "y"
	) +
	geom_label(
		aes(
			x = year,
			y = len_km_label_y,
			label = len_km_label
		),
		color = "darkgray",
		fill = "white",
		size = 3.5,
		angle = 90
	) +
	theme_minimal() +
	scale_color_manual(values = c(
		"Non-KSI" = "#00BFC4",
		"KSI" = "#F8766D",
		"Cycle Track Added" = "darkgray"
	)) +
	scale_x_continuous(
		breaks = seq(
			min(colli_counts$year, na.rm = T),
			max(colli_counts$year, na.rm = T),
			by = 1
		)
	) +
	scale_y_continuous(expand = expansion(mult = c(0.1, 0.3))) +
	labs(
		x = "Year",
		y = "Individuals Involved"
	) +
	theme(
		axis.text.x = element_text(size = 8),
		legend.title = element_blank(),
		legend.position = "top"
	)
p
```

## Map

This map shows the location of cycle tracks on Bloor Streeet, University Avenue, and Yonge Street, along with the collision density determined by the number of individuals within 25 meters of each cycle track.

```{r}

# Visualize results of target streets and cycle tracks
tmap_mode("view")
tmap_leaflet(
	tm_shape(bounds) +
		tm_polygons(alpha = 0, ) +
		tm_shape(streets) +
		tm_lines(col = "#1d1d1d", popup.vars = T) +
		tm_text("target_street", size = 1) +
		tm_shape(bike) +
		tm_lines(col = "red", lwd = 2, popup.vars = T)
	) %>%
	addFullscreenControl%>%
	addHeatmap(
		lat = colli_bike$LATITUDE,
		lng = colli_bike$LONGITUDE,
		radius = 8
	) %>%
	addLegend(
		position = "topright",
		colors = c("red", "#1d1d1d"),
		labels = c("Cycle Track", "Street")
	) %>%
	addLegend(
		position = "topright",
		colors = colorNumeric(
			c("blue", "green", "yellow", "orange"),
			domain = NULL
		)(seq(1, 100, length.out = 5)),
		values = c(1, 10),
		labels = c("Low", "", "", "", "High"),
		title = "Collision<br/>Density"
	)

```
