---
title: "Cycling Interventions Evaluation"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
- "Brice Batomen brice.kuimi@utoronto.ca"
- "Linda Rothman linda.rothman@torontomu.ca"
- "Andrew Howard andrew.howard@sickkids.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../docs",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Installation

1.  Install [R](https://www.r-project.org/)
2.  Install [RTools](https://cran.r-project.org/) if you are on Windows
3.  Install [RStudio](https://posit.co/download/rstudio-desktop/)

For more details, see [Software and Package Versions](#software-and-package-versions).

# Running This Code

1.  Ensure the installation steps above are completed
2.  Download a zip of the code and data [here](https://github.com/rrwen/recovr-eval/archive/refs/heads/main.zip) and unzip it
    - Code Repository: [github.com/rrwen/recovr-eval](https://github.com/rrwen/recovr-eval)
3.  In RStudio, open the [src/src.Rproj](https://github.com/rrwen/recovr-eval/blob/main/src/src.Rproj) file
4.  Then, open the [src/index.Rmd](https://github.com/rrwen/recovr-eval/blob/main/src/index.Rmd) file
5.  In RStudio:
    - Run all code: Click the `Run` drop down (top right of the code pane) and click `Run All`
    - Generate HTML version: Click `knit` (top left of code pane) and a file will be generated in `docs/index.html`

# Libraries

Install R packages if needed.

```{r, results = FALSE}

# Required packages
required_packages <- c(
	"rmarkdown",
	"bookdown",
	"knitr",
	"tidyverse",
	"purrr",
	"glue",
	"lubridate",
	"scales",
	"patchwork",
	"DiagrammeR",
	"DiagrammeRsvg",
	"DT",
	"webshot2",
	"magick",
	"rsvg",
	"sf",
	"tmap",
	"ggspatial",
	"prettymapr",
	"units",
	"leaflet",
	"leaflet.extras",
	"igraph"
)

# Try to install packages if not installed
default_options <- options()
tryCatch(
	{
		# Disable interactivity
		options(install.packages.compile.from.source = "always")
		
		# Install package if not installed
		for (package in required_packages) {
			is_package_installed <- require(package, character.only = TRUE)
			if (!is_package_installed) {
				cat(paste0("Installing package: ", package, "\n"))
				install.packages(package)
			} else {
				cat(paste0("Package already installed: ", package, "\n"))
			}
		}
	},
	error = function(cond) {
		stop(cond)
	},
	finally = {
		options(default_options) # reset interactivity
	}
)
```

Load R libraries.

```{r}
library(DiagrammeR)
library(ggplot2)
library(ggspatial)
library(glue)
library(igraph)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(patchwork)
library(sf)
library(tidyverse)
library(tmap)
```

# Data

Read data from the `data` folder.

```{r}
ddesc <- read_csv("../data/data.csv")
ddesc
```

## Vancouver Bikeways {.tabset}

`r ddesc %>% filter(file == "vancouver-bikeways-2024-10-01.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "vancouver-bikeways-2024-10-01.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "vancouver-bikeways-2024-10-01.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "vancouver-bikeways-2024-10-01.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "vancouver-bikeways-2024-10-01.geojson") %>% pull(notes)`

```{r}

# Read data
vancbike_raw <- read_sf("../data/vancouver-bikeways-2024-10-01.geojson")

# Get download date
vancbike_dldate <- ddesc %>% filter(
	file == "vancouver-bikeways-2024-10-01.geojson"
) %>% pull(download_date)
```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(vancbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(vancbike_raw)`
* Rows: `r nrow(vancbike_raw)`

```{r}
vancbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
vancbike_ddict <- read_csv("../data/vancouver-bikeways-2024-10-01-datadict.csv")
vancbike_ddict
```

### Details

```{r}
print(vancbike_raw)
```

### Files

The data files are available below:

- [vancouver-bikeways-2024-10-01.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/vancouver-bikeways-2024-10-01.geojson)
- [vancouver-bikeways-2024-10-01-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/vancouver-bikeways-2024-10-01-datadict.csv)

## Calgary Bikeways {.tabset}

`r ddesc %>% filter(file == "calgary-bikeways-2024-10-01.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "calgary-bikeways-2024-10-01.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "calgary-bikeways-2024-10-01.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "calgary-bikeways-2024-10-01.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "calgary-bikeways-2024-10-01.geojson") %>% pull(notes)`

```{r}

# Read data
calgbike_raw <- read_sf("../data/calgary-bikeways-2024-10-01.geojson")

# Get download date
calgbike_dldate <- ddesc %>% filter(
	file == "calgary-bikeways-2024-10-01.geojson"
) %>% pull(download_date)
```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(calgbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(calgbike_raw)`
* Rows: `r nrow(calgbike_raw)`

```{r}
calgbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
calgbike_ddict <- read_csv("../data/calgary-bikeways-2024-10-01-datadict.csv")
calgbike_ddict
```

### Details

```{r}
print(calgbike_raw)
```

### Files

The data files are available below:

- [calgary-bikeways-2024-10-01.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/calgary-bikeways-2024-10-01.geojson)
- [calgary-bikeways-2024-10-01-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/calgary-bikeways-2024-10-01-datadict.csv)

## Toronto Bikeways {.tabset}

`r ddesc %>% filter(file == "toronto-bikeways-2024-10-01.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "toronto-bikeways-2024-10-01.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-10-01.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-10-01.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "toronto-bikeways-2024-10-01.geojson") %>% pull(notes)`

```{r}

# Read data
toronbike_raw <- read_sf("../data/toronto-bikeways-2024-10-01.geojson")

# Get download date
toronbike_dldate <- ddesc %>% filter(
	file == "toronto-bikeways-2024-10-01.geojson"
) %>% pull(download_date)

```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(toronbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(toronbike_raw)`
* Rows: `r nrow(toronbike_raw)`

```{r}
toronbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
toronbike_ddict <- read_csv("../data/toronto-bikeways-2024-10-01-datadict.csv")
toronbike_ddict
```

### Details

```{r}
print(toronbike_raw)
```

### Files

The data files are available below:

- [toronto-bikeways-2024-10-01.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-10-01.geojson)
- [toronto-bikeways-2024-10-01-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-10-01-datadict.csv)

## Verified Dates {.tabset}

`r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(notes)`

```{r}

# Read data
vdates_raw <- read_csv("../data/verify-dates-2024-06-12.csv")

# Get download date
vdates_dldate <- ddesc %>% filter(
	file == "verify-dates-2024-06-12.csv"
) %>% pull(download_date)
```

### Data

* Columns: `r ncol(vdates_raw)`
* Rows: `r nrow(vdates_raw)`

```{r}
vdates_raw
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
vdates_ddict <- read_csv("../data/verify-dates-2024-06-12-datadict.csv")
vdates_ddict
```

### Files

The data files are available below:

- [verify-dates-2024-06-12.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12.csv)
- [verify-dates-2024-06-12-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12-datadict.csv)

## Toronto KSI {.tabset}

`r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(notes)`

```{r}

# Read data
toronksi_raw <- read_sf(
	"../data/toronto-ksi-2024-06-13.csv",
	options = c(
		"X_POSSIBLE_NAMES=LONGITUDE",
		"Y_POSSIBLE_NAMES=LATITUDE"
	),
	crs = 4326
)

# Get download date
toronksi_dldate <- ddesc %>% filter(
	file == "toronto-ksi-2024-06-13.csv"
) %>% pull(download_date)
```

### Map

**Note**: Due to the large number of records, only the latest year of `r year(max(toronksi_raw$DATE))` is displayed (n = `r toronksi_raw %>% filter(year(DATE) == max(year(DATE))) %>% nrow`).

```{r}
tmap_mode("view")
tm_shape(toronksi_raw %>% filter(year(DATE) == max(year(DATE)))) +
    tm_dots(
    	col = "ACCLASS",
    	clustering = TRUE,
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(toronksi_raw)`
* Rows: `r nrow(toronksi_raw)`

```{r}
toronksi_raw %>% as_tibble()
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
ksi_ddict <- read_csv("../data/toronto-ksi-2024-06-13-datadict.csv")
ksi_ddict
```

### Details

```{r}
print(toronksi_raw)
```

### Files

The data files are available below:

- [toronto-ksi-2024-06-13.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-ksi-2024-06-13.csv)
- [toronto-ksi-2024-06-13-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-ksi-2024-06-13-datadict.csv)

## Vancouver KSI

```{r}
vancksi_raw <- read.csv("../tmp/TAS persons involved in crashes in Vancouver 2010-2021.csv")
```

## Calgary KSI

```{r}
calgksi_raw <- read.csv("../tmp/Combined Collision Data for Calgary (2016 - 2021).csv") %>%
	filter(CASE_YEAR < 2021) %>%
	mutate(
		OCCURENCE_TIMESTAMP = as.Date(OCCURENCE_TIMESTAMP, origin = "1899-12-30") # excel date serial numbers
	) %>%
	add_row(
		read.csv("../tmp/Collision Data for Calgary (2021 - 2022).csv") %>%
			mutate(OCCURENCE_TIMESTAMP = OCCURENCE_TIMESTAMP %>% as_datetime)
	)
```

# Cleaning

## Combine Bikeways {.tabset}

Combine bikeway data across all cities using common columns.

**Note**: For Calgary and Toronto, there weren't any included bikeways that were not verified (e.g. Local Street Bikeways (LSBs)), and a dummy column was added to account for this case.

**Note 2**: Bikeway lengths were also calculated in meters as an additional column `geometry_len_meters`. 

```{r}

# List of city bikeway data
bike_list <- list(
	vancouver = vancbike_raw %>%
		mutate(
			build_status = "EXISTING"
		),
	calgary = calgbike_raw %>%
		mutate(
			no_verify_install_type = NA,
			build_status = str_to_upper(status)
		),
	toronto = toronbike_raw %>%
		mutate(
			no_verify_install_type = NA,
			build_status = "EXISTING"
		)
)

# Get common columns across all city bikeways
bike_cols <- bike_list %>%
	map(colnames) %>%
	reduce(intersect)

# Combine bikeway data across cities
bike_combine <- names(bike_list) %>%
	map(function(city) {
		bike_list[[city]] %>%
			select(
				all_of(bike_cols)
			) %>%
			mutate(
				city = factor(city, levels = names(bike_list)),
				.before = 1
			)
	}) %>%
	reduce(add_row) %>%
	mutate(
		geometry_len_meters = st_length(geometry) %>% as.numeric
	)

```

### Details

```{r}

# Display combined bikeway rows and cols
cat(glue(
	"\n* Rows: ", bike_combine %>% nrow,
	"\n* Columns: ", bike_combine %>% ncol
))

# Display rows per city
bike_combine_counts <- bike_combine %>%
	as_tibble %>%
	group_by(city) %>% count
bike_combine_counts

```

### Preview

```{r}

# Display combined bikeway data
bike_combine %>% as_tibble %>% head

```

## Combine KSI {.tabset}

Combine KSI data across cities using common columns.

**Note**: For Vancouver, columns were duplicated to represent individuals injured or killed.

**Note 2**: KSI records for Vancouver only had month and year - thus the day was set to the first of each month to be consistent with Toronto and Calgary data.

```{r}

# Vancouver ksi
vancksi <- vancksi_raw %>%
	uncount(VICTIM_COUNT) %>%
	st_as_sf(
		coords = c("LONGITUDE", "LATITUDE"),
		crs = 4326,
		na.fail = F
	) %>%
	mutate(
		ksi_city = "vancouver",
		ksi_id = 1:nrow(.) %>% as.character,
		ksi_date = as_date(glue("{YEAR}-{MONTH}-01")), # assume first day
		ksi_severity = INJURY_TYPE %>% as.character,
		ksi_user = ROLE %>% as.character
	) %>%
	select(
		ksi_city,
		ksi_id,
		ksi_date,
		ksi_severity,
		ksi_user
	)

# Calgary ksi
calgksi <- calgksi_raw %>%
	st_as_sf(
		coords = c("LOC_GPS_LONG", "LOC_GPS_LAT"),
		crs = 4326,
		na.fail = F
	) %>%
	mutate(
		ksi_city = "calgary",
		ksi_id = 1:nrow(.)  %>% as.character,
		ksi_date = OCCURENCE_TIMESTAMP,
		ksi_severity = INJURY_SEVERITY %>% as.character,
		ksi_user = Object.Type %>% as.character
	) %>%
	mutate(
		ksi_user = case_when(
			ksi_user == "Bicyclist" ~ "Cyclist",
			.default = ksi_user
		),
	) %>%
	select(
		ksi_city,
		ksi_id,
		ksi_date,
		ksi_severity,
		ksi_user
	)

# Toronto ksi
toronksi <- toronksi_raw %>%
	mutate(
		ksi_city = "toronto",
		ksi_id = INDEX %>% as.character,
		ksi_date = DATE %>% as_date,
		ksi_severity = ACCLASS %>% as.character,
		ksi_user = INVTYPE %>% as.character
	) %>%
	select(
		ksi_city,
		ksi_id,
		ksi_date,
		ksi_severity,
		ksi_user
	)

# Combine KSI data
ksi_combine <- calgksi %>%
	add_row(vancksi) %>%
	add_row(toronksi)

```

### Details

```{r}

# Display combined ksi rows and cols
cat(glue(
	"\n* Rows: ", ksi_combine %>% nrow,
	"\n* Columns: ", ksi_combine %>% ncol
))

# Display rows per city
ksi_combine_counts <- ksi_combine %>%
	as_tibble %>%
	group_by(ksi_city) %>%
	count
ksi_combine_counts

```

### Preview

```{r}

# Display combined ksi data
ksi_combine %>% as_tibble %>% select(-geometry) %>% head

```

## Filter KSI with Coordinates {.tabset}

Filter KSI data with coordinates only.

```{r}

# Filter for ksi with coords
ksi_coords <- ksi_combine %>%
	filter(!st_is_empty(geometry))

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_combine %>% nrow,
	"\n* Columns: ", ksi_combine %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_coords %>% nrow,
	"\n* Columns: ", ksi_coords %>% ncol
))

# Display missing coordinates per city
ksi_coords_counts <- ksi_combine %>%
	as_tibble %>%
	group_by(ksi_city) %>%
	count %>%
	left_join(
		ksi_coords %>%
			as_tibble %>%
			group_by(ksi_city) %>%
			count(name = "n_coords"),
		by = "ksi_city"
	) %>%
	mutate(
		n_miss_coords = n - n_coords,
		perc_coords = n_coords / n * 100,
		perc_miss_coords = n_miss_coords / n * 100
	)
ksi_coords_counts

```

### Preview

```{r}

# Display filtered rows
ksi_coords %>% as_tibble %>% select(-geometry) %>% head

```

## Filter KSI Severity {.tabset}

Filter KSI data for fatal and serious injuries only.

```{r}

# Filter for fatal or serious injuries only
ksi_injury <- ksi_coords %>%
	filter(ksi_severity %in% c(
		"Major Injuries",
		"Fatality",
		"Non-Fatal Injury",
		"Fatal",
		"Fatal injury",
		"Serious injury - Overnight at hospital"
	))

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_coords %>% nrow,
	"\n* Columns: ", ksi_coords %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_injury %>% nrow,
	"\n* Columns: ", ksi_injury %>% ncol
))

# Display counts
ksi_injury_counts <- ksi_coords %>%
	as_tibble %>%
	group_by(ksi_city, ksi_severity) %>%
	count
ksi_injury_counts

```

### Preview

```{r}

# Display filtered rows
ksi_injury %>% as_tibble %>% select(-geometry) %>% head

```

## Filter Verified Bikeways {.tabset}

Filter for bikeways with verified installations or upgrades.

```{r}

# Filter verified bikeways
bike_verify <- bike_combine %>%
	filter(!city %in% c("calgary", "vancouver")) %>%
	add_row( # calgary filters appendix 2
		bike_combine %>%
			filter(
				city == "calgary" &
				install_type %in% c("Bicycle Lane", "Cycle Track") &
				!build_status %in% c("INACTIVE", "PLANNED")
			)
	) %>%
	add_row( # vancouver filters appendix 2
		bike_combine %>%
	    filter(
	        city == "vancouver" &
	            install_type %in% c("Painted Lanes", "Protected Bike Lanes", "Local Street") &
	            road_type != "Off-street"
	    )
	) %>%
	filter(
		(!is.na(verify_install_year) |
		!is.na(verify_upgrade1_year) |
		!is.na(verify_upgrade2_year)) &
		is.na(no_verify_install_type) &
		is.na(verify_misclass) &
		!st_is_empty(geometry)
	) %>%
	distinct(geometry, .keep_all = TRUE) # remove duplicate geoms

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_combine %>% nrow,
	"\n* Columns: ", bike_combine %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_verify %>% nrow,
	"\n* Columns: ", bike_verify %>% ncol
))

# Display counts
bike_verify_counts <- bike_verify %>%
	as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(status, value) %>%
	count
bike_verify_counts

```

### Preview

```{r}

# Display filtered rows
bike_verify %>%
	as_tibble %>%
	select(
		verify_install_year,
		verify_upgrade1_year,
		verify_upgrade2_year,
		no_verify_install_type,
		everything()
	) %>%
	arrange(`verify_install_year`) %>%
	head

```

## Filter Infrastructure Types {.tabset}

Filter for painted lanes (`PL` and `BUF`) and cycle tracks (`PBL`) only in any of `verify_install_type`, `verify_upgrade1_type`, and `verify_upgrade2_type`.

```{r}

# Remove none or na bike types
bike_itypes <- bike_verify %>%
	filter(
		verify_install_type %in% c("PL", "BUF", "PBL") |
		verify_upgrade1_type %in% c("PL", "BUF", "PBL") |
		verify_upgrade2_type %in% c("PL", "BUF", "PBL")
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_verify %>% nrow,
	"\n* Columns: ", bike_verify %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_itypes %>% nrow,
	"\n* Columns: ", bike_itypes %>% ncol
))

# Display counts
bike_itypes_counts <- bike_itypes %>%
	as_tibble %>%
	filter(
		verify_install_type %in% c("PL", "BUF", "PBL")
	) %>%
	mutate(
		type = verify_install_type
	) %>%
	group_by(type) %>%
	count %>%
	ungroup %>%
	add_row(
		bike_itypes %>%
			as_tibble %>%
			filter(
				!verify_install_type %in% c("PL", "BUF", "PBL") &
				verify_upgrade1_type %in% c("PL", "BUF", "PBL")
			) %>%
			mutate(
				type = glue("{verify_install_type} (Install) to {verify_upgrade1_type} (1st Upgrade)")
			) %>%
			group_by(type) %>%
			count
	) %>%
	add_row(
		bike_itypes %>%
			as_tibble %>%
			filter(
				!verify_install_type %in% c("PL", "BUF", "PBL") &
				!verify_upgrade1_type %in% c("PL", "BUF", "PBL") &
				verify_upgrade2_type %in% c("PL", "BUF", "PBL")
			) %>%
			mutate(
				type = glue("{verify_install_type} to {verify_upgrade1_type} (1st Upgrade) to {verify_upgrade2_type} (2nd Upgrade)")
			) %>%
			group_by(type) %>%
			count
	)
bike_itypes_counts

```

### Preview

```{r}

# Preview filtered bike types data
bike_itypes %>% as_tibble %>% head

```

## Filter Post-2010 Bikeways {.tabset}

Filter for bikeways with a verified installations or upgrades after 2010.

```{r}

# Filter bikeways for post2010
bike_p2010 <- bike_itypes %>%
	filter(
		verify_install_year > 2010 |
		verify_upgrade1_year > 2010 |
		verify_upgrade2_year > 2010
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_itypes %>% nrow,
	"\n* Columns: ", bike_itypes %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_p2010 %>% nrow,
	"\n* Columns: ", bike_p2010 %>% ncol
))

# Display counts
bike_p2010_counts <- bike_p2010 %>%
	as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(status, value) %>%
	count
bike_p2010_counts

```

### Preview

```{r}

# Display filtered rows
bike_p2010 %>%
	as_tibble %>%
	select(
		verify_install_year,
		verify_upgrade1_year,
		verify_upgrade2_year,
		everything()
	) %>%
	arrange(`verify_install_year`) %>%
	head

```

## Filter 2009-2023 KSI {.tabset}

Filter for KSI records between 2009 and 2023 inclusively.

```{r}

# Filter ksi for post2010
ksi_y2009to2023 <- ksi_injury %>%
	filter(year(ksi_date) >= 2009 & year(ksi_date) <= 2023)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_injury %>% nrow,
	"\n* Columns: ", ksi_injury %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_y2009to2023 %>% nrow,
	"\n* Columns: ", ksi_y2009to2023 %>% ncol
))

# Display counts
ksi_y2009to2023_counts <- ksi_y2009to2023 %>%
	as_tibble %>%
	group_by(year(ksi_date)) %>%
	count
ksi_y2009to2023_counts

```

### Preview

```{r}

# Display filtered rows
ksi_y2009to2023 %>%
	as_tibble %>%
	select(
		ksi_date,
		everything()
	) %>%
	arrange(year(ksi_date)) %>%
	head

```

## Add Bikeway Dates {.tabset}

Join ambiguous verified install/upgrade dates (e.g. Jan 1/2022, 2022/02, Fall 2020) to manually cleaned dates with structured time units (e.g. days, months, quarters, semesters, ranges) and date formats (e.g. 2022-01-01, 2022-02-01).

```{r}

# Join clean dates to include structured date formats
bike_dates <- bike_p2010 %>%
	left_join( # install dates
		vdates_raw %>%
			rename_with(~ paste0("_install_", .x)),
		by = join_by(verify_install_date == `_install_verify_date_raw`)
	) %>%
	left_join( # upgrade1 dates
		vdates_raw %>%
			rename_with(~ paste0("_upgrade1_", .x)),
		by = join_by(verify_upgrade1_date == `_upgrade1_verify_date_raw`)
	) %>%
	left_join( # upgrade2 dates
		vdates_raw %>%
			rename_with(~ paste0("_upgrade2_", .x)),
		by = join_by(verify_upgrade2_date == `_upgrade2_verify_date_raw`)
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_p2010 %>% nrow,
	"\n* Columns: ", bike_p2010 %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_dates %>% nrow,
	"\n* Columns: ", bike_dates %>% ncol,
	"\n\nNew Columns (", bike_dates %>% ncol - bike_p2010 %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_dates) %>%
			.[!. %in% colnames(bike_p2010)],
		collapse = ""
	)
))

# Display counts
bike_dates_counts <- bike_dates %>%
    as_tibble %>%
    pivot_longer(
        c(
            verify_install_year,
            verify_upgrade1_year,
            verify_upgrade2_year
        ),
        names_to = "status",
        values_to = "status_year"
    ) %>%
    mutate(
        status = case_when(
            str_starts(status, "verify_install") ~ "install",
            str_starts(status, "verify_upgrade1") ~ "upgrade1",
            str_starts(status, "verify_upgrade2") ~ "upgrade2"
        )
    ) %>%
    group_by(city, status) %>%
    summarize(
        n = sum(!is.na(status_year))
    ) %>%
	left_join(
		bike_dates %>%
		    as_tibble %>%
		    pivot_longer(
		        c(
		            `_install_verify_date_type`,
		            `_upgrade1_verify_date_type`,
		            `_upgrade2_verify_date_type`
		        ),
		        names_to = "status",
		        values_to = "status_date_type"
		    ) %>%
		    mutate(
		        status = case_when(
		            str_starts(status, "_install") ~ "install",
		            str_starts(status, "_upgrade1") ~ "upgrade1",
		            str_starts(status, "_upgrade2") ~ "upgrade2"
		        )
		    ) %>%
		    group_by(city, status) %>%
		    summarize(
		        n_with_dates = sum(!is.na(status_date_type))
		    ),
		by = c("city", "status")
	) %>%
	mutate(
		n_without_dates = n - n_with_dates
	) %>%
	select(
		city,
		status,
		n_with_dates,
		n_without_dates,
		n
	)
bike_dates_counts

```

### Preview

```{r}

# Display clean dates cols
bike_dates %>%
	as_tibble %>%
	select(starts_with(c("_install_", "_upgrade1_", "_upgrade2_")), everything()) %>%
	arrange(`_install_verify_date_type`)

```

## Add Time Units {.tabset .tabset-pills}

Add the following time unit columns for installs/upgrades in bikeways and for each ksi record:

* `_time_month`: which month (1, 2, 3 ... 12) a bikeway had a verified installation/upgrade or ksi record
* `_time_quarter`: which quarter (1, 2, 3, 4) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_third`: which third (1, 2, 3) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_half`: which half (1, 2) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_semester`: which semester (1, 2) of the year a bikeway had a verified installation/upgrade or ksi record, where:
	* `1`: represents May to October of this year
	* `2`: represents November (this year) to April of next year
* `_time_year`: standardized year column that bikeway had verified installation/upgrade or ksi record
* `_time_..._group`: grouped columns of the above defining the time units in textual format
* Install records follow the patterns above prefixed with `_install`
* Upgrade records are prefixed with `_upgrade1` and `_upgrade2`
* KSI records will follow the same patterns above except prefixed with `_ksi` as well
* Additional columns for installs and upgrades only:
	* `_assign_semester`: if the date falls between semester 1 or 2 and was automatically assigned to semester 1 or 2 based on the larger number of days in either semester 1 or 2
	* `_assign_semester_1_days`: number of days in semester 1 if the date falls between semesters
	* `_assign_semester_2_days`: number of days in semester 2 if the date falls between semesters
	* If the number of days are equal, then semester 2 is assigned by default

**Note**: Date ranges that fall between months, quarters, thirds, or halves of the year were excluded from being classified as any of the time units (e.g. March 31 to April 15 will be excluded as it does not fall within either the 1st or 2nd quarter of the year).

```{r}

#' Add Time Units to Data Frame
#'
#' This function takes a data frame and adds new columns representing various time units such as month, quarter, third, half, and semester. These columns are derived from a date column or if the time unit is more than day, then additiona start and end date columns
#'
#' @param df A data frame that contains the columns `verify_date`, `verify_date_type`, `verify_date_start`, and `verify_date_end`.
#' @param date_col Name of the date column to use
#' @param start_col Name of the start date column to use
#' @param end_col Name of the end date column to use
#' @param type_col Name of the date type column to use (e.g. month, day, year, etc)
#' @param prefix Additional text to add to the output column names
#' @param assign_semester Whether to assign intervals between semesters to be in the semester with the max duration that the interval falls in. When durations are equal, semester 2 will be chosen.
#'
#' @return A data frame with additional columns (without `prefix`):
#' \describe{
#'   \item{`_time_month`}{Numeric representation of the month (1-12).}
#'   \item{`_time_month_group`}{Unique time group for month.}
#'   \item{`_time_quarter`}{Quarter of the year (1-4).}
#'   \item{`_time_quarter_group`}{Unique time group for quarter.}
#'   \item{`_time_third`}{Third of the year (1-3).}
#'   \item{`_time_third_group`}{Unique time group for third.}
#'   \item{`_time_half`}{Half of the year (1-2).}
#'   \item{`_time_half_group`}{Unique time group for half.}
#'   \item{`_time_semester`}{Custom semester (1 for May-Oct, 2 for Nov-Apr).}
#'   \item{`_assign_semester`}{Whether the date was assigned since it did not fall between either semester 1 or 2.}
#'   \item{`_assign_semester_1_days`}{Number of days that the date is in semester 1 if `assign_semester` is `TRUE`.}
#'   \item{`_assign_semester_2_days`}{Number of days that the date is in semester 2 if `assign_semester` is `TRUE`.}
#'   \item{`_time_semester_group`}{Unique time group for semester.}
#'   \item{`_time_year`}{4-digit year.}
#' }
#'
#' @examples
#' \dontrun{
#' df <- data.frame(
#'   verify_date = as.Date(c("2021-01-15", "2021-06-20", "2021-09-10")),
#'   verify_date_type = c("day", "month", "day"),
#'   verify_date_start = as.Date(c("2021-01-01", "2021-06-01", "2021-09-01")),
#'   verify_date_end = as.Date(c("2021-01-31", "2021-06-30", "2021-09-30"))
#' )
#' result <- add_time_units(df)
#' print(result)
#' }
#'
#' @import dplyr
#' @importFrom lubridate month year
#' @export
add_time_units <- function(
	df,
	date_col = "_install_verify_date",
	start_col = "_install_verify_date_start",
	end_col = "_install_verify_date_end",
	type_col = "_install_verify_date_type",
	year_col = "verify_install_year",
	prefix = NULL,
	assign_semester = T
) {
	
	# Code time units
	out <- df %>%
		mutate( # rename req cols
			`_date` = .data[[date_col]],
			`_start` = .data[[start_col]],
			`_end` = .data[[end_col]],
			`_type` = .data[[type_col]],
			`_year` = .data[[year_col]]
		) %>%
		mutate( # add time units
			
			# Months (monthly)
			`_time_month` = case_when(
				`_type` %in% c("day", "month") ~ month(`_date`, label = T, abbr = F)
			),
			`_time_month` = as.numeric(`_time_month`),
			
			# Quarters (quarterly)
			`_time_quarter` = case_when(
				month(`_date`) %in% 1:3 |
				(
					month(`_start`) %in% 1:3 &
					month(`_end`) %in% 1:3 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% 4:6 |
				(
					month(`_start`) %in% 4:6 &
					month(`_end`) %in% 4:6 &
					year(`_start`) == year(`_end`)
				) ~ 2,
				month(`_date`) %in% 7:9 |
				(
					month(`_start`) %in% 7:9 &
					month(`_end`) %in% 7:9 &
					year(`_start`) == year(`_end`)
				) ~ 3,
				month(`_date`) %in% 10:12 |
				(
					month(`_start`) %in% 10:12 &
					month(`_end`) %in% 10:12 &
					year(`_start`) == year(`_end`)
				) ~ 4
			),
			
			# Thirds (triyearly)
			`_time_third` = case_when(
				month(`_date`) %in% 1:4 |
				(
					month(`_start`) %in% 1:4 &
					month(`_end`) %in% 1:4 &
					year(`_start`) == year(`_end`)
				) ~ 1, # Fall
				month(`_date`) %in% 5:8 |
				(
					month(`_start`) %in% 5:8 &
					month(`_end`) %in% 5:8 &
					year(`_start`) == year(`_end`)
				) ~ 2, # Winter
				month(`_date`) %in% 9:12 |
				(
					month(`_start`) %in% 9:12 &
					month(`_end`) %in% 9:12 &
					year(`_start`) == year(`_end`)
				) ~ 3 # Spring/Summer
			),
			
			# Halves (biyearly)
			`_time_half` = case_when(
				month(`_date`) %in% 1:6 |
				(
					month(`_start`) %in% 1:6 &
					month(`_end`) %in% 1:6 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% 7:12 |
				(
					month(`_start`) %in% 7:12 &
					month(`_end`) %in% 7:12 &
					year(`_start`) == year(`_end`)
				) ~ 2
			),
			
			# Semester (custom range)
			`_time_semester` = case_when(
				month(`_date`) %in% 5:10 | # May to Oct (this year)
				(
					month(`_start`) %in% 5:10 &
					month(`_end`) %in% 5:10 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% c(11:12, 1:4) | # Nov (this/last year) to Apr (next/this year)
				(
					month(`_start`) %in% c(11:12, 1:4) &
					month(`_end`) %in% c(11:12, 1:4) &
					year(`_start`) <= (year(`_end`) + 1)
				) ~ 2
			),
			
			# Year
			`_time_year` = `_year`
			
		)
	
	# For intervals between semesters, assign in max duration
	if (assign_semester) {
		out <- out %>%
			mutate( # add intervals for semesters and date
				`_assign_semester_1_interval_1` = interval( # sem 1 this year
		            start = glue("{`_year`}-05-01") %>% as_date,
		            end = glue("{`_year`}-10-31") %>% as_date
		        ),
		        `_assign_semester_2_interval_1` = interval( # sem 2 this to next year
		            start = glue("{`_year`}-11-01") %>% as_date,
		            end = glue("{`_year` + 1}-04-30") %>% as_date
		        ),
				`_assign_semester_2_interval_2` = interval( # sem 2 last to this year
		            start = glue("{`_year` - 1}-11-01") %>% as_date,
		            end = glue("{`_year`}-04-30") %>% as_date
		        ),
				`_assign_date_interval` = if_else( # only for non-assigned sem
					!is.na(`_start`) & !is.na(`_end`) & is.na(`_time_semester`),
					interval(`_start`, `_end`),
					NA
				)
			) %>%
			mutate( # get duration in days
				`_assign_semester_1_days` = coalesce(
					intersect(
						`_assign_date_interval`,
						`_assign_semester_1_interval_1`
					)
				) %>% as.numeric / (3600 * 24),
				`_assign_semester_2_days` = coalesce(
					intersect(
						`_assign_date_interval`,
						`_assign_semester_2_interval_1`
					),
					intersect(
						`_assign_date_interval`,
						`_assign_semester_2_interval_2`
					)
				) %>% as.numeric / (3600 * 24)
			) %>%
			mutate( # assign semester
				`_time_semester` = if_else(
					!is.na(`_assign_date_interval`),
					case_when(
						`_assign_semester_1_days` == `_assign_semester_2_days` ~ 2, # break ties to sem 2
						`_assign_semester_1_days` > `_assign_semester_2_days` ~ 1, # sem 1
						`_assign_semester_1_days` < `_assign_semester_2_days` ~ 2 # sem 2
					),
					`_time_semester`
				),
				`_assign_semester` = if_else(
					!is.na(`_assign_semester_1_days`) | !is.na(`_assign_semester_2_days`),
					TRUE,
					FALSE
				)
			) %>%
			select(-c( # remove unwanted cols
				`_assign_semester_1_interval_1`,
				`_assign_semester_2_interval_1`,
				`_assign_semester_2_interval_2`,
				`_assign_date_interval`
			))
	}
	
	# Add time groups that include the year
	out <- out %>%
		mutate(
			
			# Month group
			`_time_month_group` = if_else(
				!is.na(`_time_month`),
				glue("{`_year`}_{`_time_month`}"),
				NA
			),
			
			# Quarter group
			`_time_quarter_group` = if_else(
				!is.na(`_time_quarter`),
				glue("{`_year`}_{`_time_quarter`}"),
				NA
			),
			
			# Third group
			`_time_third_group` = if_else(
				!is.na(`_time_third`),
				glue("{`_year`}_{`_time_third`}"),
				NA
			),
			
			# Half group
			`_time_half_group` = if_else(
				!is.na(`_time_half`),
				glue("{`_year`}_{`_time_half`}"),
				NA
			),
			
			# Semester group
			`_time_semester_group` = if_else(
				!is.na(`_time_semester`),
				glue("{`_year`}_{`_time_semester`}"),
				NA
			),
			
			# Year group
			`_time_year_group` = if_else(
				!is.na(`_time_year`),
				glue("{`_year`}"),
				NA
			)
			
		)
	
	# Add prefix to time unit cols
	if (!is.null(prefix)) {
		out <- out %>%
			rename_with( # add prefix
				~ paste0(prefix, .x),
				starts_with("_time_")
			)
	}
	
	# Remove time cols
	out <- out %>%
		select(-c(
			`_date`,
			`_start`,
			`_end`,
			`_type`,
			`_year`
		))
	return(out)
}

```

### Bikeways {.tabset}

```{r}

# Add time unit columns using func for bikeways
bike_tunits <- bike_dates %>%
	add_time_units( # install tunits
		date_col = "_install_verify_date",
		start_col = "_install_verify_date_start",
		end_col = "_install_verify_date_end",
		type_col = "_install_verify_date_type",
		year_col = "verify_install_year",
		prefix = "_install"
	) %>%
	add_time_units( # upgrade1 tunits
		date_col = "_upgrade1_verify_date",
		start_col = "_upgrade1_verify_date_start",
		end_col = "_upgrade1_verify_date_end",
		type_col = "_upgrade1_verify_date_type",
		year_col = "verify_upgrade1_year",
		prefix = "_upgrade1"
	) %>%
	add_time_units( # upgrade2 tunits
		date_col = "_upgrade2_verify_date",
		start_col = "_upgrade2_verify_date_start",
		end_col = "_upgrade2_verify_date_end",
		type_col = "_upgrade2_verify_date_type",
		year_col = "verify_upgrade2_year",
		prefix = "_upgrade2"
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_dates %>% nrow,
	"\n* Columns: ", bike_dates %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_tunits %>% nrow,
	"\n* Columns: ", bike_tunits %>% ncol,
	"\n\nNew Columns (", bike_tunits %>% ncol - bike_dates %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_tunits) %>%
			.[!. %in% colnames(bike_dates)],
		collapse = ""
	)
))

# Display counts
bike_tunits_counts <- bike_tunits %>%
    as_tibble %>%
	pivot_longer(
		ends_with("_group"),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "_install") ~ "install",
			str_starts(column, "_upgrade1") ~ "upgrade1",
			str_starts(column, "_upgrade2") ~ "upgrade2"
		),
		time_unit = str_replace_all(
			column,
			"_install_time_|_upgrade1_time_|_upgrade2_time_|_group",
			""
		)
	) %>%
	filter(!is.na(value)) %>%
	group_by(status, time_unit) %>%
	count
bike_tunits_counts

```

#### Preview

```{r}

# Display time unit cols for bikeways
bike_tunits %>%
	as_tibble %>%
	select(
		starts_with("_"),
		everything()
	) %>%
	arrange(`_install_verify_date`) %>%
	head

```

### KSI {.tabset}

```{r}

# Add time unit columns using func for ksi
ksi_tunits <- ksi_y2009to2023 %>%
	mutate( # add cols to standardize func params
		`_ksi_date` = ksi_date,
		`_ksi_date_start` = NA,
		`_ksi_date_end` = NA,
		`_ksi_date_type` = "day",
		`_ksi_year` = year(ksi_date)
	) %>%
	add_time_units(
		date_col = "_ksi_date",
		start_col = "_ksi_date_start",
		end_col = "_ksi_date_end",
		type_col = "_ksi_date_type",
		year_col = "_ksi_year",
		prefix = "_ksi",
		assign_semester = F
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_y2009to2023 %>% nrow,
	"\n* Columns: ", ksi_y2009to2023 %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_tunits %>% nrow,
	"\n* Columns: ", ksi_tunits %>% ncol,
	"\n\nNew Columns (", ksi_tunits %>% ncol - ksi_y2009to2023 %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(ksi_tunits) %>%
			.[!. %in% colnames(ksi_injury)],
		collapse = ""
	)
))

# Display counts
ksi_tunits_counts <- ksi_tunits %>%
    as_tibble %>%
	pivot_longer(
		ends_with("_group"),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		time_unit = str_replace_all(
			column,
			"_ksi_time_|_group",
			""
		)
	) %>%
	filter(!is.na(value)) %>%
	group_by(time_unit) %>%
	count
ksi_tunits_counts

```

#### Preview

```{r}

# Display time unit cols for ksi
ksi_tunits %>%
	as_tibble %>%
	select(
		ksi_date,
		starts_with("_")
	) %>%
	arrange(year(ksi_date)) %>%
	head

```

## Determine Time Unit {.tabset .tabset-pills}

For bikeways, determine the temporal resolution (unit of time) finer than yearly based on the amount of data available per time unit (sorted from the highest resolution time unit to the lowest resolution time unit):

* **Month** (high): refers to verified post-2010 installs and upgrades occurring approximately in a month (1/12) of a year
* **Quarter**: refers to verified post-2010 installs and upgrades occurring approximately in a quarter (1/4) of a year
* **Third**: refers to verified post-2010 installs and upgrades occurring approximately in a third (1/3) of a year
* **Half** (low): refers to verified post-2010 installs and upgrades occurring approximately in a half (1/2) of a year
* **Semester** (low): refers to verified post-2010 installs and upgrades occurring approximately in a half (1/2) of a year with the first half (1) being November (same year) to April (next year) and the second half being May to October (next year)

The `year` was selected as the time unit.

```{r}

plot_time_units <- function(
		df,
		month_col = "_install_time_month",
		quarter_col = "_install_time_quarter",
		third_col = "_install_time_third",
		half_col = "_install_time_half",
		semester_col = "_install_time_semester",
		year_col = "_install_time_year",
		title = "% of Verified Post-2010 Installs by Unit of Time",
		overlap_gap = 10.5,
		expand = c(0.18, 0),
		city_nudge_x = 0.25
) {
	
	# Prepare plot data
	plot_data <- df %>%
		as_tibble %>%
		mutate( # assign time unit cols
			`_time_month` = .data[[month_col]],
			`_time_quarter` = .data[[quarter_col]],
			`_time_third` = .data[[third_col]],
			`_time_half` = .data[[half_col]],
			`_time_semester` = .data[[semester_col]],
			`_time_year` = .data[[year_col]]
		) %>%
		group_by(city) %>%
		summarize( # calc installs/upgrades for each time unit
			Month = sum(!is.na(`_time_month`)),
			Quarter = sum(!is.na(`_time_quarter`)),
			Third = sum(!is.na(`_time_third`)),
			Half = sum(!is.na(`_time_half`)),
			Semester = sum(!is.na(`_time_semester`)),
			Year = sum(!is.na(`_time_year`))
		) %>%
		pivot_longer(
			cols = -city,
			names_to = "type",
			values_to = "n"
		) %>%
		mutate( # zero as NA
			n = if_else(n == 0, NA, n)
		) %>%
		left_join( # Add city totals
			df %>%
				as_tibble %>%
				group_by(city) %>%
				count(name = "total"),
			by = "city"
		) %>%
		ungroup %>%
	    mutate( # calc percentages and add labels
	        type = factor(type, levels = c("Month", "Quarter", "Third", "Half", "Semester", "Year")),
	        city = factor(str_to_title(city), levels = c("Vancouver", "Calgary", "Toronto")),
	        perc = n / total * 100,
	        perc_label = glue(
	        	"{str_sub(city, end = 1)}: {round(perc, 2)}%\n",
	        	"(n={format(n, big.mark = ',', scientific = F)})"
	        )
	    ) %>%
	    group_by(type) %>% # adjust overlapping labels
	    arrange(desc(perc)) %>% 
		mutate(
	        perc_before = lag(perc),
	        perc_after = lead(perc),
	        perc_order = 0:(length(perc) - 1)
	    ) %>%
	    mutate( # detect overlap labels and shift down and up
	        perc_label_y = case_when(
	            lag(perc) == perc | lead(perc) == perc ~ max(perc) - (perc_order * overlap_gap),
	            lag(perc) - perc < 10 ~ max(perc) - (perc_order * overlap_gap),
	            .default = perc
	        )
	    ) %>%
		group_by(city) %>%
		mutate( # add city labels at end of lines
			city_label = if_else(
				type == "Year",
				glue(
					"{city}\n",
					"(n={format(total, big.mark = ',', scientific = F)})"
				),
				NA
			)
		) %>%
		ungroup %>%
	    arrange(city, type)
	
	# Get time unit totals
	unit_total <- sum(plot_data$total %>% unique, na.rm = T)
	unit_total_label <- format(unit_total, big.mark = ",", scientific = F)
	
	# Plot time units
	out <- plot_data %>%
		ggplot(aes(
			x = type,
			y = perc,
			group = city,
			color = city,
			label = perc_label
		)) +
		geom_line(
			alpha = 0.8,
			linewidth = 1.5
		) +
		geom_label(
			aes(y = perc_label_y),
			size = 2.25,
			show.legend = F,
			label.padding = unit(0.5, "lines")
		) +
		geom_text(
			aes(
				label = city_label,
				y = perc_label_y
			),
			size = 2.5,
			show.legend = F,
			hjust = 0,
			nudge_x = city_nudge_x
		) +
		scale_y_continuous(
			limits = c(0, 100),
			labels = function(x) paste0(x, "%")
		) +
		scale_x_discrete(
			limits = levels(plot_data$type),
			position = "top",
			expand = expand
		) +
		labs(
			title = glue(
				"{title}\n",
				"(n={unit_total_label})"
			),
			x = NULL,
			y = NULL,
			color = "City",
			label = NULL,
			group = NULL
		) +
		theme_minimal() +
		theme(
			legend.position = "none",
			plot.title = element_text(hjust = 0.5)
		)
	return(out)
}

# Select time unit
tunit <- "year"

```

### Install {.tabset}

#### Plot

```{r}
plot_time_units(bike_tunits)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	group_by(city, `_install_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

### 1st Upgrade {.tabset}

#### Plot

```{r}
plot_time_units(
	bike_tunits %>% filter(!is.na(verify_upgrade1_year)),
	month_col = "_upgrade1_time_month",
	quarter_col = "_upgrade1_time_quarter",
	third_col = "_upgrade1_time_third",
	half_col = "_upgrade1_time_half",
	semester_col = "_upgrade1_time_semester",
	year_col = "_upgrade1_time_year",
	title = "% of Verified Post-2010 1st Upgrades by Unit of Time"
)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	filter(!is.na(verify_upgrade1_year)) %>%
	group_by(city, `_upgrade1_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

### 2nd Upgrade {.tabset}

#### Plot

```{r}
plot_time_units(
	bike_tunits %>% filter(!is.na(verify_upgrade2_year)),
	month_col = "_upgrade2_time_month",
	quarter_col = "_upgrade2_time_quarter",
	third_col = "_upgrade2_time_third",
	half_col = "_upgrade2_time_half",
	semester_col = "_upgrade2_time_semester",
	year_col = "_upgrade2_time_year",
	title = "% of Verified Post-2010 2nd Upgrades by Unit of Time",
	expand = c(0.2, 0),
	city_nudge_x = 0.35
)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	filter(!is.na(verify_upgrade2_year)) %>%
	group_by(city, `_upgrade1_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

## Filter Timed Bikeways {.tabset}

Filter for bikeways with time unit values that were installed/upgraded post-2010.

```{r}
bike_tfilter <- bike_tunits %>%
	filter(
		(verify_install_year > 2010 & !is.na(.data[[glue("_install_time_{tunit}")]])) |
		(verify_upgrade1_year > 2010 & !is.na(.data[[glue("_upgrade1_time_{tunit}")]])) |
		(verify_upgrade2_year > 2010 & !is.na(.data[[glue("_upgrade1_time_{tunit}")]])) 
	)
```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_tunits %>% nrow,
	"\n* Columns: ", bike_tunits %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_tfilter %>% nrow,
	"\n* Columns: ", bike_tfilter %>% ncol
))

# Display counts
bike_sem_counts <- bike_tfilter %>%
    as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(city, status, value) %>%
	count
bike_sem_counts

```

### Preview

```{r}

# Display merged bike segments
bike_tfilter %>%
	as_tibble %>%
	select(-geometry) %>%
	head

```

## Add Improvements

Add whether there were bikeway improvements or not for each bikeway segment.

Bikeway improvements occur when the installation was upgraded either once or twice, but resulted in the same bikeway type.

For the first upgrade, this is `TRUE` when the installation is the same type as the first upgrade.

For the second upgrade, this is `TRUE` when the first upgrade or installation is the same type as the second upgrade.

```{r}

# Join clean dates to include structured date formats
bike_improv <- bike_tfilter %>%
	mutate(
		`_improve1` = if_else( # whether upgrade1 was an improvement
			!is.na(verify_upgrade1_type) &
			verify_install_type == verify_upgrade1_type,
			TRUE,
			FALSE
		),
		`_improve2` = if_else( # whether upgrade2 was an improvement
			!is.na(verify_upgrade2_type) &
			(verify_upgrade2_type == verify_upgrade1_type |
			verify_upgrade2_type == verify_install_type),
			TRUE,
			FALSE
		),
		`_improve` = `_improve1` | `_improve2` # whether there was an improvement
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_tfilter %>% nrow,
	"\n* Columns: ", bike_tfilter %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_improv %>% nrow,
	"\n* Columns: ", bike_improv %>% ncol,
	"\n\nNew Columns (", bike_improv %>% ncol - bike_tfilter %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_improv) %>%
			.[!. %in% colnames(bike_tfilter)],
		collapse = ""
	)
))

# Display counts
bike_improv_counts <- bike_improv %>%
	mutate(
		history = paste0(
			verify_install_type,
			" -> ",
			verify_upgrade1_type,
			" -> ",
			verify_upgrade2_type
		)
	) %>%
	as_tibble() %>%
	select(-geometry) %>%
	arrange(history) %>%
	group_by(history) %>%
	summarize(
		improve1 = sum(`_improve1`),
		improve2 = sum(`_improve2`),
		n = n()
	) %>%
	add_row(
		history = "TOTAL",
		improve1 = sum(.[["improve1"]]),
		improve2 = sum(.[["improve2"]]),
		n = sum(.[["n"]])
	)
bike_improv_counts

```

### Preview

```{r}

# Display clean dates cols
bike_improv %>%
	as_tibble() %>%
	select(-geometry) %>%
	select(
		starts_with("_improve"),
		verify_install_type,
		verify_upgrade1_type,
		verify_upgrade2_type,
		everything()
	) %>%
	head()

```

## Merge Bikeways {.tabset}

Merge bikeway segments if they are within the same infrastructure type, time unit, and known status (e.g. install, 1st upgrade, 2nd upgrade).

The known status is the most recent change **with a known time unit** for a segment with one of the following values:

1. `install`: an installed segment with no upgrades
2. `upgrade1` a segment upgraded once and was previously installed
3. `upgrade2`: a segment upgraded twice and was previously upgrade once and installed

This also adds an additional column `_merge_group` specifying the merge group in the format of `<CITYNUM>_<STATUS>_<TYPE>_<YEAR>_<TIME>_<ADJID>`, where:

* `<CITYNUM>`: is the numeric id of the city, 1 is vancouver, 2 is calgary, and 3 is toronto
* `<STATUS>`: the status of the segment, one of `install`, `upgrade1`, or `upgrade2`
* `<TYPE>`: the infrastructure type of the segment
* `<YEAR>`: the year of the segment status
* `<TIME>`: the value of the time unit if not the same as year
* `<ADJID>`: the adjacency id, if the ids are the same, they are adjacent in space

Other merge related columns start with `_merge_...`.

**Note 1**: Only infrastructure with different types are considered an upgrade.

**Note 2**: This also re-calculates the bikeway lengths after merging segments.

```{r}

# Determine adjacent segments if they are within same status, type, and time unit
bike_adj <- bike_improv %>%
	filter( # remove segments that do not have the time unit
		!is.na(.data[[glue("_upgrade2_time_{tunit}")]]) |
		!is.na(.data[[glue("_upgrade1_time_{tunit}")]]) |
		!is.na(.data[[glue("_install_time_{tunit}")]])
	) %>%
	mutate( # add column to store status, type, and time unit
		`_merge_status` = case_when(
			!is.na(.data[[glue("_upgrade2_time_{tunit}")]]) &
				verify_upgrade2_type != verify_upgrade1_type & # up2 == up1
				verify_upgrade2_type != verify_install_type ~ # up2 == inst
				"upgrade2",
			!is.na(.data[[glue("_upgrade1_time_{tunit}")]]) &
				verify_upgrade1_type != verify_install_type ~ # up1 == inst
				"upgrade1",
			!is.na(.data[[glue("_install_time_{tunit}")]]) ~
				"install"
		),
		`_merge_type` = case_when(
			`_merge_status` == "upgrade2" ~ verify_upgrade2_type,
			`_merge_status` == "upgrade1" ~ verify_upgrade1_type,
			`_merge_status` == "install" ~ verify_install_type
		),
		`_merge_time_unit` = tunit,
		`_merge_time` = case_when(
			`_merge_status` == "upgrade2" ~ .data[[glue("_upgrade2_time_{tunit}")]],
			`_merge_status` == "upgrade1" ~ .data[[glue("_upgrade1_time_{tunit}")]],
			`_merge_status` == "install" ~ .data[[glue("_install_time_{tunit}")]]
		),
		`_merge_time_group` = case_when(
			`_merge_status` == "upgrade2" ~ .data[[glue("_upgrade2_time_{tunit}_group")]],
			`_merge_status` == "upgrade1" ~ .data[[glue("_upgrade1_time_{tunit}_group")]],
			`_merge_status` == "install" ~ .data[[glue("_install_time_{tunit}_group")]]
		),
		`_merge_year` = case_when(
			`_merge_status` == "upgrade2" ~ `_upgrade2_time_year`,
			`_merge_status` == "upgrade1" ~ `_upgrade1_time_year`,
			`_merge_status` == "install" ~ `_install_time_year`
		)
	) %>%
	group_by(city, `_merge_status`, `_merge_type`, `_merge_time_group`) %>%
	group_map(~ {
		
		# Assign group id as group_by values separated by underscore
		group_id <- .y[1,] %>%
			as.character %>%
			paste0(., collapse = "_") %>%
			str_to_lower(.)
		
		# Find adjacent segments and group them together using a group id
		adj_group <- .x %>%
			st_touches %>%
			graph_from_adj_list %>%
			components %>%
			.$membership %>%
			paste0(group_id, "_", .)
		
		# Assign adj group to data in order
		out <- .x %>% mutate(`_merge_group` = adj_group)
		return(out)
		
	}, .keep = T) %>%
	reduce(add_row)

# Merge bike segments by adjacency
bike_merge <- bike_adj %>%
	group_by(`_merge_group`) %>%
	summarize(
		id = paste0(id, collapse = ","),
		city = paste0(unique(city), collapse = ","),
		status = case_when( # assign single status if possible
			all(unique(`_merge_status`) == "upgrade2") ~ "upgrade2",
			all(unique(`_merge_status`) == "upgrade1") ~ "upgrade1",
			all(unique(`_merge_status`) == "install") ~ "install",
			.default = paste(unique(`_merge_status`), collapse = ",") # multiple status
		),
		type = if_else( # assign single type if possible
			length(unique(`_merge_type`)) == 1,
			unique(`_merge_type`) %>% as.character, # single type
			paste(unique(`_merge_type`), collapse = ",") # multiple types
		),
		year = if_else( # assign single year if possible
			length(unique(`_merge_year`)) == 1,
			unique(`_merge_year`) %>% as.character, # single year
			paste(unique(`_merge_year`), collapse = ",") # multiple years
		),
		time_unit = tunit,
		time_value = if_else( # assign single time if possible
			length(unique(`_merge_time`)) == 1,
			unique(`_merge_time`) %>% as.character, # single time
			paste(unique(`_merge_time`), collapse = ",") # multiple times
		),
		time_group = if_else( # assign single time if possible
			length(unique(`_merge_time_group`)) == 1,
			unique(`_merge_time_group`) %>% as.character, # single time
			paste(unique(`_merge_time_group`), collapse = ",") # multiple times
		),
		across(everything(), ~ paste0(.x, collapse = ","))
	) %>%
	mutate(
		geometry_len_meters = st_length(geometry) %>% as.numeric
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_improv %>% nrow,
	"\n* Columns: ", bike_improv %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nNew Columns (", bike_merge %>% ncol - bike_improv %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_merge) %>%
			.[!. %in% colnames(bike_improv)],
		collapse = ""
	)
))

# Display counts
bike_merge_counts <- bike_merge %>%
    as_tibble %>%
	group_by(city, status) %>%
	count
bike_merge_counts

```

### Preview

```{r}

# Display merged bike segments
bike_merge %>%
	as_tibble %>%
	select(-geometry) %>%
	head

```

### Map

```{r}
tmap_mode("view")
tm_shape(bike_merge %>% st_buffer(25)) +
    tm_polygons(
    	group = "Merged (Black)",
    	col = "black",
    	border.col = "white",
    	popup.vars = T,
    	lwd = 1.5
    ) +
	tm_shape(bike_adj) +
	tm_lines(
		group = "Segments (Red)",
		col = "red",
		lwd = 1.5
	) +
	tm_shape(bike_adj %>% st_cast("POINT")) +
	tm_dots(group = NULL, col = "red", size = 0.03)
```


## Join KSI to Bikeways {.tabset}

Spatially join KSI to nearest bikeways.

Each KSI point is joined to a bikeway if they are within 25 meters of the nearest bikeway

**Note**: if there is overlapping of bikeways, the same KSI can be assigned to multiple bikeways.

```{r}

# Buffer bikeways and keep ids
bike_merge_buff25 <- bike_merge %>%
	select(city, id) %>%
	st_buffer(25)

# Spatial join bikeways with ksi keeping ids
sjoin <- bike_merge_buff25 %>%
	st_join(ksi_tunits %>% select(ksi_city, ksi_id)) %>%
	as_tibble %>%
	select(-geometry)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBikeways\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nKSI\n",
	"\n* Rows: ", ksi_tunits %>% nrow,
	"\n* Columns: ", ksi_tunits %>% ncol,
	"\n\nSpatial Join Keys\n",
	"\n* Rows: ", sjoin %>% nrow,
	"\n* Columns: ", sjoin %>% ncol,
	"\n* Rows Spatially Joined (Bikeways): ",
	sjoin %>%
		filter(!is.na(id)) %>%
		pull(id) %>%
		unique %>%
		length,
	"\n* Rows Spatially Joined (KSI): ",
	sjoin %>%
		filter(!is.na(ksi_id)) %>%
		pull(ksi_id) %>%
		unique %>%
		length
))

# Display 
sjoin_counts <- sjoin %>%
	group_by(city) %>%
	summarize(
		n_bike = n_distinct(id, na.rm = T)
	) %>%
	left_join(
		sjoin %>%
			group_by(ksi_city) %>%
			summarize(
				n_ksi = n_distinct(ksi_id, na.rm = T)
			),
		by = join_by(city == ksi_city)
	)
sjoin_counts

```

### Preview

```{r}

# Display spatially joined ids
sjoin %>%
	filter(!is.na(ksi_id)) %>%
	head

```

## Count KSI Per Time Unit {.tabset}

For each merged bikeway, count KSI spatially joined to bikeways per unit of time.

This creates columns in the format of `_ksi_time_<UNIT>_<YEAR>_<VALUE>`, where:

* `<UNIT>`: is the unit of time, either `month`, `quarter`, `half`, or `semester`
* `<YEAR>`: is the year for the unit of time
* `<VALUE>`: is the number representing the portion of the time range for the unit of time with month being 1-12, quarter being 1-4, third being 1-3, half being 1-2 and semester being 1-2 (this is omitted if time unit is `year`)

**Note**: KSI by road users will also be counted.

```{r}

# Ensure ids are same data type
sjoin <- sjoin %>%
	mutate(
		id = as.character(id),
		ksi_id = as.character(ksi_id)
	)
ksi_tunits <- ksi_tunits %>%
	mutate(
		ksi_id = as.character(ksi_id)
	)

# Get time unit col
tunit_col <- glue("_ksi_time_{tunit}_group")

# Count ksi spatially joined to bike
bike_sjoin <- bike_merge %>%
	mutate(
		id = as.character(id)
	) %>%
	full_join( # add spatial joins
		sjoin,
		by = join_by(city, id)
	) %>%
	full_join( # add ksi
		ksi_tunits %>%
			as_tibble %>%
			select(-geometry),
		by = join_by(ksi_city, ksi_id)
	) %>%
	mutate( # add ksi aggregates
		ksi_user_driver = if_else(
			ksi_user %in% c("Driver", "Passenger", "Truck Driver"), 1, 0
		),
		ksi_user_pedestrian = if_else(
			ksi_user == "Pedestrian", 1, 0
		),
		ksi_user_cyclist = if_else(
			ksi_user %in% c("Cyclist", "Cyclist Passenger"), 1, 0
		),
		ksi_user_other = (!(ksi_user_driver | ksi_user_pedestrian | ksi_user_cyclist)) %>% as.integer
	) %>%
	group_by(across(c(
		"city",
		"id",
		tunit_col
	))) %>%
	summarize(
		across(
			-c("ksi_user_driver", "ksi_user_pedestrian", "ksi_user_cyclist", "ksi_user_other"),
			~ paste0(unique(.x), collapse = ",")
		),
		ksi_count = n(),
		ksi_count_driver = sum(ksi_user_driver, na.rm = T),
		ksi_count_pedestrian = sum(ksi_user_pedestrian, na.rm = T),
		ksi_count_cyclist = sum(ksi_user_cyclist, na.rm = T),
		ksi_count_other = sum(ksi_user_other, na.rm = T)
	) %>%
	mutate( # correct for no ksi joined
		ksi_id = if_else(ksi_id == "NA", NA, ksi_id),
		ksi_count = if_else(is.na(ksi_id), 0, ksi_count),
		ksi_count_driver = if_else(is.na(ksi_id), 0, ksi_count_driver),
		ksi_count_pedestrian = if_else(is.na(ksi_id), 0, ksi_count_pedestrian),
		ksi_count_cyclist = if_else(is.na(ksi_id), 0, ksi_count_cyclist),
		ksi_count_other = if_else(is.na(ksi_id), 0, ksi_count_other)
	)

# Create combination of all id and time unit combos
if (tunit != "year") {
	
	# Create time unit combos for finer than a year
	tunit_combo <- expand_grid(
	    `_merge_group` = unique(bike_merge$`_merge_group`),
	    year = min(year(ksi_tunits$ksi_date)):max(year(ksi_tunits$ksi_date)),
	    time_value = unique(ksi_tunits[[glue("_ksi_time_{tunit}")]])
	) %>%
	    mutate(
	    	!!tunit_col := glue("{year}_{time_value}")
	    ) %>%
	    select(-year, -time_value)
	
} else {
	
	# Create time unit combos for year
	tunit_combo <- expand_grid(
	    `_merge_group` = unique(bike_merge$`_merge_group`),
	    !!tunit_col := min(year(ksi_tunits$ksi_date)):max(year(ksi_tunits$ksi_date))
	) %>%
		mutate(!!tunit_col := .data[[tunit_col]] %>% as.character)
	
}


# Fill in missing time units as zero
bike_ksi <- tunit_combo %>%
	left_join(
		bike_merge,
		by = c("_merge_group")
	) %>%
	left_join(
		bike_sjoin %>%
			as_tibble %>%
			select(c(
				"_merge_group",
				tunit_col,
				starts_with("ksi_count"),
				colnames(ksi_tunits)
			)) %>%
			select(-geometry),
		by = c("_merge_group", tunit_col)
	) %>%
	mutate(
		ksi_count = replace_na(ksi_count, 0),
		ksi_count_driver = replace_na(ksi_count_driver, 0),
		ksi_count_pedestrian = replace_na(ksi_count_pedestrian, 0),
		ksi_count_cyclist = replace_na(ksi_count_cyclist, 0),
		ksi_count_other = replace_na(ksi_count_other, 0)
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_ksi %>% nrow,
	"\n* Columns: ", bike_ksi %>% ncol,
	"\n\nNew Columns (", bike_ksi %>% ncol - bike_merge %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_ksi) %>%
			.[!. %in% colnames(bike_merge)],
		collapse = ""
	)
))

```

### Preview

```{r}

# Display merged bikeways with ksi counts
bike_ksi %>%
	filter(!is.na(ksi_id)) %>%
	head

```

## Prepare Model Data {.tabset}

**Note**: Improvements and installs/upgrades can occur at different times and parts of the same merged segment. Although there is an improvement or previous install or upgrade on a merged segment, it may not be present on the entire merged segment.

```{r}

# Extract year and time unit for time ids
if (tunit != "year") {
	
	# Time units more fine than a year
	tid_data <- bike_ksi %>%
		separate(
			glue("_ksi_time_{tunit}_group"),
			c("year", tunit),
			sep = "_"
		) %>%
		select(year, time_value)
	
} else {
	
	# Time unit as year
	tid_data <- bike_ksi %>%
		mutate(time_value = .data[[glue("_ksi_time_{tunit}")]]) %>%
		select(year = .data[[glue("_ksi_time_{tunit}")]], time_value)
	
}


# Assign time ids using ordered levels
if (tunit != "year") {
	
	# Get uniq combinations as ids
	tid <- expand_grid(
		time_year = min(tid_data$year, na.rm = T):max(tid_data$year, na.rm = T),
		time_value = unique(tid_data$time_value)
	) %>%
	    arrange(time_year, time_value) %>%
	    mutate(
	    	time_orig = glue("{time_year}_{time_value}"),
	    	time = factor(time_orig) %>% as.integer
	    )
	
} else {
	
	# Get year as ids
	tid <- data.frame(
		time_year = min(tid_data$year, na.rm = T):max(tid_data$year, na.rm = T)
	) %>%
		mutate(
			time_value = time_year,
			time_orig = time_value %>% as.character,
			time = factor(time_orig) %>% as.integer
		)
	
}

# Create time id map list
tid_map <- Map(reformulate, shQuote(tid$time), shQuote(tid$time_orig))

# Rowwise func for assigning improve or upgrade cohort
get_cohort <- function (
		change_time,
		change_type,
		time_value,
		type_value = NULL
	) {
	
	# Convert times to time ids
	change_tid <- tibble(
		time_orig = str_split(change_time, ",") %>% unlist,
		type = str_split(change_type, ",") %>% unlist
	) %>%
		mutate(
			time_value = case_match(time_orig, !!!tid_map)
		) %>%
		filter(
			type != "NA" &
			!is.na(type) &
			time_orig != "NA" &
			!is.na(time_orig) &
			!is.na(time_value)
		)
	
	# Include infra type only if avail
	if (!is.null(type_value)) {
		change_tid <- change_tid %>% filter(type %in% type_value)
	}
	
	# # Ignore if no data exists
	if (nrow(change_tid) < 1) {
		return(NA)
	}
	
	# Get time ids for changes only
	change_tid <- change_tid %>%
		pull(time_value) %>%
		as.integer
	
	# Get cohort for improve
	time_value <- time_value %>% as.integer
	if (time_value %in% change_tid) { # time value is same as change time
		out <- time_value
	} else if (any(change_tid < time_value)) { # time value lt change value then use latest
		out <- change_tid[change_tid < time_value] %>% max(na.rm = T)
	} else { # time value more than improve time
		out <- NA
	}
	return(out)
	
}

# Prepare model data
model_data <- bike_ksi %>%
	mutate( # Add whether segments were improved or not
		verify_improve1 = if_else(str_detect(`_improve1`, "TRUE"), T, F),
		verify_improve1_type = if_else(verify_improve1, verify_upgrade1_type, NA),
		verify_improve1_year = if_else(verify_improve1, verify_upgrade1_year, NA),
		verify_improve1_time_orig = if_else(verify_improve1, .data[[glue("_upgrade1_time_{tunit}_group")]], NA),
		verify_improve2 = if_else(str_detect(`_improve2`, "TRUE"), T, F),
		verify_improve2_type = if_else(verify_improve2, verify_upgrade2_type, NA),
		verify_improve2_year = if_else(verify_improve2, verify_upgrade2_year, NA),
		verify_improve2_time_orig = if_else(verify_improve2, .data[[glue("_upgrade2_time_{tunit}_group")]], NA),
		verify_improve = if_else(str_detect(`_improve`, "TRUE"), T, F)
	) %>%
	select(
		bike_merge_id = `_merge_group`,
		bike_city = city,
		bike_status = status,
		bike_type = type,
		bike_len_meters = geometry_len_meters,
		bike_time_orig = time_group,
		bike_year = year,
		bike_time = time_value,
		bike_install_type = verify_install_type,
		bike_install_year = verify_install_year,
		bike_install_time_orig = .data[[glue("_install_time_{tunit}_group")]],
		bike_upgrade1_type = verify_upgrade1_type,
		bike_upgrade1_year = verify_upgrade1_year,
		bike_upgrade1_time_orig = .data[[glue("_upgrade1_time_{tunit}_group")]],
		bike_upgrade2_type = verify_upgrade2_type,
		bike_upgrade2_year = verify_upgrade2_year,
		bike_upgrade2_time_orig = .data[[glue("_upgrade2_time_{tunit}_group")]],
		bike_improve = verify_improve,
		bike_improve1 = verify_improve1,
		bike_improve1_type = verify_improve1_type,
		bike_improve1_year = verify_improve1_year,
		bike_improve1_time_orig = verify_improve1_time_orig,
		bike_improve2 = verify_improve2,
		bike_improve2_type = verify_improve2_type,
		bike_improve2_year = verify_improve2_year,
		bike_improve2_time_orig = verify_improve2_time_orig,
		ksi_time_orig = .data[[glue("_ksi_time_{tunit}_group")]],
		ksi = ksi_count,
		ksi_driver = ksi_count_driver,
		ksi_pedestrian = ksi_count_pedestrian,
		ksi_cyclist = ksi_count_cyclist,
		ksi_other = ksi_count_other
	) %>%
	left_join( # add data about range of data
		ksi_y2009to2023 %>%
			as_tibble() %>%
			group_by(ksi_city) %>%
			summarize(
				ksi_year_min = min(year(ksi_date)),
				ksi_year_max = max(year(ksi_date))
			),
		by = join_by(bike_city == ksi_city)
	) %>%
	separate( # get year and time value
		ksi_time_orig,
		c("ksi_year", "ksi_time"),
		sep = "_",
		remove = F
	) %>%
	mutate( # set ksi count to na if not within data year range
		ksi = if_else(
			ksi_year < ksi_year_min | ksi_year > ksi_year_max,
			NA,
			ksi
		),
		ksi_driver = if_else(is.na(ksi), NA, ksi_driver),
		ksi_pedestrian = if_else(is.na(ksi), NA, ksi_pedestrian),
		ksi_cyclist = if_else(is.na(ksi), NA, ksi_cyclist),
		ksi_other = if_else(is.na(ksi), NA, ksi_other)
	) %>%
	mutate( # fill in as year if unit is year
		ksi_time = case_when(
			tunit == "year" ~ ksi_year,
			.default = ksi_time
		),
		ksi_time_unit = tunit
	) %>%
	mutate( # assign ids and add time units
		id = factor(bike_merge_id) %>% as.integer,
		bike_time_unit = tunit,
		ksi_time_unit = tunit
	) %>%
	left_join( # assign ksi time ids
		tid %>%
			select(
				ksi_time_orig = time_orig,
				time
			),
		by = "ksi_time_orig"
	) %>%
	left_join( # assign cohort based on bike time ids
		tid %>%
			select(
				bike_time_orig = time_orig,
				cohort = time
			),
		by = "bike_time_orig"
	) %>%
	rowwise %>%
	mutate( # assign cohort
		cohort_pl = get_cohort(
			paste(
				bike_install_time_orig,
				bike_upgrade1_time_orig,
				bike_upgrade2_time_orig,
				sep = ","
			),
			paste(
				bike_install_type,
				bike_upgrade1_type,
				bike_upgrade2_type,
				sep = ","
			),
			time,
			c("PL", "BUF")
		),
		cohort_ct = get_cohort(
			paste(
				bike_install_time_orig,
				bike_upgrade1_time_orig,
				bike_upgrade2_time_orig,
				sep = ","
			),
			paste(
				bike_install_type,
				bike_upgrade1_type,
				bike_upgrade2_type,
				sep = ","
			),
			time,
			"PBL"
		),
		cohort_install = get_cohort(
			bike_install_time_orig,
			bike_install_type,
			time
		),
		cohort_install_pl = get_cohort(
			bike_install_time_orig,
			bike_install_type,
			time,
			c("PL", "BUF")
		),
		cohort_install_ct = get_cohort(
			bike_install_time_orig,
			bike_install_type,
			time,
			"PBL"
		),
		cohort_upgrade1 = get_cohort(
			bike_upgrade1_time_orig,
			bike_upgrade1_type,
			time
		),
		cohort_upgrade1_pl = get_cohort(
			bike_upgrade1_time_orig,
			bike_upgrade1_type,
			time,
			c("PL", "BUF")
		),
		cohort_upgrade1_ct = get_cohort(
			bike_upgrade1_time_orig,
			bike_upgrade1_type,
			time,
			"PBL"
		),
		cohort_upgrade2 = get_cohort(
			bike_upgrade2_time_orig,
			bike_upgrade2_type,
			time
		),
		cohort_upgrade2_pl = get_cohort(
			bike_upgrade2_time_orig,
			bike_upgrade2_type,
			time,
			c("PL", "BUF")
		),
		cohort_upgrade2_ct = get_cohort(
			bike_upgrade2_time_orig,
			bike_upgrade2_type,
			time,
			"PBL"
		),
		cohort_improve1 = get_cohort(
			bike_improve1_time_orig,
			bike_improve1_type,
			time
		),
		cohort_improve1_pl = get_cohort(
			bike_improve1_time_orig,
			bike_improve1_type,
			time,
			c("PL", "BUF")
		),
		cohort_improve1_ct = get_cohort(
			bike_improve1_time_orig,
			bike_improve1_type,
			time,
			"PBL"
		),
		cohort_improve2 = get_cohort(
			bike_improve2_time_orig,
			bike_improve1_type,
			time
		),
		cohort_improve2_pl = get_cohort(
			bike_improve2_time_orig,
			bike_improve2_type,
			time,
			c("PL", "BUF")
		),
		cohort_improve2_ct = get_cohort(
			bike_improve2_time_orig,
			bike_improve2_type,
			time,
			"PBL"
		)
	) %>%
	mutate( # assign post based on whether it is greater than bike time
		post = as.integer(time >= cohort),
		post_install = as.integer(time >= cohort_install) %>% replace_na(0),
		post_install_pl = as.integer(time >= cohort_install_pl) %>% replace_na(0),
		post_install_ct = as.integer(time >= cohort_install_ct) %>% replace_na(0),
		post_upgrade1 = as.integer(time >= cohort_upgrade1) %>% replace_na(0),
		post_upgrade1_pl = as.integer(time >= cohort_upgrade1_pl) %>% replace_na(0),
		post_upgrade1_ct = as.integer(time >= cohort_upgrade1_ct) %>% replace_na(0),
		post_upgrade2 = as.integer(time >= cohort_upgrade2) %>% replace_na(0),
		post_upgrade2_pl = as.integer(time >= cohort_upgrade2_pl) %>% replace_na(0),
		post_upgrade2_ct = as.integer(time >= cohort_upgrade2_ct) %>% replace_na(0),
		post_improve1 = as.integer(time >= cohort_improve1) %>% replace_na(0),
		post_improve1_pl = as.integer(time >= cohort_improve1_pl) %>% replace_na(0),
		post_improve1_ct = as.integer(time >= cohort_improve1_ct) %>% replace_na(0),
		post_improve2 = as.integer(time >= cohort_improve2) %>% replace_na(0),
		post_improve2_pl = as.integer(time >= cohort_improve2_pl) %>% replace_na(0),
		post_improve2_ct = as.integer(time >= cohort_improve2_ct) %>% replace_na(0)
	) %>%
	mutate( # add length in km
		bike_len_km = bike_len_meters / 1000,
		.after = bike_len_meters
	) %>%
	arrange(id, time) %>%
	select(
		starts_with("bike_"),
		starts_with("ksi"),
		id,
		time,
		starts_with("post"),
		starts_with("cohort"),
		everything()
	)

# Export model data and time ids to folder
model_data %>% write_csv("../data/recovr-eval-2024-10-17.csv", na = "")
```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_ksi %>% nrow,
	"\n* Columns: ", bike_ksi %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", model_data %>% nrow,
	"\n* Columns: ", model_data %>% ncol,
	"\n\nNew Columns (", model_data %>% ncol - bike_ksi %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(model_data) %>%
			.[!. %in% colnames(bike_ksi)],
		collapse = ""
	)
))

```

### Preview

```{r}

# Display model data
model_data %>%
	head

```

## Flow Chart

```{r}

# Func to calc km
calc_total_km <- function(df, places = 0) {
	out <- round(
		df %>%
			st_length %>%
			as.numeric %>%
			sum(., na.rm = T) / 1000,
		places
	)
	return(out)
}

# Calculate vancbike vars
vancbike_stats <- list(
	raw = vancbike_raw %>% nrow,
	verify = bike_verify %>% filter(city == "vancouver") %>% nrow,
	itypes = bike_itypes %>% filter(city == "vancouver") %>% nrow,
	p2010 = bike_p2010 %>% filter(city == "vancouver") %>% nrow,
	tfilter = bike_tfilter %>% filter(city == "vancouver") %>% nrow,
	merge = bike_merge %>% filter(city == "vancouver") %>% nrow,
	sjoin = sjoin %>% filter(ksi_city == "vancouver" & !is.na(ksi_id)) %>% pull(ksi_id) %>% unique %>% length,
	raw_km = calc_total_km(vancbike_raw),
	verify_km = calc_total_km(bike_verify %>% filter(city == "vancouver")),
	itypes_km = calc_total_km(bike_itypes %>% filter(city == "vancouver")),
	p2010_km = calc_total_km(bike_p2010 %>% filter(city == "vancouver")),
	tfilter_km = calc_total_km(bike_tfilter %>% filter(city == "vancouver")),
	merge_km = calc_total_km(bike_merge %>% filter(city == "vancouver")),
	years = vancbike_raw$install_year %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Calculate vancksi vars
vancksi_stats <- list(
	raw = ksi_combine %>% filter(ksi_city == "vancouver") %>% nrow %>% format(big.mark = ","),
	coords = ksi_coords %>% filter(ksi_city == "vancouver") %>% nrow %>% format(big.mark = ","),
	injury = ksi_injury %>% filter(ksi_city == "vancouver") %>% nrow %>% format(big.mark = ","),
	y2009to2023 = ksi_y2009to2023 %>% filter(ksi_city == "vancouver") %>% nrow %>% format(big.mark = ","),
	years = vancksi %>% mutate(x = year(ksi_date)) %>% pull(x) %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Calculate calgbike vars
calgbike_stats <- list(
	raw = calgbike_raw %>% nrow,
	verify = bike_verify %>% filter(city == "calgary") %>% nrow,
	itypes = bike_itypes %>% filter(city == "calgary") %>% nrow,
	p2010 = bike_p2010 %>% filter(city == "calgary") %>% nrow,
	tfilter = bike_tfilter %>% filter(city == "calgary") %>% nrow,
	merge = bike_merge %>% filter(city == "calgary") %>% nrow,
	sjoin = sjoin %>% filter(ksi_city == "calgary" & !is.na(ksi_id)) %>% pull(ksi_id) %>% unique %>% length,
	raw_km = calc_total_km(calgbike_raw),
	verify_km = calc_total_km(bike_verify %>% filter(city == "calgary")),
	itypes_km = calc_total_km(bike_itypes %>% filter(city == "calgary")),
	p2010_km = calc_total_km(bike_p2010 %>% filter(city == "calgary")),
	tfilter_km = calc_total_km(bike_tfilter %>% filter(city == "calgary")),
	merge_km = calc_total_km(bike_merge %>% filter(city == "calgary")),
	years = calgbike_raw$install_year %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Calculate calgksi vars
calgksi_stats <- list(
	raw = ksi_combine %>% filter(ksi_city == "calgary") %>% nrow %>% format(big.mark = ","),
	coords = ksi_coords %>% filter(ksi_city == "calgary") %>% nrow %>% format(big.mark = ","),
	injury = ksi_injury %>% filter(ksi_city == "calgary") %>% nrow %>% format(big.mark = ","),
	y2009to2023 = ksi_y2009to2023 %>% filter(ksi_city == "calgary") %>% nrow %>% format(big.mark = ","),
	years = calgksi %>% mutate(x = year(ksi_date)) %>% pull(x) %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Calculate toronbike n vars
toronbike_stats <- list(
	raw = toronbike_raw %>% nrow,
	verify = bike_verify %>% filter(city == "toronto") %>% nrow,
	itypes = bike_itypes %>% filter(city == "toronto") %>% nrow,
	p2010 = bike_p2010 %>% filter(city == "toronto") %>% nrow,
	tfilter = bike_tfilter %>% filter(city == "toronto") %>% nrow,
	merge = bike_merge %>% filter(city == "toronto") %>% nrow,
	sjoin = sjoin %>% filter(ksi_city == "toronto" & !is.na(ksi_id)) %>% pull(ksi_id) %>% unique %>% length,
	raw_km = calc_total_km(toronbike_raw),
	verify_km = calc_total_km(bike_verify %>% filter(city == "toronto")),
	itypes_km = calc_total_km(bike_itypes %>% filter(city == "toronto")),
	p2010_km = calc_total_km(bike_p2010 %>% filter(city == "toronto")),
	tfilter_km = calc_total_km(bike_tfilter %>% filter(city == "toronto")),
	merge_km = calc_total_km(bike_merge %>% filter(city == "toronto")),
	years = toronbike_raw$install_year %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Calculate toronksi n vars
toronksi_stats <- list(
	raw = ksi_combine %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ","),
	coords = ksi_coords %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ","),
	injury = ksi_injury %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ","),
	y2009to2023 = ksi_y2009to2023 %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ","),
	years = toronksi %>% mutate(x = year(ksi_date)) %>% pull(x) %>% range(na.rm = T) %>% paste0(., collapse = "-")
)

# Other vars for flow chart
tunit_title <- tunit %>% str_to_title

# Create flow chart
fig_flowchart <- grViz(glue('

digraph {
	
	// Settings
	layout=neato
	splines=ortho
	node [shape="box", width=2.75, height=1.75, margin=0.1]
	
	// Raw bikeways
	bike_raw[
		width=2,
		shape=cylinder,
		label=<
			<br/><b>Bikeways*</b><br/><br/>
			Vancouver: {{vancbike_stats$raw_km}} km (n={{vancbike_stats$raw}}) <br/>
			Calgary: {{calgbike_stats$raw_km}} km (n={{calgbike_stats$raw}})<br/>
			Toronto: {{toronbike_stats$raw_km}} km (n={{toronbike_stats$raw}})
		>,
		pos="0,0!"
	]
	
	// Verified bikeways
	bike_verify[
		label=<
			<b>Verified Bikeways Only</b><br/><br/>
			Vancouver: {{vancbike_stats$verify_km}} km (n={{vancbike_stats$verify}}) <br/>
			Calgary: {{calgbike_stats$verify_km}} km (n={{calgbike_stats$verify}})<br/>
			Toronto: {{toronbike_stats$verify_km}} km (n={{toronbike_stats$verify}})
		>,
		pos="3,0!"
	]
	
	// PL and CT bikeways
	bike_itypes[
		label=<
			<b>Painted Lanes and<br/>
			Cycle Tracks Only</b><br/><br/>
			Vancouver: {{vancbike_stats$itypes_km}} km (n={{vancbike_stats$itypes}}) <br/>
			Calgary: {{calgbike_stats$itypes_km}} km (n={{calgbike_stats$itypes}})<br/>
			Toronto: {{toronbike_stats$itypes_km}} km (n={{toronbike_stats$itypes}})
		>,
		pos="6,0!"
	]
	
	// Post-2010 bikeways
	bike_p2010[
		label=<
			<b>Post-2010 Installs<br/>
			or Upgrades Only</b><br/><br/>
			Vancouver: {{vancbike_stats$p2010_km}} km (n={{vancbike_stats$p2010}}) <br/>
			Calgary: {{calgbike_stats$p2010_km}} km (n={{calgbike_stats$p2010}})<br/>
			Toronto: {{toronbike_stats$p2010_km}} km (n={{toronbike_stats$p2010}})
		>,
		pos="9,0!"
	]
	
	// Within time unit
	bike_tfilter[
		label=<
			<b>Installs or Upgrades<br/>
			with {{tunit_title}}s</b><br/><br/>
			Vancouver: {{vancbike_stats$tfilter_km}} km (n={{vancbike_stats$tfilter}}) <br/>
			Calgary: {{calgbike_stats$tfilter_km}} km (n={{calgbike_stats$tfilter}})<br/>
			Toronto: {{toronbike_stats$tfilter_km}} km (n={{toronbike_stats$tfilter}})
		>,
		pos="3,-2.25!"
	]
	
	// Spatial merge
	bike_merge[
		label=<
			<b>Spatially Merged Bikeways<br/>
			by Status, Type, and {{tunit_title}}</b><br/><br/>
			Vancouver: {{vancbike_stats$merge_km}} km (n={{vancbike_stats$merge}}) <br/>
			Calgary: {{calgbike_stats$merge_km}} km (n={{calgbike_stats$merge}})<br/>
			Toronto: {{toronbike_stats$merge_km}} km (n={{toronbike_stats$merge}})
		>,
		pos="6,-2.25!"
	]
	
	// Raw KSI
	ksi_raw[
		width=2,
		shape=cylinder,
		label=<
			<br/><b>KSI**</b><br/><br/>
			Vancouver: {{vancksi_stats$raw}} KSI<br/>
			Calgary: {{calgksi_stats$raw}} KSI<br/>
			Toronto: {{toronksi_stats$raw}} KSI
		>,
		pos="0,-4.5!"
	]
	
	// KSI with coordinates
	ksi_coords[
		label=<
			<b>With Coordinates</b><br/><br/>
			Vancouver: {{vancksi_stats$coords}} KSI<br/>
			Calgary: {{calgksi_stats$coords}} KSI<br/>
			Toronto: {{toronksi_stats$coords}} KSI
		>,
		pos="3,-4.5!"
	]
	
	// KSI with fatal or serious injury only
	ksi_injury[
		label=<
			<b>Fatal or Serious<br/>Injuries Only</b><br/><br/>
			Vancouver: {{vancksi_stats$injury}} KSI<br/>
			Calgary: {{calgksi_stats$injury}} KSI<br/>
			Toronto: {{toronksi_stats$injury}} KSI
		>,
		pos="6,-4.5!"
	]
	
	// Post-2010 KSI
	ksi_y2009to2023[
		label=<
			<b>From 2009 to 2023</b><br/><br/>
			Vancouver: {{vancksi_stats$y2009to2023}} KSI<br/>
			Calgary: {{calgksi_stats$y2009to2023}} KSI<br/>
			Toronto: {{toronksi_stats$y2009to2023}} KSI
		>,
		pos="9,-4.5!"
	]
	
	// Spatial Joined KSI
	ksi_sjoin[
		label=<
			<b>KSI Within 25 Meters of<br/>
			Merged Bikeways</b><br/><br/>
			Vancouver: {{vancbike_stats$sjoin}} KSI (n={{vancbike_stats$merge}}) <br/>
			Calgary: {{calgbike_stats$sjoin}} KSI (n={{calgbike_stats$merge}})<br/>
			Toronto: {{toronbike_stats$sjoin}} KSI (n={{toronbike_stats$merge}})
		>,
		pos="9,-2.25!"
	]
	
	// Footer
	foot[
		shape="none",
		fontsize="10",
		height=0.65,
		width=1,
		labeljust="r",
		label="*Bikeway Data Year Ranges: Vancouver ({{vancbike_stats$years}}), Calgary ({{calgbike_stats$years}}), Toronto ({{toronbike_stats$years}})\n**KSI Data Year Ranges: Vancouver ({{vancksi_stats$years}}), Calgary ({{calgksi_stats$years}}), Toronto ({{toronksi_stats$years}})",
		pos="5,-5.85!"
	]
	
	
	
	// Bikeway process
	bike_raw -> bike_verify -> bike_itypes -> bike_p2010 -> bike_tfilter -> bike_merge -> ksi_sjoin
	
	// KSI process
	ksi_raw -> ksi_coords -> ksi_injury -> ksi_y2009to2023 -> ksi_sjoin
}

', .open = "{{", .close = "}}"))

# Export flowchart
fig_flowchart %>%
    export_svg %>%
    charToRaw %>%
    rsvg_png(
    	"../figures/fig-methods.png",
    	width = 1920
    )

# Display flowchart
fig_flowchart

```
