---
title: "Cycling Interventions Evaluation"
subtitle: "R Code"
author:
- "Richard Wen richard.wen@utoronto.ca"
- "Brice Batomen brice.kuimi@utoronto.ca"
- "Linda Rothman linda.rothman@torontomu.ca"
- "Andrew Howard andrew.howard@sickkids.ca"
date: "`r format(Sys.time(), '%B %d, %Y')`"
knit: |
    (function(input_rmd, ...) {
    rmarkdown::render(
        input_rmd,
        rmarkdown::html_document(
            toc = TRUE,
            toc_float = TRUE,
            highlight = "zenburn",
            code_folding = "hide",
            df_print = "paged",
            self_contained = FALSE
        ),
        output_dir = "../docs",
        output_file = "index", ...)
    })
---

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	message = FALSE
)
```

# Installation

1.  Install [R](https://www.r-project.org/)
2.  Install [RTools](https://cran.r-project.org/) if you are on Windows
3.  Install [RStudio](https://posit.co/download/rstudio-desktop/)

For more details, see [Software and Package Versions](#software-and-package-versions).

# Running This Code

1.  Ensure the installation steps above are completed
2.  Download a zip of the code and data [here](https://github.com/rrwen/recovr-eval/archive/refs/heads/main.zip) and unzip it
    - Code Repository: [github.com/rrwen/recovr-eval](https://github.com/rrwen/recovr-eval)
3.  In RStudio, open the [src/src.Rproj](https://github.com/rrwen/recovr-eval/blob/main/src/src.Rproj) file
4.  Then, open the [src/index.Rmd](https://github.com/rrwen/recovr-eval/blob/main/src/index.Rmd) file
5.  In RStudio:
    - Run all code: Click the `Run` drop down (top right of the code pane) and click `Run All`
    - Generate HTML version: Click `knit` (top left of code pane) and a file will be generated in `docs/index.html`

# Libraries

Install R packages if needed.

```{r, results = FALSE}

# Required packages
required_packages <- c(
	"rmarkdown",
	"bookdown",
	"knitr",
	"tidyverse",
	"purrr",
	"glue",
	"lubridate",
	"scales",
	"patchwork",
	"DiagrammeR",
	"DiagrammeRsvg",
	"webshot2",
	"magick",
	"rsvg",
	"sf",
	"tmap",
	"ggspatial",
	"prettymapr",
	"units",
	"leaflet",
	"leaflet.extras",
	"igraph"
)

# Try to install packages if not installed
default_options <- options()
tryCatch(
	{
		# Disable interactivity
		options(install.packages.compile.from.source = "always")
		
		# Install package if not installed
		for (package in required_packages) {
			is_package_installed <- require(package, character.only = TRUE)
			if (!is_package_installed) {
				cat(paste0("Installing package: ", package, "\n"))
				install.packages(package)
			} else {
				cat(paste0("Package already installed: ", package, "\n"))
			}
		}
	},
	error = function(cond) {
		stop(cond)
	},
	finally = {
		options(default_options) # reset interactivity
	}
)
```

Load R libraries.

```{r}
library(DiagrammeR)
library(ggplot2)
library(ggspatial)
library(glue)
library(igraph)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(patchwork)
library(sf)
library(tidyverse)
library(tmap)
```

# Data

Read data from the `data` folder.

```{r}
ddesc <- read_csv("../data/data.csv")
ddesc
```

## Vancouver Bikeways {.tabset}

`r ddesc %>% filter(file == "vancouver-bikeways-2024-06-02.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "vancouver-bikeways-2024-06-02.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "vancouver-bikeways-2024-06-02.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "vancouver-bikeways-2024-06-02.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "vancouver-bikeways-2024-06-02.geojson") %>% pull(notes)`

```{r}

# Read data
vancbike_raw <- read_sf("../data/vancouver-bikeways-2024-06-02.geojson")

# Get download date
vancbike_dldate <- ddesc %>% filter(
	file == "vancouver-bikeways-2024-06-02.geojson"
) %>% pull(download_date)
```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(vancbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(vancbike_raw)`
* Rows: `r nrow(vancbike_raw)`

```{r}
vancbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
vancbike_ddict <- read_csv("../data/vancouver-bikeways-2024-06-02-datadict.csv")
vancbike_ddict
```

### Details

```{r}
print(vancbike_raw)
```

### Files

The data files are available below:

- [vancouver-bikeways-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/vancouver-bikeways-2024-06-02.geojson)
- [vancouver-bikeways-2024-06-02-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/vancouver-bikeways-2024-06-02-datadict.csv)

## Calgary Bikeways {.tabset}

`r ddesc %>% filter(file == "calgary-bikeways-2024-06-05.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "calgary-bikeways-2024-06-05.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "calgary-bikeways-2024-06-05.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "calgary-bikeways-2024-06-05.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "calgary-bikeways-2024-06-05.geojson") %>% pull(notes)`

```{r}

# Read data
calgbike_raw <- read_sf("../data/calgary-bikeways-2024-06-05.geojson")

# Get download date
calgbike_dldate <- ddesc %>% filter(
	file == "calgary-bikeways-2024-06-05.geojson"
) %>% pull(download_date)
```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(calgbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(calgbike_raw)`
* Rows: `r nrow(calgbike_raw)`

```{r}
calgbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
calgbike_ddict <- read_csv("../data/calgary-bikeways-2024-06-05-datadict.csv")
calgbike_ddict
```

### Details

```{r}
print(calgbike_raw)
```

### Files

The data files are available below:

- [calgary-bikeways-2024-06-05.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/calgary-bikeways-2024-06-05.geojson)
- [calgary-bikeways-2024-06-05-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/calgary-bikeways-2024-06-05-datadict.csv)

## Toronto Bikeways {.tabset}

`r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "toronto-bikeways-2024-06-02.geojson") %>% pull(notes)`

```{r}

# Read data
toronbike_raw <- read_sf("../data/toronto-bikeways-2024-06-02.geojson")

# Get download date
toronbike_dldate <- ddesc %>% filter(
	file == "toronto-bikeways-2024-06-02.geojson"
) %>% pull(download_date)

```

### Map

Only the first 1000 records are shown.

```{r}
tmap_mode("view")
tm_shape(toronbike_raw %>% head(1000)) +
    tm_lines(
    	col = "#336699",
    	border.col = "white",
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(toronbike_raw)`
* Rows: `r nrow(toronbike_raw)`

```{r}
toronbike_raw %>% as_tibble
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
toronbike_ddict <- read_csv("../data/toronto-bikeways-2024-06-02-datadict.csv")
toronbike_ddict
```

### Details

```{r}
print(toronbike_raw)
```

### Files

The data files are available below:

- [toronto-bikeways-2024-06-02.geojson](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-06-02.geojson)
- [toronto-bikeways-2024-06-02-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-bikeways-2024-06-02-datadict.csv)

## Verified Dates {.tabset}

`r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "verify-dates-2024-06-12.csv") %>% pull(notes)`

```{r}

# Read data
vdates_raw <- read_csv("../data/verify-dates-2024-06-12.csv")

# Get download date
vdates_dldate <- ddesc %>% filter(
	file == "verify-dates-2024-06-12.csv"
) %>% pull(download_date)
```

### Data

* Columns: `r ncol(vdates_raw)`
* Rows: `r nrow(vdates_raw)`

```{r}
vdates_raw
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
vdates_ddict <- read_csv("../data/verify-dates-2024-06-12-datadict.csv")
vdates_ddict
```

### Files

The data files are available below:

- [verify-dates-2024-06-12.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12.csv)
- [verify-dates-2024-06-12-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/verify-dates-2024-06-12-datadict.csv)

## Toronto KSI {.tabset}

`r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(description)`

* **Download Link**: `r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(url)`
* **Download Date**: `r format(ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(download_date), '%B %d, %Y')`
* **Data Updated**: `r format(ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(update_date), '%B %d, %Y')`
* **Notes**: `r ddesc %>% filter(file == "toronto-ksi-2024-06-13.csv") %>% pull(notes)`

```{r}

# Read data
toronksi_raw <- read_sf(
	"../data/toronto-ksi-2024-06-13.csv",
	options = c(
		"X_POSSIBLE_NAMES=LONGITUDE",
		"Y_POSSIBLE_NAMES=LATITUDE"
	),
	crs = 4326
)

# Get download date
toronksi_dldate <- ddesc %>% filter(
	file == "toronto-ksi-2024-06-13.csv"
) %>% pull(download_date)
```

### Map

**Note**: Due to the large number of records, only the latest year of `r year(max(toronksi_raw$DATE))` is displayed (n = `r toronksi_raw %>% filter(year(DATE) == max(year(DATE))) %>% nrow`).

```{r}
tmap_mode("view")
tm_shape(toronksi_raw %>% filter(year(DATE) == max(year(DATE)))) +
    tm_dots(
    	col = "ACCLASS",
    	clustering = TRUE,
    	popup.vars = TRUE
    )
```

### Data

* Columns: `r ncol(toronksi_raw)`
* Rows: `r nrow(toronksi_raw)`

```{r}
toronksi_raw %>% as_tibble()
```

### Dictionary

The data contains the following columns:

```{r, cols.print = 3}
ksi_ddict <- read_csv("../data/toronto-ksi-2024-06-13-datadict.csv")
ksi_ddict
```

### Details

```{r}
print(toronksi_raw)
```

### Files

The data files are available below:

- [toronto-ksi-2024-06-13.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-ksi-2024-06-13.csv)
- [toronto-ksi-2024-06-13-datadict.csv](https://github.com/rrwen/recovr-eval/blob/main/data/toronto-ksi-2024-06-13-datadict.csv)

# Cleaning

## Combine Bikeways {.tabset}

Combine bikeway data across all cities using common columns.

**Note**: For Calgary and Toronto, there weren't any included bikeways that were not verified (e.g. Local Street Bikeways (LSBs)), and a dummy column was added to account for this case.

```{r}

# List of city bikeway data
bike_list <- list(
	vancouver = vancbike_raw,
	calgary = calgbike_raw %>%
		mutate(no_verify_install_type = NA),
	toronto = toronbike_raw %>%
		mutate(no_verify_install_type = NA)
)

# Get common columns across all city bikeways
bike_cols <- bike_list %>%
	map(colnames) %>%
	reduce(intersect)

# Combine bikeway data across cities
bike_combine <- names(bike_list) %>%
	map(function(city) {
		bike_list[[city]] %>%
			select(
				all_of(bike_cols)
			) %>%
			mutate(
				city = factor(city, levels = names(bike_list)),
				.before = 1
			)
	}) %>%
	reduce(add_row)

```

### Details

```{r}

# Display combined bikeway rows and cols
cat(glue(
	"\n* Rows: ", bike_combine %>% nrow,
	"\n* Columns: ", bike_combine %>% ncol
))

# Display rows per city
bike_combine_counts <- bike_combine %>%
	as_tibble %>%
	group_by(city) %>% count
bike_combine_counts

```

### Preview

```{r}

# Display combined bikeway data
bike_combine %>% as_tibble %>% head

```

## Combine KSI

Combine KSI data across cities using common columns.

**Note**: Only Toronto KSI was included currently.

```{r}

# Combine KSI data
ksi_combine <- toronksi_raw %>%
	mutate(
		ksi_city = "toronto",
		ksi_id = INDEX,
		ksi_date = DATE
	)

```

### Details

```{r}

# Display combined ksi rows and cols
cat(glue(
	"\n* Rows: ", ksi_combine %>% nrow,
	"\n* Columns: ", ksi_combine %>% ncol
))

# Display rows per city
ksi_combine_counts <- ksi_combine %>%
	as_tibble %>%
	group_by(ksi_city) %>%
	count
ksi_combine_counts

```

### Preview

```{r}

# Display combined ksi data
ksi_combine %>% as_tibble %>% head

```

## Filter Verified Bikeways

Filter for bikeways with a verified installations or upgrades.

```{r}

# Filter bikeways for post2011
bike_verify <- bike_combine %>%
	filter(
		!is.na(verify_install_year) |
		!is.na(verify_upgrade1_year) |
		!is.na(verify_upgrade2_year) &
		is.na(no_verify_install_type)
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_combine %>% nrow,
	"\n* Columns: ", bike_combine %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_verify %>% nrow,
	"\n* Columns: ", bike_verify %>% ncol
))

# Display counts
bike_verify_counts <- bike_verify %>%
	as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(status, value) %>%
	count
bike_verify_counts

```

#### Preview

```{r}

# Display filtered rows
bike_verify %>%
	as_tibble %>%
	select(
		verify_install_year,
		verify_upgrade1_year,
		verify_upgrade2_year,
		no_verify_install_type,
		everything()
	) %>%
	arrange(`verify_install_year`) %>%
	head

```

## Filter Infrastructure Types

Filter for painted lanes (`PL` and `BUF`) and cycle tracks (`PBL`) only in any of `verify_install_type`, `verify_upgrade1_type`, and `verify_upgrade2_type`.

```{r}

# Remove none or na bike types
bike_itypes <- bike_verify %>%
	filter(
		verify_install_type %in% c("PL", "BUF", "PBL") |
		verify_upgrade1_type %in% c("PL", "BUF", "PBL") |
		verify_upgrade2_type %in% c("PL", "BUF", "PBL")
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_verify %>% nrow,
	"\n* Columns: ", bike_verify %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_itypes %>% nrow,
	"\n* Columns: ", bike_itypes %>% ncol
))

# Display counts
bike_itypes_counts <- bike_itypes %>%
	as_tibble %>%
	filter(
		verify_install_type %in% c("PL", "BUF", "PBL")
	) %>%
	mutate(
		type = verify_install_type
	) %>%
	group_by(type) %>%
	count %>%
	ungroup %>%
	add_row(
		bike_itypes %>%
			as_tibble %>%
			filter(
				!verify_install_type %in% c("PL", "BUF", "PBL") &
				verify_upgrade1_type %in% c("PL", "BUF", "PBL")
			) %>%
			mutate(
				type = glue("{verify_install_type} (Install) to {verify_upgrade1_type} (1st Upgrade)")
			) %>%
			group_by(type) %>%
			count
	) %>%
	add_row(
		bike_itypes %>%
			as_tibble %>%
			filter(
				!verify_install_type %in% c("PL", "BUF", "PBL") &
				!verify_upgrade1_type %in% c("PL", "BUF", "PBL") &
				verify_upgrade2_type %in% c("PL", "BUF", "PBL")
			) %>%
			mutate(
				type = glue("{verify_install_type} to {verify_upgrade1_type} (1st Upgrade) to {verify_upgrade2_type} (2nd Upgrade)")
			) %>%
			group_by(type) %>%
			count
	)
bike_itypes_counts

```

### Preview

```{r}

# Preview filtered bike types data
bike_itypes %>% as_tibble %>% head

```

## Filter Post-2011 Bikeways

Filter for bikeways with a verified installations or upgrades after 2011.

```{r}

# Filter bikeways for post2011
bike_p2011 <- bike_itypes %>%
	filter(
		verify_install_year > 2011 |
		verify_upgrade1_year > 2011 |
		verify_upgrade2_year > 2011
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_itypes %>% nrow,
	"\n* Columns: ", bike_itypes %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_p2011 %>% nrow,
	"\n* Columns: ", bike_p2011 %>% ncol
))

# Display counts
bike_p2011_counts <- bike_p2011 %>%
	as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(status, value) %>%
	count
bike_p2011_counts

```

#### Preview

```{r}

# Display filtered rows
bike_p2011 %>%
	as_tibble %>%
	select(
		verify_install_year,
		verify_upgrade1_year,
		verify_upgrade2_year,
		everything()
	) %>%
	arrange(`verify_install_year`) %>%
	head

```

## Filter 2009-2023 KSI

Filter for KSI records between 2009 and 2023 inclusively.

```{r}

# Filter ksi for post2011
ksi_y2009to2023 <- ksi_combine %>%
	filter(year(ksi_date) >= 2009 & year(ksi_date) <= 2023)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_combine %>% nrow,
	"\n* Columns: ", ksi_combine %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_y2009to2023 %>% nrow,
	"\n* Columns: ", ksi_y2009to2023 %>% ncol
))

# Display counts
ksi_y2009to2023_counts <- ksi_y2009to2023 %>%
	as_tibble %>%
	group_by(year(ksi_date)) %>%
	count
ksi_y2009to2023_counts

```

#### Preview

```{r}

# Display filtered rows
ksi_y2009to2023 %>%
	as_tibble %>%
	select(
		ksi_date,
		everything()
	) %>%
	arrange(year(ksi_date)) %>%
	head

```

## Add Bikeway Dates

Join ambiguous verified install/upgrade dates (e.g. Jan 1/2022, 2022/02, Fall 2020) to manually cleaned dates with structured time units (e.g. days, months, quarters, semesters, ranges) and date formats (e.g. 2022-01-01, 2022-02-01).

```{r}

# Join clean dates to include structured date formats
bike_dates <- bike_p2011 %>%
	left_join( # install dates
		vdates_raw %>%
			rename_with(~ paste0("_install_", .x)),
		by = join_by(verify_install_date == `_install_verify_date_raw`)
	) %>%
	left_join( # upgrade1 dates
		vdates_raw %>%
			rename_with(~ paste0("_upgrade1_", .x)),
		by = join_by(verify_upgrade1_date == `_upgrade1_verify_date_raw`)
	) %>%
	left_join( # upgrade2 dates
		vdates_raw %>%
			rename_with(~ paste0("_upgrade2_", .x)),
		by = join_by(verify_upgrade2_date == `_upgrade2_verify_date_raw`)
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_p2011 %>% nrow,
	"\n* Columns: ", bike_p2011 %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_dates %>% nrow,
	"\n* Columns: ", bike_dates %>% ncol,
	"\n\nNew Columns (", bike_dates %>% ncol - bike_p2011 %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_dates) %>%
			.[!. %in% colnames(bike_p2011)],
		collapse = ""
	)
))

# Display counts
bike_dates_counts <- bike_dates %>%
    as_tibble %>%
    pivot_longer(
        c(
            verify_install_year,
            verify_upgrade1_year,
            verify_upgrade2_year
        ),
        names_to = "status",
        values_to = "status_year"
    ) %>%
    mutate(
        status = case_when(
            str_starts(status, "verify_install") ~ "install",
            str_starts(status, "verify_upgrade1") ~ "upgrade1",
            str_starts(status, "verify_upgrade2") ~ "upgrade2"
        )
    ) %>%
    group_by(city, status) %>%
    summarize(
        n = sum(!is.na(status_year))
    ) %>%
	left_join(
		bike_dates %>%
		    as_tibble %>%
		    pivot_longer(
		        c(
		            `_install_verify_date_type`,
		            `_upgrade1_verify_date_type`,
		            `_upgrade2_verify_date_type`
		        ),
		        names_to = "status",
		        values_to = "status_date_type"
		    ) %>%
		    mutate(
		        status = case_when(
		            str_starts(status, "_install") ~ "install",
		            str_starts(status, "_upgrade1") ~ "upgrade1",
		            str_starts(status, "_upgrade2") ~ "upgrade2"
		        )
		    ) %>%
		    group_by(city, status) %>%
		    summarize(
		        n_with_dates = sum(!is.na(status_date_type))
		    ),
		by = c("city", "status")
	) %>%
	mutate(
		n_without_dates = n - n_with_dates
	) %>%
	select(
		city,
		status,
		n_with_dates,
		n_without_dates,
		n
	)
bike_dates_counts

```

### Preview

```{r}

# Display clean dates cols
bike_dates %>%
	as_tibble %>%
	select(starts_with(c("_install_", "_upgrade1_", "_upgrade2_")), everything()) %>%
	arrange(`_install_verify_date_type`)

```

## Add Time Units {.tabset .tabset-pills}

Add the following time unit columns for installs/upgrades in bikeways and for each ksi record:

* `_time_month`: which month (1, 2, 3 ... 12) a bikeway had a verified installation/upgrade or ksi record
* `_time_quarter`: which quarter (1, 2, 3, 4) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_third`: which third (1, 2, 3) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_half`: which half (1, 2) of the year a bikeway had a verified installation/upgrade or ksi record
* `_time_semester`: which semester (1, 2) of the year a bikeway had a verified installation/upgrade or ksi record, where:
	* `1`: represents May to October of this year
	* `2`: represents November (this year) to April of next year
* `_time_year`: standardized year column that bikeway had verified installation/upgrade or ksi record
* `_time_..._group`: grouped columns of the above defining the time units in textual format
* Install records follow the patterns above prefixed with `_install`
* Upgrade records are prefixed with `_upgrade1` and `_upgrade2`
* KSI records will follow the same patterns above except prefixed with `_ksi` as well
* Additional columns for installs and upgrades only:
	* `_assign_semester`: if the date falls between semester 1 or 2 and was automatically assigned to semester 1 or 2 based on the larger number of days in either semester 1 or 2
	* `_assign_semester_1_days`: number of days in semester 1 if the date falls between semesters
	* `_assign_semester_2_days`: number of days in semester 2 if the date falls between semesters
	* If the number of days are equal, then semester 2 is assigned by default

**Note**: Date ranges that fall between months, quarters, thirds, or halves of the year were excluded from being classified as any of the time units (e.g. March 31 to April 15 will be excluded as it does not fall within either the 1st or 2nd quarter of the year).

```{r}

#' Add Time Units to Data Frame
#'
#' This function takes a data frame and adds new columns representing various time units such as month, quarter, third, half, and semester. These columns are derived from a date column or if the time unit is more than day, then additiona start and end date columns
#'
#' @param df A data frame that contains the columns `verify_date`, `verify_date_type`, `verify_date_start`, and `verify_date_end`.
#' @param date_col Name of the date column to use
#' @param start_col Name of the start date column to use
#' @param end_col Name of the end date column to use
#' @param type_col Name of the date type column to use (e.g. month, day, year, etc)
#' @param prefix Additional text to add to the output column names
#' @param assign_semester Whether to assign intervals between semesters to be in the semester with the max duration that the interval falls in. When durations are equal, semester 2 will be chosen.
#'
#' @return A data frame with additional columns (without `prefix`):
#' \describe{
#'   \item{`_time_month`}{Numeric representation of the month (1-12).}
#'   \item{`_time_month_group`}{Unique time group for month.}
#'   \item{`_time_quarter`}{Quarter of the year (1-4).}
#'   \item{`_time_quarter_group`}{Unique time group for quarter.}
#'   \item{`_time_third`}{Third of the year (1-3).}
#'   \item{`_time_third_group`}{Unique time group for third.}
#'   \item{`_time_half`}{Half of the year (1-2).}
#'   \item{`_time_half_group`}{Unique time group for half.}
#'   \item{`_time_semester`}{Custom semester (1 for May-Oct, 2 for Nov-Apr).}
#'   \item{`_assign_semester`}{Whether the date was assigned since it did not fall between either semester 1 or 2.}
#'   \item{`_assign_semester_1_days`}{Number of days that the date is in semester 1 if `assign_semester` is `TRUE`.}
#'   \item{`_assign_semester_2_days`}{Number of days that the date is in semester 2 if `assign_semester` is `TRUE`.}
#'   \item{`_time_semester_group`}{Unique time group for semester.}
#'   \item{`_time_year`}{4-digit year.}
#' }
#'
#' @examples
#' \dontrun{
#' df <- data.frame(
#'   verify_date = as.Date(c("2021-01-15", "2021-06-20", "2021-09-10")),
#'   verify_date_type = c("day", "month", "day"),
#'   verify_date_start = as.Date(c("2021-01-01", "2021-06-01", "2021-09-01")),
#'   verify_date_end = as.Date(c("2021-01-31", "2021-06-30", "2021-09-30"))
#' )
#' result <- add_time_units(df)
#' print(result)
#' }
#'
#' @import dplyr
#' @importFrom lubridate month year
#' @export
add_time_units <- function(
	df,
	date_col = "_install_verify_date",
	start_col = "_install_verify_date_start",
	end_col = "_install_verify_date_end",
	type_col = "_install_verify_date_type",
	year_col = "verify_install_year",
	prefix = NULL,
	assign_semester = T
) {
	
	# Code time units
	out <- df %>%
		mutate( # rename req cols
			`_date` = .data[[date_col]],
			`_start` = .data[[start_col]],
			`_end` = .data[[end_col]],
			`_type` = .data[[type_col]],
			`_year` = .data[[year_col]]
		) %>%
		mutate( # add time units
			
			# Months (monthly)
			`_time_month` = case_when(
				`_type` %in% c("day", "month") ~ month(`_date`, label = T, abbr = F)
			),
			`_time_month` = as.numeric(`_time_month`),
			
			# Quarters (quarterly)
			`_time_quarter` = case_when(
				month(`_date`) %in% 1:3 |
				(
					month(`_start`) %in% 1:3 &
					month(`_end`) %in% 1:3 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% 4:6 |
				(
					month(`_start`) %in% 4:6 &
					month(`_end`) %in% 4:6 &
					year(`_start`) == year(`_end`)
				) ~ 2,
				month(`_date`) %in% 7:9 |
				(
					month(`_start`) %in% 7:9 &
					month(`_end`) %in% 7:9 &
					year(`_start`) == year(`_end`)
				) ~ 3,
				month(`_date`) %in% 10:12 |
				(
					month(`_start`) %in% 10:12 &
					month(`_end`) %in% 10:12 &
					year(`_start`) == year(`_end`)
				) ~ 4
			),
			
			# Thirds (triyearly)
			`_time_third` = case_when(
				month(`_date`) %in% 1:4 |
				(
					month(`_start`) %in% 1:4 &
					month(`_end`) %in% 1:4 &
					year(`_start`) == year(`_end`)
				) ~ 1, # Fall
				month(`_date`) %in% 5:8 |
				(
					month(`_start`) %in% 5:8 &
					month(`_end`) %in% 5:8 &
					year(`_start`) == year(`_end`)
				) ~ 2, # Winter
				month(`_date`) %in% 9:12 |
				(
					month(`_start`) %in% 9:12 &
					month(`_end`) %in% 9:12 &
					year(`_start`) == year(`_end`)
				) ~ 3 # Spring/Summer
			),
			
			# Halves (biyearly)
			`_time_half` = case_when(
				month(`_date`) %in% 1:6 |
				(
					month(`_start`) %in% 1:6 &
					month(`_end`) %in% 1:6 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% 7:12 |
				(
					month(`_start`) %in% 7:12 &
					month(`_end`) %in% 7:12 &
					year(`_start`) == year(`_end`)
				) ~ 2
			),
			
			# Semester (custom range)
			`_time_semester` = case_when(
				month(`_date`) %in% 5:10 | # May to Oct (this year)
				(
					month(`_start`) %in% 5:10 &
					month(`_end`) %in% 5:10 &
					year(`_start`) == year(`_end`)
				) ~ 1,
				month(`_date`) %in% c(11:12, 1:4) | # Nov (this/last year) to Apr (next/this year)
				(
					month(`_start`) %in% c(11:12, 1:4) &
					month(`_end`) %in% c(11:12, 1:4) &
					year(`_start`) <= (year(`_end`) + 1)
				) ~ 2
			),
			
			# Year
			`_time_year` = `_year`
			
		)
	
	# For intervals between semesters, assign in max duration
	if (assign_semester) {
		out <- out %>%
			mutate( # add intervals for semesters and date
				`_assign_semester_1_interval_1` = interval( # sem 1 this year
		            start = glue("{`_year`}-05-01") %>% as_date,
		            end = glue("{`_year`}-10-31") %>% as_date
		        ),
		        `_assign_semester_2_interval_1` = interval( # sem 2 this to next year
		            start = glue("{`_year`}-11-01") %>% as_date,
		            end = glue("{`_year` + 1}-04-30") %>% as_date
		        ),
				`_assign_semester_2_interval_2` = interval( # sem 2 last to this year
		            start = glue("{`_year` - 1}-11-01") %>% as_date,
		            end = glue("{`_year`}-04-30") %>% as_date
		        ),
				`_assign_date_interval` = if_else( # only for non-assigned sem
					!is.na(`_start`) & !is.na(`_end`) & is.na(`_time_semester`),
					interval(`_start`, `_end`),
					NA
				)
			) %>%
			mutate( # get duration in days
				`_assign_semester_1_days` = coalesce(
					intersect(
						`_assign_date_interval`,
						`_assign_semester_1_interval_1`
					)
				) %>% as.numeric / (3600 * 24),
				`_assign_semester_2_days` = coalesce(
					intersect(
						`_assign_date_interval`,
						`_assign_semester_2_interval_1`
					),
					intersect(
						`_assign_date_interval`,
						`_assign_semester_2_interval_2`
					)
				) %>% as.numeric / (3600 * 24)
			) %>%
			mutate( # assign semester
				`_time_semester` = if_else(
					!is.na(`_assign_date_interval`),
					case_when(
						`_assign_semester_1_days` == `_assign_semester_2_days` ~ 2, # break ties to sem 2
						`_assign_semester_1_days` > `_assign_semester_2_days` ~ 1, # sem 1
						`_assign_semester_1_days` < `_assign_semester_2_days` ~ 2 # sem 2
					),
					`_time_semester`
				),
				`_assign_semester` = if_else(
					!is.na(`_assign_semester_1_days`) | !is.na(`_assign_semester_2_days`),
					TRUE,
					FALSE
				)
			) %>%
			select(-c( # remove unwanted cols
				`_assign_semester_1_interval_1`,
				`_assign_semester_2_interval_1`,
				`_assign_semester_2_interval_2`,
				`_assign_date_interval`
			))
	}
	
	# Add time groups that include the year
	out <- out %>%
		mutate(
			
			# Month group
			`_time_month_group` = if_else(
				!is.na(`_time_month`),
				glue("{`_year`}_{`_time_month`}"),
				NA
			),
			
			# Quarter group
			`_time_quarter_group` = if_else(
				!is.na(`_time_quarter`),
				glue("{`_year`}_{`_time_quarter`}"),
				NA
			),
			
			# Third group
			`_time_third_group` = if_else(
				!is.na(`_time_third`),
				glue("{`_year`}_{`_time_third`}"),
				NA
			),
			
			# Half group
			`_time_half_group` = if_else(
				!is.na(`_time_half`),
				glue("{`_year`}_{`_time_half`}"),
				NA
			),
			
			# Semester group
			`_time_semester_group` = if_else(
				!is.na(`_time_semester`),
				glue("{`_year`}_{`_time_semester`}"),
				NA
			),
			
			# Year group
			`_time_year_group` = if_else(
				!is.na(`_time_year`),
				glue("{`_year`}"),
				NA
			)
			
		)
	
	# Add prefix to time unit cols
	if (!is.null(prefix)) {
		out <- out %>%
			rename_with( # add prefix
				~ paste0(prefix, .x),
				starts_with("_time_")
			)
	}
	
	# Remove time cols
	out <- out %>%
		select(-c(
			`_date`,
			`_start`,
			`_end`,
			`_type`,
			`_year`
		))
	return(out)
}

```

### Bikeways {.tabset}

```{r}

# Add time unit columns using func for bikeways
bike_tunits <- bike_dates %>%
	add_time_units( # install tunits
		date_col = "_install_verify_date",
		start_col = "_install_verify_date_start",
		end_col = "_install_verify_date_end",
		type_col = "_install_verify_date_type",
		year_col = "verify_install_year",
		prefix = "_install"
	) %>%
	add_time_units( # upgrade1 tunits
		date_col = "_upgrade1_verify_date",
		start_col = "_upgrade1_verify_date_start",
		end_col = "_upgrade1_verify_date_end",
		type_col = "_upgrade1_verify_date_type",
		year_col = "verify_upgrade1_year",
		prefix = "_upgrade1"
	) %>%
	add_time_units( # upgrade2 tunits
		date_col = "_upgrade2_verify_date",
		start_col = "_upgrade2_verify_date_start",
		end_col = "_upgrade2_verify_date_end",
		type_col = "_upgrade2_verify_date_type",
		year_col = "verify_upgrade2_year",
		prefix = "_upgrade2"
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_dates %>% nrow,
	"\n* Columns: ", bike_dates %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_tunits %>% nrow,
	"\n* Columns: ", bike_tunits %>% ncol,
	"\n\nNew Columns (", bike_tunits %>% ncol - bike_dates %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_tunits) %>%
			.[!. %in% colnames(bike_dates)],
		collapse = ""
	)
))

# Display counts
bike_tunits_counts <- bike_tunits %>%
    as_tibble %>%
	pivot_longer(
		ends_with("_group"),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "_install") ~ "install",
			str_starts(column, "_upgrade1") ~ "upgrade1",
			str_starts(column, "_upgrade2") ~ "upgrade2"
		),
		time_unit = str_replace_all(
			column,
			"_install_time_|_upgrade1_time_|_upgrade2_time_|_group",
			""
		)
	) %>%
	filter(!is.na(value)) %>%
	group_by(status, time_unit) %>%
	count
bike_tunits_counts

```

#### Preview

```{r}

# Display time unit cols for bikeways
bike_tunits %>%
	as_tibble %>%
	select(
		starts_with("_"),
		everything()
	) %>%
	arrange(`_install_verify_date`) %>%
	head

```

### KSI {.tabset}

```{r}

# Add time unit columns using func for ksi
ksi_tunits <- ksi_y2009to2023 %>%
	mutate( # add cols to standardize func params
		`_ksi_date` = ksi_date,
		`_ksi_date_start` = NA,
		`_ksi_date_end` = NA,
		`_ksi_date_type` = "day",
		`_ksi_year` = year(ksi_date)
	) %>%
	add_time_units(
		date_col = "_ksi_date",
		start_col = "_ksi_date_start",
		end_col = "_ksi_date_end",
		type_col = "_ksi_date_type",
		year_col = "_ksi_year",
		prefix = "_ksi",
		assign_semester = F
	)

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", ksi_y2009to2023 %>% nrow,
	"\n* Columns: ", ksi_y2009to2023 %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", ksi_tunits %>% nrow,
	"\n* Columns: ", ksi_tunits %>% ncol,
	"\n\nNew Columns (", ksi_tunits %>% ncol - ksi_y2009to2023 %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(ksi_tunits) %>%
			.[!. %in% colnames(ksi_combine)],
		collapse = ""
	)
))

# Display counts
ksi_tunits_counts <- ksi_tunits %>%
    as_tibble %>%
	pivot_longer(
		ends_with("_group"),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		time_unit = str_replace_all(
			column,
			"_ksi_time_|_group",
			""
		)
	) %>%
	filter(!is.na(value)) %>%
	group_by(time_unit) %>%
	count
ksi_tunits_counts

```

#### Preview

```{r}

# Display time unit cols for ksi
ksi_tunits %>%
	as_tibble %>%
	select(
		ksi_date,
		starts_with("_")
	) %>%
	arrange(year(ksi_date)) %>%
	head

```

## Determine Time Unit {.tabset .tabset-pills}

For bikeways, determine the temporal resolution (unit of time) finer than yearly based on the amount of data available per time unit (sorted from the highest resolution time unit to the lowest resolution time unit):

* **Month** (high): refers to verified post-2011 installs and upgrades occurring approximately in a month (1/12) of a year
* **Quarter**: refers to verified post-2011 installs and upgrades occurring approximately in a quarter (1/4) of a year
* **Third**: refers to verified post-2011 installs and upgrades occurring approximately in a third (1/3) of a year
* **Half** (low): refers to verified post-2011 installs and upgrades occurring approximately in a half (1/2) of a year
* **Semester** (low): refers to verified post-2011 installs and upgrades occurring approximately in a half (1/2) of a year with the first half (1) being November (same year) to April (next year) and the second half being May to October (next year)

```{r}

plot_time_units <- function(
		df,
		month_col = "_install_time_month",
		quarter_col = "_install_time_quarter",
		third_col = "_install_time_third",
		half_col = "_install_time_half",
		semester_col = "_install_time_semester",
		title = "% of Verified Post-2011 Installs by Unit of Time",
		overlap_eq_lag = 3,
		overlap_eq_lead = 3,
		overlap_small_lag = 2,
		overlap_small_lead = 2,
		overlap_lag = 1.5,
		overlap_lead = 1.5,
		expand = c(0.18, 0),
		city_nudge_x = 0.25
) {
	
	# Prepare plot data
	plot_data <- df %>%
		as_tibble %>%
		mutate( # assign time unit cols
			`_time_month` = .data[[month_col]],
			`_time_quarter` = .data[[quarter_col]],
			`_time_third` = .data[[third_col]],
			`_time_half` = .data[[half_col]],
			`_time_semester` = .data[[semester_col]]
		) %>%
		group_by(city) %>%
		summarize( # calc installs/upgrades for each time unit
			Month = sum(!is.na(`_time_month`)),
			Quarter = sum(!is.na(`_time_quarter`)),
			Third = sum(!is.na(`_time_third`)),
			Half = sum(!is.na(`_time_half`)),
			Semester = sum(!is.na(`_time_semester`))
		) %>%
		pivot_longer(
			cols = -city,
			names_to = "type",
			values_to = "n"
		) %>%
		mutate( # zero as NA
			n = if_else(n == 0, NA, n)
		) %>%
		left_join( # Add city totals
			df %>%
				as_tibble %>%
				group_by(city) %>%
				count(name = "total"),
			by = "city"
		) %>%
		ungroup %>%
	    mutate( # calc percentages and add labels
	        type = factor(type, levels = c("Month", "Quarter", "Third", "Half", "Semester")),
	        city = factor(str_to_title(city), levels = c("Vancouver", "Calgary", "Toronto")),
	        perc = n / total * 100,
	        perc_label = glue(
	        	"{str_sub(city, end = 1)}: {round(perc, 2)}%\n",
	        	"(n={format(n, big.mark = ',', scientific = F)})"
	        )
	    ) %>%
	    group_by(type) %>% # adjust overlapping labels
	    arrange(desc(perc)) %>% 
	    mutate( # detect overlap labels and shift down and up
	        perc_label_y = case_when(
	        	lag(perc) == perc ~ perc - overlap_eq_lag,
	            lead(perc) == perc ~ perc + overlap_eq_lead,
	        	lag(perc) - perc < 2 ~ perc - overlap_small_lag,
	            lag(perc) - perc < 7 ~ perc - overlap_lag,
	            lead(perc) - perc > -2 ~ perc + overlap_small_lead,
	        	lead(perc) - perc > -7 ~ perc + overlap_lead,
	            .default = perc
	        )
	    ) %>%
		group_by(city) %>%
		mutate( # add city labels at end of lines
			city_label = if_else(
				type == "Semester",
				glue(
					"{city}\n",
					"(n={format(total, big.mark = ',', scientific = F)})"
				),
				NA
			)
		) %>%
		ungroup %>%
	    arrange(city, type)
	
	# Get time unit totals
	unit_total <- sum(plot_data$total %>% unique, na.rm = T)
	unit_total_label <- format(unit_total, big.mark = ",", scientific = F)
	
	# Plot time units
	out <- plot_data %>%
		ggplot(aes(
			x = type,
			y = perc,
			group = city,
			color = city,
			label = perc_label
		)) +
		geom_line(
			alpha = 0.8,
			linewidth = 1.5
		) +
		geom_label(
			aes(y = perc_label_y),
			size = 2.25,
			show.legend = F,
			label.padding = unit(0.5, "lines")
		) +
		geom_text(
			aes(
				label = city_label,
				y = perc_label_y
			),
			size = 2.5,
			show.legend = F,
			hjust = 0,
			nudge_x = city_nudge_x
		) +
		scale_y_continuous(labels = function(x) paste0(x, "%")) +
		scale_x_discrete(
			limits = levels(plot_data$type),
			position = "top",
			expand = expand
		) +
		labs(
			title = glue(
				"{title}\n",
				"(n={unit_total_label})"
			),
			x = NULL,
			y = NULL,
			color = "City",
			label = NULL,
			group = NULL
		) +
		theme_minimal() +
		theme(
			legend.position = "none",
			plot.title = element_text(hjust = 0.5)
		)
	return(out)
}

```

### Install {.tabset}

#### Plot

```{r}
plot_time_units(bike_tunits)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	group_by(city, `_install_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

### 1st Upgrade {.tabset}

#### Plot

```{r}
plot_time_units(
	bike_tunits %>% filter(!is.na(verify_upgrade1_year)),
	month_col = "_upgrade1_time_month",
	quarter_col = "_upgrade1_time_quarter",
	third_col = "_upgrade1_time_third",
	half_col = "_upgrade1_time_half",
	semester_col = "_upgrade1_time_semester",
	title = "% of Verified Post-2011 1st Upgrades by Unit of Time"
)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	filter(!is.na(verify_upgrade1_year)) %>%
	group_by(city, `_upgrade1_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

### 2nd Upgrade {.tabset}

#### Plot

```{r}
plot_time_units(
	bike_tunits %>% filter(!is.na(verify_upgrade2_year)),
	month_col = "_upgrade2_time_month",
	quarter_col = "_upgrade2_time_quarter",
	third_col = "_upgrade2_time_third",
	half_col = "_upgrade2_time_half",
	semester_col = "_upgrade2_time_semester",
	title = "% of Verified Post-2011 2nd Upgrades by Unit of Time",
	overlap_eq_lag = 6.25,
	overlap_eq_lead = -1.75,
	overlap_small_lag = 1.75,
	overlap_small_lead = 1.75,
	overlap_lag = 1,
	overlap_lead = 1,
	expand = c(0.2, 0),
	city_nudge_x = 0.35
)
```

#### Details

```{r}
bike_tunits %>%
	as_tibble %>%
	filter(!is.na(verify_upgrade2_year)) %>%
	group_by(city, `_upgrade1_time_semester`) %>%
	count %>%
	group_by(city) %>%
	mutate(perc = round(n / sum(n) * 100, 2))
```

## Filter Semester Bikeways {.tabset}

Filter for bikeways with semesters that were installed/upgraded post-2011.

```{r}
bike_sem <- bike_tunits %>%
	filter(
		(verify_install_year > 2011 & !is.na(`_install_time_semester`)) |
		(verify_upgrade1_year > 2011 & !is.na(`_upgrade1_time_semester`)) |
		(verify_upgrade2_year > 2011 & !is.na(`_upgrade1_time_semester`)) 
	)
```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_tunits %>% nrow,
	"\n* Columns: ", bike_tunits %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_sem%>% nrow,
	"\n* Columns: ", bike_sem %>% ncol
))

# Display counts
bike_sem_counts <- bike_sem %>%
    as_tibble %>%
	pivot_longer(
		c(
			verify_install_year,
			verify_upgrade1_year,
			verify_upgrade2_year
		),
		names_to = "column",
		values_to = "value"
	) %>%
	mutate(
		status = case_when(
			str_starts(column, "verify_install") ~ "install",
			str_starts(column, "verify_upgrade1") ~ "upgrade1",
			str_starts(column, "verify_upgrade2") ~ "upgrade2"
		)
	) %>%
	group_by(city, status, value) %>%
	count
bike_sem_counts

```

### Preview

```{r}

# Display merged bike segments
bike_sem %>%
	as_tibble %>%
	select(-geometry) %>%
	head

```

## Merge Bikeways {.tabset}

Merge bikeway segments if they are within the same infrastructure type, time unit, and known status (e.g. install, 1st upgrade, 2nd upgrade).

As `semester` seemed to have the most available data, it was chosen as the time unit.

The known status is the most recent change **with a known time unit** for a segment with one of the following values:

1. `install`: an installed segment with no upgrades
2. `upgrade1` a segment upgraded once and was previously installed
3. `upgrade2`: a segment upgraded twice and was previously upgrade once and installed

This also adds an additional column `_merge_group` specifying the merge group in the format of `<CITYNUM>_<STATUS>_<TYPE>_<YEAR>_<TIME>_<ADJID>`, where:

* `<CITYNUM>`: is the numeric id of the city, 1 is vancouver, 2 is calgary, and 3 is toronto
* `<STATUS>`: the status of the segment, one of `install`, `upgrade1`, or `upgrade2`
* `<TYPE>`: the infrastructure type of the segment
* `<YEAR>`: the year of the segment status
* `<TIME>`: the value of the time unit
* `<ADJID>`: the adjacency id, if the ids are the same, they are adjacent in space

Other merge related columns start with `_merge_...`.

```{r}

# Select time unit
tunit <- "semester"

# Determine adjacent segments if they are within same status, type, and time unit
bike_adj <- bike_sem %>%
	filter( # remove segments that do not have the time unit
		!is.na(.data[[glue("_upgrade2_time_{tunit}")]]) |
		!is.na(.data[[glue("_upgrade1_time_{tunit}")]]) |
		!is.na(.data[[glue("_install_time_{tunit}")]])
	) %>%
	mutate( # add column to store status, type, and time unit
		`_merge_status` = case_when(
			!is.na(.data[[glue("_upgrade2_time_{tunit}")]]) ~ "upgrade2",
			!is.na(.data[[glue("_upgrade1_time_{tunit}")]])~ "upgrade1",
			!is.na(.data[[glue("_install_time_{tunit}")]]) ~ "install"
		),
		`_merge_type` = case_when(
			`_merge_status` == "upgrade2" ~ verify_upgrade2_type,
			`_merge_status` == "upgrade1" ~ verify_upgrade1_type,
			`_merge_status` == "install" ~ verify_install_type
		),
		`_merge_time_unit` = tunit,
		`_merge_time` = case_when(
			`_merge_status` == "upgrade2" ~ .data[[glue("_upgrade2_time_{tunit}")]],
			`_merge_status` == "upgrade1" ~ .data[[glue("_upgrade1_time_{tunit}")]],
			`_merge_status` == "install" ~ .data[[glue("_install_time_{tunit}")]]
		),
		`_merge_time_group` = case_when(
			`_merge_status` == "upgrade2" ~ .data[[glue("_upgrade2_time_{tunit}_group")]],
			`_merge_status` == "upgrade1" ~ .data[[glue("_upgrade1_time_{tunit}_group")]],
			`_merge_status` == "install" ~ .data[[glue("_install_time_{tunit}_group")]]
		),
		`_merge_year` = case_when(
			`_merge_status` == "upgrade2" ~ `_upgrade2_time_year`,
			`_merge_status` == "upgrade1" ~ `_upgrade1_time_year`,
			`_merge_status` == "install" ~ `_install_time_year`
		)
	) %>%
	group_by(city, `_merge_status`, `_merge_type`, `_merge_time_group`) %>%
	group_map(~ {
		
		# Assign group id as group_by values separated by underscore
		group_id <- .y[1,] %>%
			as.character %>%
			paste0(., collapse = "_") %>%
			str_to_lower(.)
		
		# Find adjacent segments and group them together using a group id
		adj_group <- .x %>%
			st_touches %>%
			graph_from_adj_list %>%
			components %>%
			.$membership %>%
			paste0(group_id, "_", .)
		
		# Assign adj group to data in order
		out <- .x %>% mutate(`_merge_group` = adj_group)
		return(out)
		
	}, .keep = T) %>%
	reduce(add_row)

# Merge bike segments by adjacency
bike_merge <- bike_adj %>%
	group_by(`_merge_group`) %>%
	summarize(
		id = paste0(id, collapse = ","),
		city = paste0(unique(city), collapse = ","),
		status = case_when( # assign single status if possible
			all(unique(`_merge_status`) == "upgrade2") ~ "upgrade2",
			all(unique(`_merge_status`) == "upgrade1") ~ "upgrade1",
			all(unique(`_merge_status`) == "install") ~ "install",
			.default = paste(unique(`_merge_status`), collapse = ",") # multiple status
		),
		type = if_else( # assign single type if possible
			length(unique(`_merge_type`)) == 1,
			unique(`_merge_type`) %>% as.character, # single type
			paste(unique(`_merge_type`), collapse = ",") # multiple types
		),
		year = if_else( # assign single year if possible
			length(unique(`_merge_year`)) == 1,
			unique(`_merge_year`) %>% as.character, # single year
			paste(unique(`_merge_year`), collapse = ",") # multiple years
		),
		time_unit = tunit,
		time_value = if_else( # assign single time if possible
			length(unique(`_merge_time`)) == 1,
			unique(`_merge_time`) %>% as.character, # single time
			paste(unique(`_merge_time`), collapse = ",") # multiple times
		),
		time_group = if_else( # assign single time if possible
			length(unique(`_merge_time_group`)) == 1,
			unique(`_merge_time_group`) %>% as.character, # single time
			paste(unique(`_merge_time_group`), collapse = ",") # multiple times
		)
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_sem %>% nrow,
	"\n* Columns: ", bike_sem %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nNew Columns (", bike_merge %>% ncol - bike_sem %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_merge) %>%
			.[!. %in% colnames(bike_sem)],
		collapse = ""
	)
))

# Display counts
bike_merge_counts <- bike_merge %>%
    as_tibble %>%
	group_by(city, status) %>%
	count
bike_merge_counts

```

### Preview

```{r}

# Display merged bike segments
bike_merge %>%
	as_tibble %>%
	select(-geometry) %>%
	head

```

### Map

```{r}
tmap_mode("view")
tm_shape(bike_merge %>% st_buffer(25)) +
    tm_polygons(
    	group = "Merged (Black)",
    	col = "black",
    	border.col = "white",
    	popup.vars = T,
    	lwd = 1.5
    ) +
	tm_shape(bike_adj) +
	tm_lines(
		group = "Segments (Red)",
		col = "red",
		lwd = 1.5
	) +
	tm_shape(bike_adj %>% st_cast("POINT")) +
	tm_dots(group = NULL, col = "red", size = 0.03)
```


## Join KSI to Bikeways {.tabset}

Spatially join KSI to nearest bikeways.

Each KSI point is joined to a bikeway if they are within 25 meters of the nearest bikeway

**Note**: if there is overlapping of bikeways, the same KSI can be assigned to multiple bikeways.

```{r}

# Set cache file for spatiotemp join
sjoin_date <- max(c(
	vancbike_dldate,
	calgbike_dldate,
	toronbike_dldate,
	toronksi_dldate
))
sjoin_file <- glue("../data/cache/sjoin-{sjoin_date}.csv")

# Run spatial join if not cached
if (!file.exists(sjoin_file)) {
	
	# Buffer bikeways and keep ids
	bike_merge_buff25 <- bike_merge %>%
		select(city, id) %>%
		st_buffer(25)
	
	# Spatial join bikeways with ksi keeping ids
	sjoin <- bike_merge_buff25 %>%
		st_join(ksi_tunits %>% select(ksi_city, ksi_id)) %>%
		as_tibble %>%
		select(-geometry)
	
	# Cache spatial joined bike and ksi ids
	sjoin %>% write_csv(sjoin_file)
	
} else {
	
	# Read spatial joined bike and ksi ids from cache
	sjoin <- read_csv(sjoin_file, col_types = cols(id = "c"))
	
}

```

#### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBikeways\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nKSI\n",
	"\n* Rows: ", ksi_tunits %>% nrow,
	"\n* Columns: ", ksi_tunits %>% ncol,
	"\n\nSpatial Join Keys\n",
	"\n* Rows: ", sjoin %>% nrow,
	"\n* Columns: ", sjoin %>% ncol,
	"\n* Rows Spatially Joined (Bikeways): ",
	sjoin %>%
		filter(!is.na(id)) %>%
		pull(id) %>%
		unique %>%
		length,
	"\n* Rows Spatially Joined (KSI): ",
	sjoin %>%
		filter(!is.na(ksi_id)) %>%
		pull(ksi_id) %>%
		unique %>%
		length
))

# Display 
sjoin_counts <- sjoin %>%
	group_by(city) %>%
	summarize(
		n_bike = n_distinct(id, na.rm = T)
	) %>%
	left_join(
		sjoin %>%
			group_by(ksi_city) %>%
			summarize(
				n_ksi = n_distinct(ksi_id, na.rm = T)
			),
		by = join_by(city == ksi_city)
	)
sjoin_counts

```

#### Preview

```{r}

# Display spatially joined ids
sjoin %>%
	filter(!is.na(ksi_id)) %>%
	head

```

## Count KSI Per Time Unit {.tabset}

For each merged bikeway, count KSI spatially joined to bikeways per unit of time.

This creates columns in the format of `_ksi_time_<UNIT>_<YEAR>_<VALUE>`, where:

* `<UNIT>`: is the unit of time, either `month`, `quarter`, `half`, or `semester`
* `<YEAR>`: is the year for the unit of time
* `<VALUE>`: is the number representing the portion of the time range for the unit of time with month being 1-12, quarter being 1-4, third being 1-3, half being 1-2 and semester being 1-2 (this is omitted if time unit is `year`)

```{r}

# Ensure ids are same data type
sjoin <- sjoin %>%
	mutate(
		id = as.character(id),
		ksi_id = as.character(ksi_id)
	)
ksi_tunits <- ksi_tunits %>%
	mutate(
		ksi_id = as.character(ksi_id)
	)

# Get time unit col
tunit_col <- glue("_ksi_time_{tunit}_group")

# Count ksi spatially joined to bike
bike_sjoin <- bike_merge %>%
	mutate(
		id = as.character(id)
	) %>%
	full_join(
		sjoin,
		by = join_by(city, id)
	) %>%
	full_join(
		ksi_tunits %>%
			as_tibble %>%
			select(-geometry),
		by = join_by(ksi_city, ksi_id)
	) %>%
	group_by(across(c(
		"city",
		"id",
		tunit_col
	))) %>%
	summarize(
		across(everything(), ~ paste0(unique(.x), collapse = ",")),
		`ksi_count` = n()
	)

# Create combination of all id and time unit combos
tunit_combo <- expand_grid(
    `_merge_group` = unique(bike_merge$`_merge_group`),
    year = min(year(ksi_tunits$ksi_date)):max(year(ksi_tunits$ksi_date)),
    time_value = unique(ksi_tunits[[glue("_ksi_time_{tunit}")]])
) %>%
    mutate(
    	!!tunit_col := glue("{year}_{time_value}")
    ) %>%
    select(-year, -time_value)

# Fill in missing time units as zero
bike_ksi <- tunit_combo %>%
	left_join(
		bike_merge,
		by = c("_merge_group")
	) %>%
	left_join(
		bike_sjoin %>%
			as_tibble %>%
			select(c(
				"_merge_group",
				tunit_col,
				"ksi_count",
				colnames(ksi_tunits)
			)) %>%
			select(-geometry),
		by = c("_merge_group", tunit_col)
	) %>%
	mutate(
		ksi_count = replace_na(ksi_count, 0)
	)

```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_merge %>% nrow,
	"\n* Columns: ", bike_merge %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", bike_ksi %>% nrow,
	"\n* Columns: ", bike_ksi %>% ncol,
	"\n\nNew Columns (", bike_ksi %>% ncol - bike_merge %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(bike_ksi) %>%
			.[!. %in% colnames(bike_merge)],
		collapse = ""
	)
))

```

### Preview

```{r}

# Display merged bikeways with ksi counts
bike_ksi %>%
	filter(!is.na(ksi_id)) %>%
	head

```

## Format Model Data

```{r}

# Extract year and semester for time ids
tid_data <- bike_ksi %>%
	separate(
		`_ksi_time_semester_group`,
		c("year", "semester"),
		sep = "_"
	) %>%
	select(year, semester)

# Assign time ids using ordered levels
tid <- expand_grid(
	time_year = min(tid_data$year):max(tid_data$year),
	time_semester = unique(tid_data$semester)
) %>%
    arrange(time_year, time_semester) %>%
    mutate(
    	time_orig = glue("{time_year}_{time_semester}"),
    	time = factor(time_orig) %>% as.integer
    )

# Prepare model data
model_data <- bike_ksi %>%
	select(
		bike_merge_id = `_merge_group`,
		bike_city = city,
		bike_status = status,
		bike_type = type,
		bike_time_orig = time_group,
		bike_year = year,
		bike_semester = time_value,
		ksi_time_orig = `_ksi_time_semester_group`,
		ksi = `ksi_count`
	) %>%
	separate( # get year and semester
		ksi_time_orig,
		c("ksi_year", "ksi_semester"),
		sep = "_",
		remove = F
	) %>%
	mutate( # assign ids
		id = factor(bike_merge_id) %>% as.integer
	) %>%
	left_join( # assign ksi time ids
		tid %>%
			select(
				ksi_time_orig = time_orig,
				time
			),
		by = "ksi_time_orig"
	) %>%
	left_join( # assign cohort based on bike time ids
		tid %>%
			select(
				bike_time_orig = time_orig,
				cohort = time
			),
		by = "bike_time_orig"
	) %>%
	mutate( # assign cohort based on pl, ct, and upgrade
		cohort_pl = if_else(bike_type %in% c("PL", "BUF"), cohort, NA),
		cohort_ct = if_else(bike_type == "PBL", cohort, NA),
		cohort_upgrade = if_else(str_starts(bike_status, "upgrade"), cohort, NA)
	) %>%
	mutate( # assign post based on whether it is greater than bike time
		post = as.integer(time >= cohort),
		post_pl = as.integer(time >= cohort_pl) %>% replace_na(0),
		post_ct = as.integer(time >= cohort_ct) %>% replace_na(0),
		post_upgrade = as.integer(time >= cohort_upgrade) %>% replace_na(0)
	) %>%
	arrange(id, time) %>%
	filter(bike_city == "toronto") %>%
	select(
		starts_with("bike_"),
		starts_with("ksi_"),
		id,
		time,
		starts_with("post"),
		ksi,
		starts_with("cohort"),
		everything()
	)

# Export model data and time ids to folder
model_data %>% write_csv("../data/recovr-eval-data-2024-08-07.csv")
```

### Details

```{r}

# Display rows and cols
cat(glue(
	"\nBefore\n",
	"\n* Rows: ", bike_ksi %>% nrow,
	"\n* Columns: ", bike_ksi %>% ncol,
	"\n\nAfter\n",
	"\n* Rows: ", model_data %>% nrow,
	"\n* Columns: ", model_data %>% ncol,
	"\n\nNew Columns (", model_data %>% ncol - bike_ksi %>% ncol, ")\n",
	paste0(
		"\n* ",
		colnames(model_data) %>%
			.[!. %in% colnames(bike_ksi)],
		collapse = ""
	)
))

```

#### Preview

```{r}

# Display model data
model_data %>%
	head

```

## Flow Chart

```{r}

# Create flow chart
fig_flowchart <- grViz('

digraph {

	layout=neato
	splines=ortho
	
	node [shape="box", width=2.75, height=1 margin=0.1]
	
	toronbike_raw[width=2, shape=cylinder, label="Toronto Bikeways\n(n=@@1)", pos="0,0!"]
	toronbike_verify[label="Verified Bikeways Only\n(n=@@2)", pos="3,0!"]
	toronbike_itypes[label="Painted Lanes and\nCycle Tracks Only\n(n=@@3)", pos="6,0!"]
	toronbike_p2011[label="Post-2011 Installs\nor Upgrades Only\n(n=@@4)", pos="0,-1.5!"]
	toronbike_sem[label="Installs or Upgrades\nwith Semesters Only\n(n=@@5)", pos="3,-1.5!"]
	toronbike_merge[label="Spatially Merged Bikeways\nby Status, Type, and Semester\n(n=@@6)", pos="6,-1.5!"]
	
	toronksi_raw[width=2, shape=cylinder, label="Toronto KSI\n(n=@@7)", pos="0,-3!"]
	toronksi_y2009to2023[label="From 2009 to 2023 Only\n(n=@@8)", pos="3,-3!"]
	toronksi_sjoin[label="Spatially Joined to\nMerged Bikeways\n(n_bikeways=@@6, n_ksi=@@9)", pos="6,-3!"]
	
	toronbike_raw -> toronbike_verify -> toronbike_itypes -> toronbike_p2011 -> toronbike_sem -> toronbike_merge

	toronbike_merge -> toronksi_sjoin
	toronksi_raw -> toronksi_y2009to2023 [minlen=3]
	toronksi_y2009to2023 -> toronksi_sjoin [minlen=3]
}

[1]: toronbike_raw %>% nrow
[2]: bike_verify %>% filter(city == "toronto") %>% nrow
[3]: bike_itypes %>% filter(city == "toronto") %>% nrow
[4]: bike_p2011 %>% filter(city == "toronto") %>% nrow
[5]: bike_sem %>% filter(city == "toronto" & (!is.na(`_install_time_semester`) | !is.na(`_upgrade1_time_semester`) | !is.na(`_upgrade2_time_semester`))) %>% nrow
[6]: bike_merge %>% filter(city == "toronto") %>% nrow
[7]: ksi_combine %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ",")
[8]: ksi_y2009to2023 %>% filter(ksi_city == "toronto") %>% nrow %>% format(big.mark = ",")
[9]: sjoin %>% filter(ksi_city == "toronto" & !is.na(ksi_id)) %>% pull(ksi_id) %>% unique %>% length

')

# Export flowchart
fig_flowchart %>%
    export_svg %>%
    charToRaw %>%
    rsvg_png("../figures/fig-methods-toronto.png")

# Display flowchart
fig_flowchart
```
